### 2015-08-14 ms.tex
abstract

%\documentclass{emulateapj}
\documentclass[letterpaper,12pt,preprint]{aastex}

% packages
\usepackage{amssymb,amsmath,amsbsy}
\usepackage{booktabs}
\usepackage[caption=false]{subfig}

% commands
\newcommand{\given}{\,|\,}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\transpose}[1]{{#1}^{\mathsf{T}}}
\newcommand{\inverse}[1]{{#1}^{-1}}
\newcommand{\msun}{\mathrm{M}_\odot}

\begin{document}

\title{Will it blend: Is the Milky Way bar causing chaotic fanning of the Ophiuchus stream?}
\author{Adrian M. Price-Whelan\altaffilmark{\colum,\adrn}, +others}

% Affiliations
\newcommand{\colum}{1}
\newcommand{\adrn}{2}

\altaffiltext{\colum}{Department of Astronomy,
                      Columbia University,
                      550 W 120th St.,
                      New York, NY 10027, USA}
\altaffiltext{\adrn}{To whom correspondence should be addressed: adrn@astro.columbia.edu}

\begin{abstract}

% Context
The Ophiuchus stream was recently discovered as an over-density of main sequence stars in Pan-STARRS (PS1) photometry and was later detected in BHB stars using line-of-sight velocities: XX of the YY BHB stars around the sky position of the stream appear kinematically distinct from the background halo population.
A small number of stars (XX) near the ends of the stream have similarly distinct velocities but with much larger velocity and spatial dispersions compared to the short segment of the stream first identified in the PS1 data.
In recent work, we have shown that streams formed on chaotic orbits may display similar kinematic signatures---short segments of recently-disrupted stellar debris flanked by high-dispersion, low-density ``stream-fans'' at the ends of the stream.
The stream's proximity to the triaxial, time-dependent potential of the Galactic bar and the peculiar kinematics and morphology of the stream suggest that the stream may have formed on a chaotic orbit.
% Aims
Here we study whether the Galactic bar can lead to chaotic orbits near (in phase-space) to the Ophiuchus stream. 
% Methods
We construct a potential model for the Milky Way bar based on recent measurements of the density of XXX stars.
We then assess the level of chaos for orbits like that of the Ophiuchus stream and generate mock streams along these orbits that qualitatively reproduce the observed kinematics of the stream.
% Results
We find that the apparent shortness and nearby low-density, high-velocity-dispersion ...
% Conclusions
These results motivate efforts to obtain deep photometry of the region surrounding the Ophiuchus stream to detect the predicted low-surface-brightness fans of stellar debris.
The existence of or lack of stream fans may provide an interesting constraint on the triaxiality and pattern speed of the bar.

\end{abstract}

% \keywords{}

\section{Introduction}\label{sec:introduction}

(a short projected length of $\approx$XX deg)

The stream is $\approx$XX kpc from the Sun at a Galactocentric radius and height $(R,z) \approx (XX, YY)~{\rm kpc}$.

much larger than the measured $\lesssim 1~{\rm km}~{\rm s}^{-1}$ dispersion of the

by releasing ensembles of test particles when the progenitor orbit crosses pericenter.

\section{Methods}\label{sec:method}

\section{Potential model}\label{sec:potential}

The three-dimensional shape and density distribution of the inner galaxy is not well known.

\section{Conclusions}\label{sec:conclusions}

\acknowledgements
APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155.
This work was supported in part by the National Science Foundation under Grant No. PHYS-1066293.
This research made use of Astropy, a community-developed core Python package for Astronomy \citep{astropy13}.
This work additionally relied on Columbia University's \emph{Hotfoot} and \emph{Yeti} compute clusters, and we acknowledge the Columbia HPC support staff for assistance. \\

%\bibliographystyle{apj}
%\bibliography{refs}

\end{document}

### 2015-08-14 ms.tex
paper outline

@@
 
+The Ophiuchus stream \citep{bernard14} is located in an interesting region of the Milky Way.
+Many streams have been found in the halo at Galactocentric distances of XX--YY kpc, but Ophiuchus is situated near the Galactic center where the gravitational potential should have significant components from the bulge and bar of the Milky Way.
+Recent modeling fit for the orbit, found discrepancy between dynamical age and age of stellar populations \citep{sesar15}, however they assume an axisymmetric disk plus spherical bulge and halo.
+If the Galactic bar contributes significantly to the density distribution of the inner Galaxy, the orbit of the cluster could be entirely different.
+
 (a short projected length of $\approx$XX deg)
@@
 
-\section{Potential model}\label{sec:potential}
+\subsection{Potential model}\label{sec:potential}
 
 The three-dimensional shape and density distribution of the inner galaxy is not well known.
+
+\subsection{Mock streams}\label{sec:mocks}
+
+\section{Results}
+
+\subsection{The degree of chaos}\label{sec:results1}
+
+\subsection{Reproducing the observed properties of the stream}\label{sec:results2}
 
@@
 APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155.
-This work was supported in part by the National Science Foundation under Grant No. PHYS-1066293.
+APW acknowledges the ... MPIA ...
 This research made use of Astropy, a community-developed core Python package for Astronomy \citep{astropy13}.
-This work additionally relied on Columbia University's \emph{Hotfoot} and \emph{Yeti} compute clusters, and we acknowledge the Columbia HPC support staff for assistance. \\
+This work additionally relied on Columbia University's \emph{Yeti} compute cluster, and we acknowledge the Columbia HPC support staff for assistance. \\
 

### 2015-08-14 ms.tex
adding more section outlines

@@
 
-The Ophiuchus stream \citep{bernard14} is located in an interesting region of the Milky Way.
-Many streams have been found in the halo at Galactocentric distances of XX--YY kpc, but Ophiuchus is situated near the Galactic center where the gravitational potential should have significant components from the bulge and bar of the Milky Way.
-Recent modeling fit for the orbit, found discrepancy between dynamical age and age of stellar populations \citep{sesar15}, however they assume an axisymmetric disk plus spherical bulge and halo.
-If the Galactic bar contributes significantly to the density distribution of the inner Galaxy, the orbit of the cluster could be entirely different.
+* Observational work on the stream (Bernard, Sesar)
 
@@
 
-by releasing ensembles of test particles when the progenitor orbit crosses pericenter.
+* Summarize chaos work
+
+* Why is Ophiuchus interesting?
+
+- Shortness of the stream
+
+- Possible fanned debris
+
+- Discrepancy between dynamical / stellar population age of the stream
+
+The Ophiuchus stream is located in a unique position relative to other known stellar tidal streams: Many streams have been found in the halo of the Milky Way at Galactocentric distances of XX--YY kpc.
+For this stream, it is therefore expected that the baryonic components of the Milky Way---the combination of the bulge, disk, and Galactic bar---should dominate the kinematic history of the progenitor system.
+Unfortunately, the density distribution or gravitational potential of the inner Galaxy is not well known.
+
+Recent modeling fit for the orbit, found discrepancy between dynamical age and age of stellar populations \citep{sesar15}, however they assume an axisymmetric disk plus spherical bulge and halo.
+If the Galactic bar contributes significantly to the density distribution of the inner Galaxy, the orbit of the cluster could be entirely different.
 
@@
 
+Our goal is to assess whether the Galactic bar can produce significant numbers of chaotic orbits in the inner Galaxy and see if the Ophiuchus stream could plausibly be on a strongly chaotic orbit.
+If chaos is important for this stream, we would then like to show that chaotic stream-fanning can reproduce the observed morphology and velocity signatures of ...
+In this section, we describe the methods we will use to detect and quantify the strength of chaos for orbits and then explain the method we will use to generate mock tidal streams.
+
 \subsection{Potential model}\label{sec:potential}
 
-The three-dimensional shape and density distribution of the inner galaxy is not well known.
+\subsection{Mock streams}\label{sec:mocks}
 
-\subsection{Mock streams}\label{sec:mocks}
+by releasing ensembles of test particles when the progenitor orbit crosses pericenter.
 

### 2015-08-21 ms.tex
added to methods

@@
 
+We use ...
+[Bulge from Portail et a. (2015), bar from Wegg and Gerhard (2015)?]
+
+Though these density distributions are analytic and relatively simple, the potential and accelerations must be computed numerically.
+Rather than solve Poisson's equation at every evaluation of the potential or forces, we instead use the Self-Consistent Field (SCF) method \citep{hernquist92} to expand the density and potential on a series of orthogonal potential-density pairs.
+This method has been previously used to represent the potential from the Galactic bar \citep{zhao96, wangzhao12}.
+The advantage to this method is that the expansion coefficients (expensive integrals) only need to be computed once.
+
+We follow the formalism introduced in \cite{hernquist92} and used in \cite{lowing11}:
+
 \subsection{Mock streams}\label{sec:mocks}

### 2015-09-02 ms.tex
text woohoo

--- 2015-08-14 ms.tex
@@
 
-\title{Will it blend: Is the Milky Way bar causing chaotic fanning of the Ophiuchus stream?}
-\author{Adrian M. Price-Whelan\altaffilmark{\colum,\adrn}, +others}
+\title{}
+\author{Adrian M. Price-Whelan\altaffilmark{\colum,\adrn},
+            Branimir Sesar\altaffilmark{\mpia},
+            Sarah Pearson\altaffilmark{\colum},
+            Kathryn V. Johnston\altaffilmark{\colum},
+            Hans-Walter Rix\altaffilmark{\mpia},
+}
 
@@
 \newcommand{\adrn}{2}
+\newcommand{\mpia}{3}
 
@@
 \altaffiltext{\adrn}{To whom correspondence should be addressed: adrn@astro.columbia.edu}
+\altaffiltext{\mpia}{Max-Planck-Institut f\"ur Astronomie,
+                     K\"onigstuhl 17, D-69117 Heidelberg, Germany}
 
@@
 
-The Ophiuchus stream \citep{bernard14} is located in an interesting region of the Milky Way.
-Many streams have been found in the halo at Galactocentric distances of XX--YY kpc, but Ophiuchus is situated near the Galactic center where the gravitational potential should have significant components from the bulge and bar of the Milky Way.
-Recent modeling fit for the orbit, found discrepancy between dynamical age and age of stellar populations \citep{sesar15}, however they assume an axisymmetric disk plus spherical bulge and halo.
-If the Galactic bar contributes significantly to the density distribution of the inner Galaxy, the orbit of the cluster could be entirely different.
+The Ophiuchus stream \citep{bernard14, sesar15a} is a recently discovered stellar tidal stream that sits above the Galactic bulge at a Galactocentric radius and height $(R,z) \approx (1, 4.3)~{\rm kpc}$.
+All observational evidence suggests that the stream is a totally disrupted globular cluster: The stream stars have (1) a small positional dispersion orthogonal to the extended direction of the stream ($\approx$10 pc); (2) a small velocity dispersion $\approx$0.4 ${\rm km}~{\rm s}^{-1}$; (3) an old stellar population ($\approx$12 Gyr) estimated from isochrone fitting; and (4) no detectable over-density along the stream that could be  the progenitor system \citep[][hereafter S15a]{sesar15a}.
 
-(a short projected length of $\approx$XX deg)
+The deprojected length of the visible part of the stream ($\approx$1.5 kpc) is short for the age of its stellar population.
+S15a fit an orbit to the kinematics of the stream stars in a static, axisymmetric model for the gravitational field of the inner Galaxy and ran N-body simulations of globular clusters on this orbit.
+S15a find that---on this orbit---the observed portion of the stream must have been formed in the last $\lesssim$400 Myr for the stream to remain as short as it is observed.
+This dynamical age is at odds with the old stellar population.
+More recently, a significant observational effort has followed-up nearly all BHB star candidates within XX deg of the stream to obtain radial velocities to identify more stream members \citep{sesar15b}.
+The stream has a very distinct and large line-of-sight velocity ($\approx$290 ${\rm km}~{\rm s}^{-1}$) and is therefore easily detected above the background halo population.
+Of the XX stars followed up, six have velocities significantly discrepant with the halo population.
+Of these six stars, two lie within the previously detected extent of the stream and have velocities consistent with the previously identified stream members.
+The remaining four stars lie close to an extrapolation of the stream, have velocities much larger than halo stars in this region ($v_{\rm los}>230 {\rm km}~{\rm s}^{-1}$), but have a velocity dispersion $\approx$75 times the measured internal velocity dispersion of the stream stars and positional spread much larger than the orthogonal positional dispersion of the main part of the stream.
+These low-density, high-dispersion features were not modeled in S15a and N-body simulations on the orbit fit from the previous work do not naturally reproduce these features.
 
-The stream is $\approx$XX kpc from the Sun at a Galactocentric radius and height $(R,z) \approx (XX, YY)~{\rm kpc}$.
+It is not surprising that the orbit fit in the static, axisymmetric potential model used in S15a leads to a dynamical puzzle and failed to predict the new features: it is now well-known that the Galactic bulge contains a triaxial, bar-like structure several kpc in size \citep[e.g.,][]{blitzXX, wegg13, MANY}.
+Given the proximity of the stream to the center of the Galaxy, the time-dependent, triaxial potential of the Galactic bar must be taken into account when modeling the orbit of the progenitor system.
+The presence of a bar-like perturbation to the potential will change the possible orbit of the stream progenitor and will drastically change the resonance structure of orbits in the inner galaxy \citep{athanassoula, MANY}.
+The bar may also induce a significant number of chaotic orbits in this region.
 
-much larger than the measured $\lesssim 1~{\rm km}~{\rm s}^{-1}$ dispersion of the
+[Review of chaos / fanning stuff]
+
+In this work, we study whether stream-fanning---chaotic or simply from density evolution in a triaxial, time-dependent potential---can explain the observed properties of the Ophiuchus stream.
+In Section~\ref{sec:method} we describe the methods used in this work: in Section~\ref{sec:potential} we describe the models we use for the gravitational potential of the Galaxy, in Section~\ref{sec:mocks} we explain the simple method we use to generate mock streams, and in Section~\ref{sec:chaos-indicator} we briefly describe the methods we use for detecting and characterizing chaotic orbits.
+In Section~\ref{sec:results1} we discuss the differences in the orbit structure between a static, axisymmetric potential model and a model with a time-dependent bar potential, and in Section~\ref{sec:results2} we use mock streams to argue that [chaotic stream-fanning is a plausible explanation for the observational peculiarities of the Ophiuchus stream.] We conclude in Section~\ref{sec:conclusions}.
+
+\section{Methods}\label{sec:method}
+
+Our goal is to (1) assess whether the Galactic bar can produce significant numbers of chaotic orbits in the inner Galaxy and (2) see if the Ophiuchus stream could plausibly be on a strongly chaotic orbit.
+In this section, we describe the potential model we use for the galaxy and describe the methods we use to detect and quantify the strength of chaos for orbits.
+
+\subsection{Potential models}\label{sec:potential}
+
+To integrate orbits and to compute chaos indicators we must choose a gravitational potential that represents the potential of the Milky Way. [...]
+
+We use a four-component, analytic model for the potential consisting of a spherical Navarro-Frenk-White halo \citep{nfw96}, a Miyamoto-Nagai disk \citep{miyamoto75}, a Hernquist spheroid \citep{hernquist90}, and
+a basis function expansion representation of the Galactic bar used in \citep{wang12}.
+
+[Bulge from Portail et a. (2015), bar from Wegg and Gerhard (2015)?]
+
+Though these density distributions are analytic and relatively simple, the potential and accelerations must be computed numerically.
+Rather than solve Poisson's equation at every evaluation of the potential or forces, we instead use the Self-Consistent Field (SCF) method \citep{hernquist92} to expand the density and potential on a series of orthogonal potential-density pairs.
+This method has been previously used to represent the potential from the Galactic bar \citep{zhao96, wangzhao12}.
+The advantage to this method is that the expansion coefficients (expensive integrals) only need to be computed once.
+
+We follow the formalism introduced in \cite{hernquist92} and used in \cite{lowing11}:
+
+\subsection{Generating mock streams}\label{sec:mocks}
 
@@
 
-\section{Methods}\label{sec:method}
-
-\subsection{Potential model}\label{sec:potential}
-
-The three-dimensional shape and density distribution of the inner galaxy is not well known.
-
-\subsection{Mock streams}\label{sec:mocks}
+\subsection{Detecting and characterizing chaos with orbital frequency diffusion}\label{sec:chaos-indicator}
 

### 2015-09-03 ms.tex
more text

@@
 \newcommand{\msun}{\mathrm{M}_\odot}
+
+\newcommand{\project}[1]{\textsl{#1}}
+
+% TO DO
+\newcommand{\todo}[2]{{\color{red} TODO: (\MakeUppercase{#1}) #2}}
 
@@
 Of these six stars, two lie within the previously detected extent of the stream and have velocities consistent with the previously identified stream members.
-The remaining four stars lie close to an extrapolation of the stream, have velocities much larger than halo stars in this region ($v_{\rm los}>230 {\rm km}~{\rm s}^{-1}$), but have a velocity dispersion $\approx$75 times the measured internal velocity dispersion of the stream stars and positional spread much larger than the orthogonal positional dispersion of the main part of the stream.
-These low-density, high-dispersion features were not modeled in S15a and N-body simulations on the orbit fit from the previous work do not naturally reproduce these features.
+The remaining four stars lie close to an extrapolation of the stream, have velocities much larger than halo stars in this region ($v_{\rm los}>230~{\rm km}~{\rm s}^{-1}$), but have a velocity dispersion $\approx$75 times the measured internal velocity dispersion of the stream stars and positional spread much larger than the orthogonal positional dispersion of the main part of the stream.
+These low-density, high-dispersion features were not modeled in S15a and N-body simulations on the orbit fit from the previous work do not  reproduce these features.
 
@@
 Given the proximity of the stream to the center of the Galaxy, the time-dependent, triaxial potential of the Galactic bar must be taken into account when modeling the orbit of the progenitor system.
-The presence of a bar-like perturbation to the potential will change the possible orbit of the stream progenitor and will drastically change the resonance structure of orbits in the inner galaxy \citep{athanassoula, MANY}.
-The bar may also induce a significant number of chaotic orbits in this region.
+The presence of a bar-like perturbation to the potential will change the possible orbit of the stream progenitor and will drastically change the resonance structure of orbits in the inner galaxy \citep{athanassoula, MANY}, potentially introducing a significant number of chaotic orbits in this region.
 
-[Review of chaos / fanning stuff]
+Recent work has shown that chaos can dramatically alter the density evolution of tidal streams.
+Along certain chaotic orbits, the stream stars will spread much faster in 3D position than from ordinary phase-mixing and, depending on the orbital phase at which the stream is observed, may develop large, low-density blobs of stars at the ends of a stream \citep{apw15-chaos}.
+[...]
 
@@
 
-Our goal is to (1) assess whether the Galactic bar can produce significant numbers of chaotic orbits in the inner Galaxy and (2) see if the Ophiuchus stream could plausibly be on a strongly chaotic orbit.
+Our goal is to (1) assess whether the Galactic bar can produce significant numbers of chaotic orbits in the inner Galaxy; (2) see if the Ophiuchus stream could plausibly be on a strongly chaotic orbit; and (3) if chaotic density evolution can explain the apparent shortness of the stream and low-density, high-dispersion ends.
 In this section, we describe the potential model we use for the galaxy and describe the methods we use to detect and quantify the strength of chaos for orbits.
+We then describe a simplified version of the \emph{Streakline} method \citep{kuepper12} that we use to generate mock streams.
 
@@
 
-To integrate orbits and to compute chaos indicators we must choose a gravitational potential that represents the potential of the Milky Way. [...]
+To integrate orbits and to compute chaos indicators we must choose a gravitational potential model to represent the potential of the Milky Way.
+The key feature of the potential that we would like to capture is the time-dependence and triaxiality of the Galactic bar.
+We construct two parametrized potential models: one with a triaxial, time-dependent bar and one with a purely spherical bulge, both added to simple models for the disk and halo of the Milky Way.
+Recent work has mapped the 3D density distribution of the region around the bulge using number counts of red clump stars and have measured the bar angle and axis ratios of the bar structure \citep{wegg13}.
+The authors do not fit an analytic density model to the data and do not provide enough information about the density field to reconstruct the full-space stellar density distribution of the bar.
+Further, this is not a dynamical model for the bar and therefore may not account for all of the mass but rather provides information about the number density of stars in the Galactic bar.
 
-We use a four-component, analytic model for the potential consisting of a spherical Navarro-Frenk-White halo \citep{nfw96}, a Miyamoto-Nagai disk \citep{miyamoto75}, a Hernquist spheroid \citep{hernquist90}, and
-a basis function expansion representation of the Galactic bar used in \citep{wang12}.
+We instead use an earlier model that uses a basis function expansion (BFE) to represent the potential and density of the bar \citep[][hereafter W12]{wang12}.
+The authors provide a table of expansion coefficients for a low-order expansion of the triaxial bar density used in \citet{dwek95}.
+The BFE formalism used in W12 neglects the sine expansion terms because of the symmetries of the triaxial density; the coefficients presented in W12 are for just the cosine terms (the $A_{lm}$ in \citet{hernquist92} or the $S_{nlm}$ in \citet{lowing11}).
+We have implemented the BFE computation of the potential, density, and gradient of the potential in C and Python and the code is publicly available on \project{GitHub}.\footnote{\url{https://github.com/adrn/biff}} The BFE representation effectively fixes the axis ratios of the bar (the exponential scale lengths along the three axes of the bar were adopted from \cite{dwek95} when the expansion coefficients were calculated in W12), but the mass of the bar, pattern speed, and angle with respect to the Galactic $x$ axis are free parameters.
 
-[Bulge from Portail et a. (2015), bar from Wegg and Gerhard (2015)?]
+In both models, we use a spherical Navarro-Frenk-White potential for the halo \citep{nfw96}, a Miyamoto-Nagai potential for the disk \citep{miyamoto75}, and a Hernquist spheroid for the central spherical component of the bulge \citep{hernquist90}.
+Model A consists of just these three components and the total potential is therefore static and axisymmetric.
+Models B and C include the BFE bar potential described above: Model B contains a light bar and Model C contains a heavy bar.
+Except for the pattern speed and bar angle (which are varied), the parameter values for all potential models are given in Table~\ref{tbl:potential-params}.
+In all cases, the parameter values were chosen to qualitatively match the rotation curve of \citet{bovy12}
 
-Though these density distributions are analytic and relatively simple, the potential and accelerations must be computed numerically.
-Rather than solve Poisson's equation at every evaluation of the potential or forces, we instead use the Self-Consistent Field (SCF) method \citep{hernquist92} to expand the density and potential on a series of orthogonal potential-density pairs.
-This method has been previously used to represent the potential from the Galactic bar \citep{zhao96, wangzhao12}.
-The advantage to this method is that the expansion coefficients (expensive integrals) only need to be computed once.
+\begin{table*}[ht]
+\begin{center}
+	\begin{tabular}{ c  c | c  c}
+	         \toprule
+	         \multicolumn{2}{c|}{\bf Model A} & \multicolumn{2}{c}{\bf Model B} \\
+	         Parameter & Value & Parameter & Value \\\toprule
+		{\it Disk} &  &  & \\
+		a & b & c & d\\
+		a & b & c & d\\\midrule
+		a & b & c & d\\
+		\bottomrule
+		\end{tabular}
+	\caption{ \label{tbl:potential}}
+\end{center}
+\end{table*}
 
-We follow the formalism introduced in \cite{hernquist92} and used in \cite{lowing11}:
+\subsection{Detecting and characterizing chaos with orbital frequency diffusion}\label{sec:chaos-indicator}
 
@@
 by releasing ensembles of test particles when the progenitor orbit crosses pericenter.
-
-\subsection{Detecting and characterizing chaos with orbital frequency diffusion}\label{sec:chaos-indicator}
 
@@
 APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155.
-APW acknowledges the ... MPIA ...
+APW acknowledges the staff at the MPIA for...
 This research made use of Astropy, a community-developed core Python package for Astronomy \citep{astropy13}.
@@
 
-%\bibliographystyle{apj}
-%\bibliography{refs}
+\bibliographystyle{apj}
+\bibliography{refs}
 

### 2015-09-04 ms.tex
text

@@
 Recent work has mapped the 3D density distribution of the region around the bulge using number counts of red clump stars and have measured the bar angle and axis ratios of the bar structure \citep{wegg13}.
-The authors do not fit an analytic density model to the data and do not provide enough information about the density field to reconstruct the full-space stellar density distribution of the bar.
+The authors do not fit an analytic density model to the data and do not provide enough information about the density field to reconstruct the full-space stellar density distribution of the bulge.
 Further, this is not a dynamical model for the bar and therefore may not account for all of the mass but rather provides information about the number density of stars in the Galactic bar.
+\cite{portail15}
 

### 2015-09-05 ms.tex
potential description

--- 2015-09-03 ms.tex
@@
 \usepackage[caption=false]{subfig}
+\usepackage{multirow}
 
@@
 \newcommand{\inverse}[1]{{#1}^{-1}}
-\newcommand{\msun}{\mathrm{M}_\odot}
+\newcommand{\msun}{\ensuremath{\mathrm{M}_\odot}}
 
@@
 Given the proximity of the stream to the center of the Galaxy, the time-dependent, triaxial potential of the Galactic bar must be taken into account when modeling the orbit of the progenitor system.
-The presence of a bar-like perturbation to the potential will change the possible orbit of the stream progenitor and will drastically change the resonance structure of orbits in the inner galaxy \citep{athanassoula, MANY}, potentially introducing a significant number of chaotic orbits in this region.
+The presence of a bar-like perturbation to the potential will change the possible orbit of the stream progenitor and will drastically change the resonance structure of orbits in the inner galaxy \citep{athanassoula, portail15b MANY}, potentially introducing a significant number of chaotic orbits in this region \citep{weinberg15}.
 
@@
 The key feature of the potential that we would like to capture is the time-dependence and triaxiality of the Galactic bar.
-We construct two parametrized potential models: one with a triaxial, time-dependent bar and one with a purely spherical bulge, both added to simple models for the disk and halo of the Milky Way.
-Recent work has mapped the 3D density distribution of the region around the bulge using number counts of red clump stars and have measured the bar angle and axis ratios of the bar structure \citep{wegg13}.
-The authors do not fit an analytic density model to the data and do not provide enough information about the density field to reconstruct the full-space stellar density distribution of the bar.
-Further, this is not a dynamical model for the bar and therefore may not account for all of the mass but rather provides information about the number density of stars in the Galactic bar.
+We construct two parametrized potential models: one with a triaxial, time-dependent (rotating) bulge component and one with a purely spherical bulge, both added to simple models for the disk and halo of the Milky Way.
+These potentials are meant to be representative rather than definitive models as the global potential is not known.
 
-We instead use an earlier model that uses a basis function expansion (BFE) to represent the potential and density of the bar \citep[][hereafter W12]{wang12}.
-The authors provide a table of expansion coefficients for a low-order expansion of the triaxial bar density used in \citet{dwek95}.
-The BFE formalism used in W12 neglects the sine expansion terms because of the symmetries of the triaxial density; the coefficients presented in W12 are for just the cosine terms (the $A_{lm}$ in \citet{hernquist92} or the $S_{nlm}$ in \citet{lowing11}).
-We have implemented the BFE computation of the potential, density, and gradient of the potential in C and Python and the code is publicly available on \project{GitHub}.\footnote{\url{https://github.com/adrn/biff}} The BFE representation effectively fixes the axis ratios of the bar (the exponential scale lengths along the three axes of the bar were adopted from \cite{dwek95} when the expansion coefficients were calculated in W12), but the mass of the bar, pattern speed, and angle with respect to the Galactic $x$ axis are free parameters.
+In both models, we use a spherical Navarro-Frenk-White potential to represent the dark matter \citep{nfw96}, a Miyamoto-Nagai potential for the disk \citep{miyamoto75}, and a Hernquist spheroid for the central spherical component of the bulge \citep{hernquist90}.
+Model A consists of just these three components and the total potential is therefore static and axisymmetric.
+Model B includes a component to represent the Galactic bar.
+For the bar component, we use a basis function expansion (BFE) of the potential and density of the bar with expansion coefficients derived for a triaxial, exponential bar density \citep[][hereafter W12]{wang12}.
+We use the pre-computed expansion coefficients used in W12, which were computed from a low-order expansion of the triaxial bar density used in \citet{dwek95}.\footnote{The coefficients presented in W12 are for just the cosine terms (the $A_{lm}$ in \citet{hernquist92} or the $S_{nlm}$ in \citet{lowing11}) because all sine terms have zero coefficients for a triaxial density function.} We have implemented the BFE computation of the potential, density, and gradient of the potential in C and Python and the code is publicly available on \project{GitHub}.\footnote{\url{https://github.com/adrn/biff}}
 
-In both models, we use a spherical Navarro-Frenk-White potential for the halo \citep{nfw96}, a Miyamoto-Nagai potential for the disk \citep{miyamoto75}, and a Hernquist spheroid for the central spherical component of the bulge \citep{hernquist90}.
-Model A consists of just these three components and the total potential is therefore static and axisymmetric.
-Models B and C include the BFE bar potential described above: Model B contains a light bar and Model C contains a heavy bar.
-Except for the pattern speed and bar angle (which are varied), the parameter values for all potential models are given in Table~\ref{tbl:potential-params}.
-In all cases, the parameter values were chosen to qualitatively match the rotation curve of \citet{bovy12}
+The BFE representation fixes the axis ratios of the bar---that is, the the exponential scale lengths along the three axes of the bar were adopted from \cite{dwek95} when the expansion coefficients were calculated in W12)---and the pattern speed and angle with respect to the Galactic $x$ axis are left as free parameters; all other potential parameter values are given in Table~\ref{tbl:potential-params}.
+In both Model A and Model B, the mass of the halo is fixed and the mass of the disk, spheroid, and bar are varied in order to qualitatively reproduce the flatness and amplitude of the circular velocity of the Milky Way \citep{bovy12}.
+Figure~\ref{fig:circ-vel} shows the circular velocity curves and decompositions for each component for models A and B.
+Figure~\ref{fig:density} shows isodensity contours of the face-on (left) and edge-on (right) surface density for the two potential models \citep[compare to, e.g., Figure 3 in][]{portail15}.
 
@@
 \begin{center}
-	\begin{tabular}{ c  c | c  c}
+	\begin{tabular}{ c | c | c | c}
 	         \toprule
-	         \multicolumn{2}{c|}{\bf Model A} & \multicolumn{2}{c}{\bf Model B} \\
-	         Parameter & Value & Parameter & Value \\\toprule
-		{\it Disk} &  &  & \\
-		a & b & c & d\\
-		a & b & c & d\\\midrule
-		a & b & c & d\\
+	         Component & Parameter & {\bf Model A} & {\bf Model B}\\\toprule
+		Disk & $M_{\rm disk}$ &  & \\
+		& $a$ &  & \\
+		& $b$ &  &  \\\midrule
+	         Halo & $v_h$ &  & \\
+		& $r_s$ &  & \\\midrule
+		Spheroid & $M_{\rm sph}$ &  & \\
+		& $c$ &  & \\\midrule
+		Bar & $M_{\rm bar}$ & -- & $1.8 \times 10^{10}~\msun$ \\
 		\bottomrule
 		\end{tabular}
-	\caption{ \label{tbl:potential}}
+	\caption{The disk potential scale lengths ($a$, $b$) were adopted following \citep{bovy15-galpy} to match the exponential scale length of the disk \citep{bovyrix13} and local dark-matter density \citep[e.g.,][]{bovytremaine12}.
+The halo, NFWpotential is parametrized as $\Phi(r) = -v_h^2\,\frac{\ln{(1 + r/r_s)}}{r/r_s}$.
+The bar mass for Model B is taken from recent 3D density modeling of red clump stars in the Galactic bulge \citep{portail15}.
+\label{tbl:potential}}
 \end{center}
@@
 
+% Recent work has mapped the 3D density distribution of the region around the bulge using number counts of red clump stars and have measured the bar angle and axis ratios of the bar structure \citep{wegg13}.
+The authors do not fit an analytic density model to the data and do not provide enough information about the density field to reconstruct the full-space stellar density distribution of the bulge.
+Further, this is not a dynamical model for the bar and therefore may not account for all of the mass but rather provides information about the number density of stars in the Galactic bar.
+\cite{portail15}
+
 \acknowledgements
@@
 APW acknowledges the staff at the MPIA for...
+The authors wish to acknowledge Victor Debattista, ...
 This research made use of Astropy, a community-developed core Python package for Astronomy \citep{astropy13}.

### 2015-09-05 ms.tex
refs

@@
 
-In both models, we use a spherical Navarro-Frenk-White potential to represent the dark matter \citep{nfw96}, a Miyamoto-Nagai potential for the disk \citep{miyamoto75}, and a Hernquist spheroid for the central spherical component of the bulge \citep{hernquist90}.
+In both models, we use a spherical Navarro-Frenk-White potential to represent the dark matter \citep{navarro96}, a Miyamoto-Nagai potential for the disk \citep{miyamoto75}, and a Hernquist spheroid for the central spherical component of the bulge \citep{hernquist90}.
 Model A consists of just these three components and the total potential is therefore static and axisymmetric.

### 2015-09-11 ms.tex
figures

--- 2015-09-05 ms.tex
@@
 \newcommand{\msun}{\ensuremath{\mathrm{M}_\odot}}
+\newcommand{\kms}{\ensuremath{\mathrm{km}~\mathrm{s}^{-1}}}
 
@@
 
-\title{}
+%\title{Tapping the keg: the Galactic bar peculiarity of the Ophiuchus stream and the Galactic bar}
+\title{Tapping the keg in the Galactic bar}
 \author{Adrian M. Price-Whelan\altaffilmark{\colum,\adrn},
@@
             Kathryn V. Johnston\altaffilmark{\colum},
-            Hans-Walter Rix\altaffilmark{\mpia},
+            Hans-Walter Rix\altaffilmark{\mpia}
 }
@@
 The key feature of the potential that we would like to capture is the time-dependence and triaxiality of the Galactic bar.
-We construct two parametrized potential models: one with a triaxial, time-dependent (rotating) bulge component and one with a purely spherical bulge, both added to simple models for the disk and halo of the Milky Way.
-These potentials are meant to be representative rather than definitive models as the global potential is not known.
+Recent work has used stellar number counts of XX stars in the Galactic bulge to constrain dynamical models of the bar \citep{portail15}.
+Measurements of the total mass of the bar feature from this study are largely consistent with past work \citep[e.g.,][]{XX}, however the measured pattern speed and present bar angle are significantly discrepant and this difference is not fully understood.
+We therefore construct a parametrized potential model consisting of a triaxial, time-dependent (rotating) bulge component added to simple models for the disk and halo of the Milky Way.
+We describe below how we fix the parameters of the disk and halo, but explore different choices for the time-dependence and orientation of the bar.
+This potential model is meant to be representative rather than definitive as the potential of the Milky Way in the inner and outer regions of the galaxy are not well known.
 
-In both models, we use a spherical Navarro-Frenk-White potential to represent the dark matter \citep{nfw96}, a Miyamoto-Nagai potential for the disk \citep{miyamoto75}, and a Hernquist spheroid for the central spherical component of the bulge \citep{hernquist90}.
-Model A consists of just these three components and the total potential is therefore static and axisymmetric.
-Model B includes a component to represent the Galactic bar.
+We use a spherical Navarro-Frenk-White potential to represent the dark matter halo \citep{navarro96} (parametrized as $\Phi(r) = -v_h^2\,\frac{\ln{(1 + r/r_s)}}{r/r_s}$) and a Miyamoto-Nagai potential for the disk \citep{miyamoto75}.
 For the bar component, we use a basis function expansion (BFE) of the potential and density of the bar with expansion coefficients derived for a triaxial, exponential bar density \citep[][hereafter W12]{wang12}.
@@
 
-The BFE representation fixes the axis ratios of the bar---that is, the the exponential scale lengths along the three axes of the bar were adopted from \cite{dwek95} when the expansion coefficients were calculated in W12)---and the pattern speed and angle with respect to the Galactic $x$ axis are left as free parameters; all other potential parameter values are given in Table~\ref{tbl:potential-params}.
-In both Model A and Model B, the mass of the halo is fixed and the mass of the disk, spheroid, and bar are varied in order to qualitatively reproduce the flatness and amplitude of the circular velocity of the Milky Way \citep{bovy12}.
-Figure~\ref{fig:circ-vel} shows the circular velocity curves and decompositions for each component for models A and B.
-Figure~\ref{fig:density} shows isodensity contours of the face-on (left) and edge-on (right) surface density for the two potential models \citep[compare to, e.g., Figure 3 in][]{portail15}.
+The BFE representation fixes the axis ratios of the bar---that is, the the exponential scale lengths along the three axes of the bar were adopted from \cite{dwek95} when the expansion coefficients were calculated in W12)---but the pattern speed and angle with respect to the Galactic $x$ axis are left as free parameters; all other potential parameter values are given in Table~\ref{tbl:potential-params}.
+The mass of the halo is fixed and the mass of the disk and bar are varied in order to qualitatively reproduce the flatness and amplitude of the circular velocity curve of the Milky Way \citep{bovy12}.
+Figure~\ref{fig:circ-vel} shows the circular velocity along the line connecting the Sun to the Galactic center in this model (the Galactic $x$ axis).
+Figure~\ref{fig:density} shows isodensity contours of the face-on (left) and edge-on (right) surface density for the model \citep[compare to, e.g., Figure 3 in][]{portail15}.
 
@@
 \begin{center}
-	\begin{tabular}{ c | c | c | c}
+	\begin{tabular}{ c | c | c }
 	         \toprule
-	         Component & Parameter & {\bf Model A} & {\bf Model B}\\\toprule
-		Disk & $M_{\rm disk}$ &  & \\
-		& $a$ &  & \\
-		& $b$ &  &  \\\midrule
-	         Halo & $v_h$ &  & \\
-		& $r_s$ &  & \\\midrule
-		Spheroid & $M_{\rm sph}$ &  & \\
-		& $c$ &  & \\\midrule
-		Bar & $M_{\rm bar}$ & -- & $1.8 \times 10^{10}~\msun$ \\
+	         Component & Parameter & Value \\\toprule
+		Disk & $M_{\rm disk}$ & $5 \times 10^{10}~\msun$ \\
+		& $a$ & 3~{\rm kpc}\\
+		& $b$ & 0.28~{\rm kpc} \\\midrule
+	         Halo & $v_h$ & 185.8~\kms\\
+		& $r_s$ & 30~kpc \\\midrule
+%		Spheroid & $M_{\rm sph}$ & \\
+%		& $c$ &  \\\midrule
+		Bar & $M_{\rm bar}$ & $1.8 \times 10^{10}~\msun$ \\
 		\bottomrule
@@
 	\caption{The disk potential scale lengths ($a$, $b$) were adopted following \citep{bovy15-galpy} to match the exponential scale length of the disk \citep{bovyrix13} and local dark-matter density \citep[e.g.,][]{bovytremaine12}.
-The halo, NFWpotential is parametrized as $\Phi(r) = -v_h^2\,\frac{\ln{(1 + r/r_s)}}{r/r_s}$.
-The bar mass for Model B is taken from recent 3D density modeling of red clump stars in the Galactic bulge \citep{portail15}.
-\label{tbl:potential}}
+The bar mass is taken from recent 3D density modeling of red clump stars in the Galactic bulge \citep{portail15}.
+\label{tbl:potential-params}}
 \end{center}
@@
 
+% ---------------------------------------------------------------------------------
+\clearpage
+\begin{figure}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/circ-vel.pdf}
+\caption{} 
+\label{fig:circ-vel}
+\end{center}
+\end{figure}
+
+% ---------------------------------------------------------------------------------
+\clearpage
+\begin{figure}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/surface-density.png}
+\caption{} 
+\label{fig:surface-density}
+\end{center}
+\end{figure}
+
 \end{document}

### 2015-10-03 ms.tex
changed figures

@@
 % TO DO
+\usepackage{graphicx} 
 \newcommand{\todo}[2]{{\color{red} TODO: (\MakeUppercase{#1}) #2}}
@@
 %\title{Tapping the keg: the Galactic bar peculiarity of the Ophiuchus stream and the Galactic bar}
-\title{Tapping the keg in the Galactic bar}
+\title{Tapping the keg: Chaos induced by the Galactic bar (working title)}
 \author{Adrian M. Price-Whelan\altaffilmark{\colum,\adrn},
-            Branimir Sesar\altaffilmark{\mpia},
-            Sarah Pearson\altaffilmark{\colum},
-            Kathryn V. Johnston\altaffilmark{\colum},
-            Hans-Walter Rix\altaffilmark{\mpia}
+            Branimir Sesar\altaffilmark{\mpia}, ...
+%            Sarah Pearson\altaffilmark{\colum},
+%            Kathryn V. Johnston\altaffilmark{\colum},
+%            Hans-Walter Rix\altaffilmark{\mpia}
 }
@@
 Recent work has shown that chaos can dramatically alter the density evolution of tidal streams.
-Along certain chaotic orbits, the stream stars will spread much faster in 3D position than from ordinary phase-mixing and, depending on the orbital phase at which the stream is observed, may develop large, low-density blobs of stars at the ends of a stream \citep{apw15-chaos}.
-[...]
+Along certain chaotic orbits, the stream stars will spread much faster in 3D position than from ordinary phase-mixing and, depending on the orbital phase at which the stream is observed, may develop large, low-density ``fans'' of stars at the ends of a stream \citep{apw15-chaos}.
+In this work, we study whether stream-fanning---chaotic or simply from density evolution in a triaxial, time-dependent potential---can explain the observed properties of the Ophiuchus stream.
 
-In this work, we study whether stream-fanning---chaotic or simply from density evolution in a triaxial, time-dependent potential---can explain the observed properties of the Ophiuchus stream.
 In Section~\ref{sec:method} we describe the methods used in this work: in Section~\ref{sec:potential} we describe the models we use for the gravitational potential of the Galaxy, in Section~\ref{sec:mocks} we explain the simple method we use to generate mock streams, and in Section~\ref{sec:chaos-indicator} we briefly describe the methods we use for detecting and characterizing chaotic orbits.
@@
 
-We use a spherical Navarro-Frenk-White potential to represent the dark matter halo \citep{navarro96} (parametrized as $\Phi(r) = -v_h^2\,\frac{\ln{(1 + r/r_s)}}{r/r_s}$) and a Miyamoto-Nagai potential for the disk \citep{miyamoto75}.
+\subsubsection{Barred potential}
+We use a flattened Navarro-Frenk-White potential to represent the dark matter halo \citep{navarro96} parametrized as 
+\begin{align}
+	\Phi(r) &= -v_h^2\,\frac{\ln{(1 + r/r_s)}}{r/r_s}\\
+	r^2 &= x^2 + y^2 + \frac{z^2}{q_z^2}
+\end{align}
+and a Miyamoto-Nagai potential for the disk \citep{miyamoto75}.
 For the bar component, we use a basis function expansion (BFE) of the potential and density of the bar with expansion coefficients derived for a triaxial, exponential bar density \citep[][hereafter W12]{wang12}.
@@
 The mass of the halo is fixed and the mass of the disk and bar are varied in order to qualitatively reproduce the flatness and amplitude of the circular velocity curve of the Milky Way \citep{bovy12}.
-Figure~\ref{fig:circ-vel} shows the circular velocity along the line connecting the Sun to the Galactic center in this model (the Galactic $x$ axis).
-Figure~\ref{fig:density} shows isodensity contours of the face-on (left) and edge-on (right) surface density for the model \citep[compare to, e.g., Figure 3 in][]{portail15}.
+Figure~\ref{fig:circ-vel-barred} shows the circular velocity along the line connecting the Sun to the Galactic center in this model (the Galactic $x$ axis).
+Figure~\ref{fig:surface-density-barred} shows contours of constant surface density for a face-on (left) and edge-on (right) view of this potential model \citep[compare to, e.g., Figure 3 in][]{portail15}.
 
@@
 	         Halo & $v_h$ & 185.8~\kms\\
-		& $r_s$ & 30~kpc \\\midrule
+		& $r_s$ & 30~kpc \\
+		& $q_z$ & 0.9 \\\midrule
 %		Spheroid & $M_{\rm sph}$ & \\
@@
 
-\subsection{Detecting and characterizing chaos with orbital frequency diffusion}\label{sec:chaos-indicator}
+\subsubsection{Static potential}
+
+For comparison, we also define a time-independent potential model with a purely spherical bulge.
+
+Figure~\ref{fig:circ-vel-barred} shows the circular velocity along the line connecting the Sun to the Galactic center in this model (the Galactic $x$ axis).
+Figure~\ref{fig:surface-density-barred} shows contours of constant surface density for a face-on (left) and edge-on (right) view of this potential model \citep[compare to, e.g., Figure 3 in][]{portail15}.
+
+\subsection{Orbit grid initial conditions}\label{sec:orbit-grids}
+
+\subsection{Detecting and characterizing chaos with frequency diffusion}\label{sec:chaos-indicator}
+
+A regular orbit can be characterized by a set of fundamental frequencies.
 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/circ-vel.pdf}
-\caption{} 
-\label{fig:circ-vel}
+\includegraphics[width=\textwidth]{figures/barred-circ-vel}
+\caption{Circular velocity curve along the Sun-Galactic center line for the barred MW potential model.
+Solid black line shows total (sum of all components), lines below show a decomposition by potential component.
+Vertical grey bar shows approximate position of the Sun, horizontal grey bar shows roughly the range in measured circular velocity of the Sun.}
+\label{fig:circ-vel-barred}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/surface-density.png}
-\caption{} 
-\label{fig:surface-density}
+\includegraphics[width=\textwidth]{figures/barred-surface-density-contour}
+\caption{Contours of constant surface density for two projections of the barred MW potential model.
+For both face-on view (left panel) and edge-on view (right panel), four contours are drawn per decade in surface density between $10^7$ and $10^{12}~\msun~{\rm kpc}^{-3}$.
+Note the perturbation from the bar potential within Galactocentric radius $r \lesssim 4~{\rm kpc}$.
+The Sun's position is indicated by the `$\odot$' symbol.}
+\label{fig:surface-density-barred}
+\end{center}
+\end{figure}
+
+% ---------------------------------------------------------------------------------
+\clearpage
+\begin{figure}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/static-circ-vel}
+\caption{Same as Figure~\ref{fig:circ-vel-barred} for the static, axisymmetric MW potential model. }
+\label{fig:circ-vel-static}
+\end{center}
+\end{figure}
+
+% ---------------------------------------------------------------------------------
+\clearpage
+\begin{figure}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/static-surface-density-contour}
+\caption{Same as Figure~\ref{fig:surface-density-barred} for the static, axisymmetric MW potential model. }
+\label{fig:surface-density-static}
 \end{center}

### 2015-10-03 ms.tex
fixed table

@@
 \subsubsection{Barred potential}
-We use a flattened Navarro-Frenk-White potential to represent the dark matter halo \citep{navarro96} parametrized as 
+We use a flattened Navarro-Frenk-White potential to represent the dark matter halo \citep[e.g.,][]{kuepper15} parametrized as 
 \begin{align}
@@
 
-The BFE representation fixes the axis ratios of the bar---that is, the the exponential scale lengths along the three axes of the bar were adopted from \cite{dwek95} when the expansion coefficients were calculated in W12)---but the pattern speed and angle with respect to the Galactic $x$ axis are left as free parameters; all other potential parameter values are given in Table~\ref{tbl:potential-params}.
+The BFE representation fixes the axis ratios of the bar---that is, the the exponential scale lengths along the three axes of the bar were adopted from \cite{dwek95} when the expansion coefficients were calculated in W12)---but the pattern speed and angle with respect to the Galactic $x$ axis are left as free parameters; all other potential parameter values are given in Table~\ref{tbl:potential-params-barred}.
 The mass of the halo is fixed and the mass of the disk and bar are varied in order to qualitatively reproduce the flatness and amplitude of the circular velocity curve of the Milky Way \citep{bovy12}.
@@
 The bar mass is taken from recent 3D density modeling of red clump stars in the Galactic bulge \citep{portail15}.
-\label{tbl:potential-params}}
+Halo flattening is taken from recent work that suggests the inner Milky Way halo is slightly oblate \citep{koposov10, kuepper15}.
+\label{tbl:potential-params-barred}}
 \end{center}
@@
 For comparison, we also define a time-independent potential model with a purely spherical bulge.
+In this model, we set the bar mass to 0 and instead add a spheroidal component represented with a Hernquist potential \citep{hernquist90}.
+Parameters for this potential model are given in Table~\ref{tbl:potential-params-static}.
+Figure~\ref{fig:circ-vel-barred} shows the circular velocity along the line connecting the Sun to the Galactic center in this model (the Galactic $x$ axis).
+Figure~\ref{fig:surface-density-static} shows contours of constant surface density for a face-on (left) and edge-on (right) view of this potential model.
 
-Figure~\ref{fig:circ-vel-barred} shows the circular velocity along the line connecting the Sun to the Galactic center in this model (the Galactic $x$ axis).
-Figure~\ref{fig:surface-density-barred} shows contours of constant surface density for a face-on (left) and edge-on (right) view of this potential model \citep[compare to, e.g., Figure 3 in][]{portail15}.
+\begin{table*}[ht]
+\begin{center}
+	\begin{tabular}{ c | c | c }
+	         \toprule
+	         Component & Parameter & Value \\\toprule
+		Disk & $M_{\rm disk}$ & $6.5 \times 10^{10}~\msun$ \\
+		& $a$ & 3~{\rm kpc}\\
+		& $b$ & 0.28~{\rm kpc} \\\midrule
+	         Halo & $v_h$ & 185.8~\kms\\
+		& $r_s$ & 30~kpc \\
+		& $q_z$ & 0.9 \\\midrule
+		Spheroid & $M_{\rm sph}$ & $10^{10}~\msun$ \\
+		& $c$ & 0.2 \\
+		\bottomrule
+		\end{tabular}
+	\caption{Same as Table~\ref{tbl:potential-params-barred}, except: the disk mass is increased to account for removing the bar component, a spheroidal bulge component is added.
+\label{tbl:potential-params-static}}
+\end{center}
+\end{table*}
 

### 2015-10-07 ms.tex
Words

@@
- %\documentclass{emulateapj}
+% \documentclass{emulateapj}
 \documentclass[letterpaper,12pt,preprint]{aastex}
+
+\input{../paper-helpers/preamble.tex}
 
@@
 
-% commands
-\newcommand{\given}{\,|\,}
-\newcommand{\dd}{\mathrm{d}}
-\newcommand{\transpose}[1]{{#1}^{\mathsf{T}}}
-\newcommand{\inverse}[1]{{#1}^{-1}}
-\newcommand{\msun}{\ensuremath{\mathrm{M}_\odot}}
-\newcommand{\kms}{\ensuremath{\mathrm{km}~\mathrm{s}^{-1}}}
-
-\newcommand{\project}[1]{\textsl{#1}}
-
 % TO DO
 \usepackage{graphicx} 
-\newcommand{\todo}[2]{{\color{red} TODO: (\MakeUppercase{#1}) #2}}
+\usepackage{color}
+\newcommand{\todo}[1]{{\color{red} TODO: #1}}
 
@@
 %\title{Tapping the keg: the Galactic bar peculiarity of the Ophiuchus stream and the Galactic bar}
-\title{Tapping the keg: Chaos induced by the Galactic bar (working title)}
+\title{Spending too much time at the Galactic bar: did the Ophiuchus stream form on a chaotic orbit?}
 \author{Adrian M. Price-Whelan\altaffilmark{\colum,\adrn},
@@
 
-The Ophiuchus stream \citep{bernard14, sesar15a} is a recently discovered stellar tidal stream that sits above the Galactic bulge at a Galactocentric radius and height $(R,z) \approx (1, 4.3)~{\rm kpc}$.
-All observational evidence suggests that the stream is a totally disrupted globular cluster: The stream stars have (1) a small positional dispersion orthogonal to the extended direction of the stream ($\approx$10 pc); (2) a small velocity dispersion $\approx$0.4 ${\rm km}~{\rm s}^{-1}$; (3) an old stellar population ($\approx$12 Gyr) estimated from isochrone fitting; and (4) no detectable over-density along the stream that could be  the progenitor system \citep[][hereafter S15a]{sesar15a}.
+The Ophiuchus stream \citep{bernard14, sesar15} is a recently discovered stellar tidal stream that sits above the Galactic bulge at a Galactocentric radius and height $(R,z) \approx (1, 4.3)~{\rm kpc}$.
+All observational evidence suggests that the stream is a totally disrupted globular cluster: The stream stars have (1) a small positional dispersion orthogonal to the extended direction of the stream ($\approx$10 pc); (2) a small velocity dispersion $\approx$0.4 ${\rm km}~{\rm s}^{-1}$; (3) an old stellar population ($\approx$12 Gyr) estimated from isochrone fitting; and (4) no detectable over-density along the stream that could be  the progenitor system \citep[][hereafter S15a]{sesar15}.
 
@@
 More recently, a significant observational effort has followed-up nearly all BHB star candidates within XX deg of the stream to obtain radial velocities to identify more stream members \citep{sesar15b}.
-The stream has a very distinct and large line-of-sight velocity ($\approx$290 ${\rm km}~{\rm s}^{-1}$) and is therefore easily detected above the background halo population.
-Of the XX stars followed up, six have velocities significantly discrepant with the halo population.
+The stream has a very distinct and large line-of-sight velocity ($\approx$290 \kms) and is therefore easily detected above the background halo population.
+Of the \todo{XX} stars followed up, six have velocities significantly discrepant with the halo population.
 Of these six stars, two lie within the previously detected extent of the stream and have velocities consistent with the previously identified stream members.
-The remaining four stars lie close to an extrapolation of the stream, have velocities much larger than halo stars in this region ($v_{\rm los}>230~{\rm km}~{\rm s}^{-1}$), but have a velocity dispersion $\approx$75 times the measured internal velocity dispersion of the stream stars and positional spread much larger than the orthogonal positional dispersion of the main part of the stream.
+The remaining four stars lie close to an extrapolation of the stream, have velocities much larger than halo stars in this region ($v_{\rm los}>230$ \kms), but have a velocity dispersion $\approx$75 times the measured internal velocity dispersion of the stream stars and positional spread much larger than the orthogonal positional dispersion of the main part of the stream.
 These low-density, high-dispersion features were not modeled in S15a and N-body simulations on the orbit fit from the previous work do not  reproduce these features.
@@
 
-Our goal is to (1) assess whether the Galactic bar can produce significant numbers of chaotic orbits in the inner Galaxy; (2) see if the Ophiuchus stream could plausibly be on a strongly chaotic orbit; and (3) if chaotic density evolution can explain the apparent shortness of the stream and low-density, high-dispersion ends.
-In this section, we describe the potential model we use for the galaxy and describe the methods we use to detect and quantify the strength of chaos for orbits.
-We then describe a simplified version of the \emph{Streakline} method \citep{kuepper12} that we use to generate mock streams.
+Our goal is to (1) assess whether the Galactic bar can produce significant numbers of chaotic orbits in the inner Galaxy in the vicinity of the Ophiuchus stream and (2) if chaotic density evolution of tidal debris stripped from the progenitor of this stream can explain the apparent shortness of the stream and low-density, high-dispersion ends.
+In this section, we describe the potential models we use to represent the galaxy and describe the methods we use to detect and quantify the strength of chaos for orbits.
+We then describe a simplified version of the \streakline\ method that we use to generate mock stellar streams \citep{kuepper12}.
 
@@
 The key feature of the potential that we would like to capture is the time-dependence and triaxiality of the Galactic bar.
-Recent work has used stellar number counts of XX stars in the Galactic bulge to constrain dynamical models of the bar \citep{portail15}.
+Recent work has used stellar number counts of \todo{XX} stars in the Galactic bulge to constrain dynamical models of the bar \citep{portail15}.
 Measurements of the total mass of the bar feature from this study are largely consistent with past work \citep[e.g.,][]{XX}, however the measured pattern speed and present bar angle are significantly discrepant and this difference is not fully understood.
-We therefore construct a parametrized potential model consisting of a triaxial, time-dependent (rotating) bulge component added to simple models for the disk and halo of the Milky Way.
-We describe below how we fix the parameters of the disk and halo, but explore different choices for the time-dependence and orientation of the bar.
-This potential model is meant to be representative rather than definitive as the potential of the Milky Way in the inner and outer regions of the galaxy are not well known.
+We construct a parametrized potential model consisting of a triaxial, time-dependent (rotating) bulge component added to simple models for the disk and halo of the Milky Way.
+We describe below how we fix the parameters of the disk, halo, and bar or bulge component, but explore different choices for the time-dependence and orientation of the bar.
+We also define a static potential with a spherical bulge for comparison.
+These potential models are meant to be representative rather than definitive since the potential of the inner and outer Milky Way are not well known.
 
@@
 \begin{align}
-	\Phi(r) &= -v_h^2\,\frac{\ln{(1 + r/r_s)}}{r/r_s}\\
+	\Phi(r) &= -v_h^2\,\frac{\ln{(1 + r/r_s)}}{r/r_s}\label{eq:flatnfw}\\
 	r^2 &= x^2 + y^2 + \frac{z^2}{q_z^2}
@@
 For the bar component, we use a basis function expansion (BFE) of the potential and density of the bar with expansion coefficients derived for a triaxial, exponential bar density \citep[][hereafter W12]{wang12}.
-We use the pre-computed expansion coefficients used in W12, which were computed from a low-order expansion of the triaxial bar density used in \citet{dwek95}.\footnote{The coefficients presented in W12 are for just the cosine terms (the $A_{lm}$ in \citet{hernquist92} or the $S_{nlm}$ in \citet{lowing11}) because all sine terms have zero coefficients for a triaxial density function.} We have implemented the BFE computation of the potential, density, and gradient of the potential in C and Python and the code is publicly available on \project{GitHub}.\footnote{\url{https://github.com/adrn/biff}}
+We use the pre-computed expansion coefficients used in W12, which were computed from a low-order expansion of the triaxial bar density used in \citet{dwek95}.\footnote{The coefficients presented in W12 are for just the cosine terms (the $A_{lm}$ in \citet{hernquist92} or the $S_{nlm}$ in \citet{lowing11}) because all sine terms have zero coefficients for a triaxial density function.} We have implemented the BFE computation of the potential, density, and gradient of the potential in \texttt{C} and \python\ and the code is publicly available on \github.\footnote{\url{https://github.com/adrn/biff}}
 
-The BFE representation fixes the axis ratios of the bar---that is, the the exponential scale lengths along the three axes of the bar were adopted from \cite{dwek95} when the expansion coefficients were calculated in W12)---but the pattern speed and angle with respect to the Galactic $x$ axis are left as free parameters; all other potential parameter values are given in Table~\ref{tbl:potential-params-barred}.
+The BFE representation fixes the axis ratios of the bar---that is, the the exponential scale lengths along the three axes of the bar were adopted from \cite{dwek95} when the expansion coefficients were calculated in W12); all other potential parameter values are given in Table~\ref{tbl:potential-params-barred}.
 The mass of the halo is fixed and the mass of the disk and bar are varied in order to qualitatively reproduce the flatness and amplitude of the circular velocity curve of the Milky Way \citep{bovy12}.
 Figure~\ref{fig:circ-vel-barred} shows the circular velocity along the line connecting the Sun to the Galactic center in this model (the Galactic $x$ axis).
-Figure~\ref{fig:surface-density-barred} shows contours of constant surface density for a face-on (left) and edge-on (right) view of this potential model \citep[compare to, e.g., Figure 3 in][]{portail15}.
+Figure~\ref{fig:surface-density-barred} shows contours of constant surface density for a face-on (left) and edge-on (right) view of this potential model with the bar angle set to $20^\circ$ \citep[compare to, e.g., Figure 3 in][]{portail15}.
 
@@
 		& $b$ & 0.28~{\rm kpc} \\\midrule
-	         Halo & $v_h$ & 185.8~\kms\\
+	         Halo & $v_c$ & 185.8~\kms\\
 		& $r_s$ & 30~kpc \\
 		& $q_z$ & 0.9 \\\midrule
-%		Spheroid & $M_{\rm sph}$ & \\
-%		& $c$ &  \\\midrule
 		Bar & $M_{\rm bar}$ & $1.8 \times 10^{10}~\msun$ \\
+		& $\alpha$ & $20^\circ$\\
+		& $\Omega$ & $60~{\rm km}~{\rm s}^{-1}~{\rm kpc}^{-1}$ \\
 		\bottomrule
@@
 The bar mass is taken from recent 3D density modeling of red clump stars in the Galactic bulge \citep{portail15}.
+The bar angle, $\alpha$, and pattern speed, $\Omega$, are taken taken from \cite{dwek95} and \cite{wang12}, respectively.
 Halo flattening is taken from recent work that suggests the inner Milky Way halo is slightly oblate \citep{koposov10, kuepper15}.
+The halo mass scale is set by specifying the circular velocity at the scale radius, $v_c$, and the scale velocity in Equation~\ref{eq:flatnfw} is given by $v_h^2 = v_c^2 / (\ln2 - 1/2)$.
 \label{tbl:potential-params-barred}}
@@
 		& $b$ & 0.28~{\rm kpc} \\\midrule
-	         Halo & $v_h$ & 185.8~\kms\\
+	         Halo & $v_c$ & 185.8~\kms\\
 		& $r_s$ & 30~kpc \\
@@
 
+We wish to classify and analyze orbits around the orbit of the progenitor of the Ophiuchus stream, however the progenitor has not been found.
+We therefore take the medan position and velocity of stream members  from previous work and sample from the error distribution to generate initial conditions (these values are given in Table 3 of S15).
+We assume the mean Galactic longitude and latitude are fixed but generate samples in distance modulus, proper motions, and line-of-sight velocity assuming all uncertainties are Gaussian and independent.
+For each sample, we compute the proper motion angle (on the sky, in Galactic coordinates) and reject any sample with an angle relative to the elongation of the stream $>15^\circ$.
+We convert all sampled heliocentric coordinates to Galactocentric, cartesian coordinates by fixing the Sun to $(x,y,z) = (-8,0,0)~{\rm kpc}$ with $(v_x,v_y,v_z) = (-11.1, 220+24, 7.25)~{\rm km}~{\rm s}^{-1}$ \citep{schonrich10, bovy12}.
+
 \subsection{Detecting and characterizing chaos with frequency diffusion}\label{sec:chaos-indicator}
 
-A regular orbit can be characterized by a set of fundamental frequencies.
+Regular orbits are quasi-periodic: any phase-space component of the orbit can be decomposed in terms of a Fourier series, e.g.,
+\begin{equation}
+	x(t) = \sum^\infty_k \, a_{x,k}\, e^{i\, \omega_k \, t}.
+\end{equation}
+The amplitudes, $\bs{a}_k$ are constants which determine the strength of each mode, and the frequencies in each Fourier mode, $\omega_k$, are linear, integer combinations of the fundamental frequencies of the orbit, $\bs{\Omega}$---that is,
+\begin{equation}
+	\omega_k = \bs{n}_k \cdot \bs{\Omega}
+\end{equation}
+where $\bs{n}_k$ is a integer vector with dimensionality equal to the number of degrees of freedom of the gravitational potential \citep[e.g.,][]{binney82, binneytremaine}.
+The fundamental frequencies of a regular orbit are directly related to the orbital actions, a set of integrals of motion intricately related to the [...] of Hamiltonian and nonlinear dynamics \citep[e.g.,][]{??}.
+The fundamental frequencies can be approximated for numerically-integrated orbital time series by measuring the positions and amplitudes of spectral lines in the Fourier transform of the orbital components.
+A very accurate procedure for numerically computing the frequencies was developed by \citet{laskar93} and has been re-implemented in \python\ and released as open-source software \citep[\superfreq;][]{superfreq}.
+
+Chaotic orbits have no actions and therefore no fundamental frequencies.
+The frequencies and actions drift erratically in time with rates that depend on the structure of resonances surrounding such orbits.
+The evolution of the frequencies changes the configuration-space appearance of finite sub-sections of chaotic orbits and greatly enhances the density evolution of ensembles of nearby chaotic orbits; this has a dramatic effect on the morphological evolution of tidal streams \citep[e.g.,][]{apw15-chaos}.
+Numerically-integrated trajectories can be identified as chaotic orbits by the detection of frequency diffusion \citep[e.g.,][]{laskar96, valluri98}.
+
+The accuracy of the frequency-finding method (\superfreq) allows for detecting very small changes in the frequencies of even mildly chaotic orbits.
+Measuring the rate of frequency diffusion is therefore a powerful tool for detecting and characterizing dynamical chaos \citep[e.g.,][]{many}.
+From measuring the frequencies in many consecutive, overlapping time-windows along a numerically integrated orbit, we define the frequency diffusion rate, $DERP$, as
+\begin{equation}
+	yo
+\end{equation}
+where the index $k$ refers to the window number and $\tau$ is the shift in time between two consecutive windows.
+We note that this definition differs from previous, coarse approximations of the diffusion rate that simply differenced the frequencies computed in two consecutive time windows.
+We use this frequency diffusion rate to identify chaotic orbits integrated in the potential models described above.
 
@@
 
-by releasing ensembles of test particles when the progenitor orbit crosses pericenter.
+To generate mock stellar streams, we use a method similar to \streakline\ \citep{kuepper12}.
+For a given set of ``final conditions''---here, a position and velocity sampled from the observational error distribution for the Ophiuchus stream; Section~\ref{sec:orbit-grids}---we integrate the orbit backwards in time for a given integration period, $T$, in the specified potential.
+To generate a mock stream, we must set a progenitor mass, $m$, \todo{and mass-loss rate, $\dot{m}$} which sets the tidal radius of the mock disrupting globular cluster.
+From the endpoint of the backwards-integration (e.g., the initial conditions), we begin integrating the orbit forward in time, but now at each time-step a test-particle is released at each Lagrange point of the parent, progenitor orbit.
+The position of the test-particle is sampled from an isotropic Gaussian centered on each of the Lagrange points and the velocity of the particle is set equal to the velocity of the progenitor with a small Gaussian scatter in the radial direction.
+\todo{why these choices?} This method has been shown to reproduce the morphologies of N-body simulations of globular cluster streams, but requires far less computing time because it relies only on integrating test-particle orbits.
 
@@
 
-\subsection{The degree of chaos}\label{sec:results1}
+Using the orbit initial conditions described above (Section~\ref{sec:orbit-grids}), we integrate orbits and compute the frequency diffusion rate for the static potential and for XX barred models with varied initial bar angle and pattern speed.
+We find that [...]
 
@@
 
-% Recent work has mapped the 3D density distribution of the region around the bulge using number counts of red clump stars and have measured the bar angle and axis ratios of the bar structure \citep{wegg13}.
-The authors do not fit an analytic density model to the data and do not provide enough information about the density field to reconstruct the full-space stellar density distribution of the bulge.
-Further, this is not a dynamical model for the bar and therefore may not account for all of the mass but rather provides information about the number density of stars in the Galactic bar.
-\cite{portail15}
+We have found that a qualitative but observationally-motivated potential model for the Galactic bar generates a significant number of chaotic orbits in the vicinity of the recently-discovered Ophiuchus stream.
+By sampling many possible phase-space 6D positions and velocities for the stream centroid from the observational error distribution, we have shown that most possible orbits are strongly chaotic in the time-dependent, barred potential model for the Milky Way, but are mostly regular, eccentric orbits in a static, axisymmetric potential.
+We have shown that---using \streakline\ models of the stream density---the apparent shortness of the stream and the existence of nearby, high-dispersion stars can be explained by chaotic density evolution of the stripped stars.
+This is the first time chaos has been used to explain the morphology and [...] of a stellar stream.
+This work demonstrates the importance of the Galactic bar on the potential of the inner Milky Way and has important implications for future modeling of streams near in this region.
+This result also motivates deep imaging around the Ophiuchus stream to measure the density contrast of the ``fanned'' debris compared to the main stream and to find more candidate stars associated with chaotic tidal debris.
 

### 2015-11-16 ms.tex
words

@@
 
+\newcommand{\DM}{{\rm DM}}
+
 % TO DO
@@
 \author{Adrian M. Price-Whelan\altaffilmark{\colum,\adrn},
-            Branimir Sesar\altaffilmark{\mpia}, ...
-%            Sarah Pearson\altaffilmark{\colum},
-%            Kathryn V. Johnston\altaffilmark{\colum},
-%            Hans-Walter Rix\altaffilmark{\mpia}
+            Branimir Sesar\altaffilmark{\mpia}, 
+            Kathryn V. Johnston\altaffilmark{\colum},
+            Hans-Walter Rix\altaffilmark{\mpia}
 }
@@
 
-The deprojected length of the visible part of the stream ($\approx$1.5 kpc) is short for the age of its stellar population.
+The deprojected length of the visible part of the stream ($\approx$1.5 kpc) is short given the age of its stellar population.
 S15a fit an orbit to the kinematics of the stream stars in a static, axisymmetric model for the gravitational field of the inner Galaxy and ran N-body simulations of globular clusters on this orbit.
@@
 This dynamical age is at odds with the old stellar population.
-More recently, a significant observational effort has followed-up nearly all BHB star candidates within XX deg of the stream to obtain radial velocities to identify more stream members \citep{sesar15b}.
+More recently, a significant observational effort has followed-up nearly all blue horizontal branch (BHB) star candidates within XX deg of the stream to obtain radial velocities to identify more stream members \citep{sesar15b}.
 The stream has a very distinct and large line-of-sight velocity ($\approx$290 \kms) and is therefore easily detected above the background halo population.
@@
 
-Our goal is to (1) assess whether the Galactic bar can produce significant numbers of chaotic orbits in the inner Galaxy in the vicinity of the Ophiuchus stream and (2) if chaotic density evolution of tidal debris stripped from the progenitor of this stream can explain the apparent shortness of the stream and low-density, high-dispersion ends.
+Our goal is to (1) assess whether the Galactic bar can produce chaotic orbits in the inner Galaxy in the vicinity of the Ophiuchus stream and (2) if chaotic density evolution of tidal debris stripped from the progenitor of this stream can explain the apparent shortness of the stream and low-density, high-dispersion ends.
 In this section, we describe the potential models we use to represent the galaxy and describe the methods we use to detect and quantify the strength of chaos for orbits.
-We then describe a simplified version of the \streakline\ method that we use to generate mock stellar streams \citep{kuepper12}.
+We then describe the method we use for fitting orbits to the stream stars.
+Finally, we describe the method we use to generate mock stellar streams.
 
@@
 
-\subsection{Orbit grid initial conditions}\label{sec:orbit-grids}
+\subsection{Fitting orbits to the Ophiuchus stream}\label{sec:orbit-fit}
 
-We wish to classify and analyze orbits around the orbit of the progenitor of the Ophiuchus stream, however the progenitor has not been found.
+In each of the potentials described above, we fit orbits to observed kinematic properties of BHB stars that are confirmed members of the Ophiuchus stream \citep{sesar2015a, sesar2015b}.
+We use a methodology similar to that used in \cite{koposov10} and \citep{sesar2015a}.
+Our goal is to infer the posterior probability distributions over the orbital initial conditions given a potential, $\Phi$, and kinematic data for stream stars---$\bs{x}_i=(l, b, \DM, \mu_l, \mu_b, v_r)_i$ for the $i$-th star, where ($l$, $b$) are Galactic coordinates, $\DM$ is the distance modulus, ($\mu_l$, $\mu_b$) are proper motions in the Galactic frame, and $v_r$ is the radial velocity.
+We assume that the sky coordinates for each star are known perfectly well (have zero uncertainty) and transform the data to a rotated, heliocentric coordinate system that is aligned with the extent of the stream and centered on the median sky position of the BHB stars of \cite{sesar2015a}.
+We represent the longitude and latitude in these coordinates as $(\phi_1, \phi_2)$ and the rotation matrix to transform from Galactic to these coordinates is given in Appendix~\ref{sec:rotationmatrix}.
+We treat the stream longitude, $\phi_1$, as the perfectly-known, independent variable so that all other coordinates can be expressed as functions of this longitude (e.g., $\phi_2(\phi_1)$, ${\rm DM}(\phi_1)$, etc.).
+
+We include three nuisance parameters in our likelihood to account for the internal dispersion of the stream: $\bs{s} = (s_{\phi_2}, s_{\DM}, s_{v_r})$ (the proper motion uncertainties are sufficiently large that we can't resolve the velocity dispersion in these coordinates).\footnote{We assume that the dispersion in these coordinates is constant over the observed (short) section of the stream.
+This may be a bad assumption.} We add two additional nuisance parameters for controlling the amount of time to integrate forwards, $t_f$, and backwards, $t_b$, from the given initial conditions, which ultimately controls the length of the section of orbit that is compared to the stream star data.
+
+We assume that each observed kinematic component is independent so that the likelihood of the data for a given star, $\bs{x}_i$, with uncertainties $\bs{\sigma}_i$ is given by the product over the likelihoods for each dimension of the data.
+Defining $\bs{\theta} = (\Phi, t_b, t_f)$,
+\begin{multline}
+	p(\bs{x}_i \given \bs{\sigma}_i, \bs{s}, \bs{\theta}) = p(\phi_{2,i} \given \phi_{1,i}, s_{\phi_2},\bs{\theta}) \, p(\DM_i \given \phi_{1,i}, \sigma_{\DM,i}, s_\DM, \bs{\theta})\\ 
+	\times p(\mu_{l,i} \given \phi_{1,i}, \sigma_{\mu_{l},i}, \bs{\theta}) \, p(\mu_{b,i} \given \phi_{1,i}, \sigma_{\mu_{b},i}, \bs{\theta}) \, p(v_{r,i} \given \phi_{1,i}, \sigma_{v_r,i}, s_{v_r}, \bs{\theta})
+\end{multline}
+
+, however the progenitor has not been found.
 We therefore take the medan position and velocity of stream members  from previous work and sample from the error distribution to generate initial conditions (these values are given in Table 3 of S15).
@@
 We convert all sampled heliocentric coordinates to Galactocentric, cartesian coordinates by fixing the Sun to $(x,y,z) = (-8,0,0)~{\rm kpc}$ with $(v_x,v_y,v_z) = (-11.1, 220+24, 7.25)~{\rm km}~{\rm s}^{-1}$ \citep{schonrich10, bovy12}.
+
+In what follows, we assume the Galactocentric position and velocity of the Sun are: 
+\begin{align}
+	(x,y,z)_\odot &= (-8,0,0)~{\rm kpc}\\
+	(v_x,v_y,v_z)_\odot &= (-11.1, 220+24, 7.25)~{\rm km}~{\rm s}^{-1}
+\end{align}
+\citep{8kpref??, schonrich10, bovy12}.
 
@@
 
-To generate mock stellar streams, we use a method similar to \streakline\ \citep{kuepper12}.
-For a given set of ``final conditions''---here, a position and velocity sampled from the observational error distribution for the Ophiuchus stream; Section~\ref{sec:orbit-grids}---we integrate the orbit backwards in time for a given integration period, $T$, in the specified potential.
-To generate a mock stream, we must set a progenitor mass, $m$, \todo{and mass-loss rate, $\dot{m}$} which sets the tidal radius of the mock disrupting globular cluster.
-From the endpoint of the backwards-integration (e.g., the initial conditions), we begin integrating the orbit forward in time, but now at each time-step a test-particle is released at each Lagrange point of the parent, progenitor orbit.
+To generate mock stellar streams, we use the method presented in \citep{fardal15} (similar to the \streakline\ method presented in \citet{kuepper12}, but here the stars are released with some dispersion in position and velocity).
+For a given set of ``final conditions''---here, a position and velocity sampled from the observational error distribution from the orbit fit to the Ophiuchus stream (Section~\ref{sec:orbit-grids})---we integrate the orbit backwards in time for a given integration period, $T$, in each potential model.
+From the endpoint of the backwards-integration (e.g., the initial conditions), we begin integrating the orbit forward in time, but now at each time-step a star particle is released near each Lagrange point of the progenitor.
+The position of the Lagrange points and the scale of the dispersion in position and velocity are set by the progenitor mass, $m$.
 The position of the test-particle is sampled from an isotropic Gaussian centered on each of the Lagrange points and the velocity of the particle is set equal to the velocity of the progenitor with a small Gaussian scatter in the radial direction.
@@
 
+What is a successful model?
+- Sharp density drop
+	- BHB density falls off sooner? Is that just small-number statistics?
+	- Density profile along the stream / how sharp is the drop?
+	- Density contrast for BHB stars?
+- Increased positional dispersion
+	- The 4 new stars
+- Velocity dispersion of the 4 new stars
+	- velocity dispersion ~ 40 times internal velocity dispersion of stream stars
+- What can we say about the connection between chaotic timescale and the length of the stream? Is it the onset time for chaos, or are all orbits strongly chaotic?
+	- for a given orbit and model, paint the particles to say what the mass loss rate would have to be to match the obs.
+- "It's stupid to make it more complicated" -- Kathryn Johnston
+- from painting particles, Length of stream --> dispersion
+
 \acknowledgements
@@
 \bibliography{refs}
+
+\appendix
+\section{Stream coordinates} \label{sec:rotationmatrix}
+\todo{Add this!}
 

### 2015-11-17 ms.tex
writing about orbit fitting

@@
 \newcommand{\DM}{{\rm DM}}
+\newcommand{\norm}{\mathcal{N}}
 
@@
 
-In each of the potentials described above, we fit orbits to observed kinematic properties of BHB stars that are confirmed members of the Ophiuchus stream \citep{sesar2015a, sesar2015b}.
-We use a methodology similar to that used in \cite{koposov10} and \citep{sesar2015a}.
-Our goal is to infer the posterior probability distributions over the orbital initial conditions given a potential, $\Phi$, and kinematic data for stream stars---$\bs{x}_i=(l, b, \DM, \mu_l, \mu_b, v_r)_i$ for the $i$-th star, where ($l$, $b$) are Galactic coordinates, $\DM$ is the distance modulus, ($\mu_l$, $\mu_b$) are proper motions in the Galactic frame, and $v_r$ is the radial velocity.
-We assume that the sky coordinates for each star are known perfectly well (have zero uncertainty) and transform the data to a rotated, heliocentric coordinate system that is aligned with the extent of the stream and centered on the median sky position of the BHB stars of \cite{sesar2015a}.
+In each of the potentials described above, we fit orbits to observed kinematic properties of BHB stars that are confirmed members of the Ophiuchus stream \citep{sesar15a, sesar15b}.
+Our goal is to infer the posterior probability distributions over orbital initial conditions, $\bs{w}_0=(l, b, \DM, \mu_l, \mu_b, v_r)_0$, given a potential, $\Phi$, and kinematic data for each $i$ stream star, $\bs{x}_i=(l, b, \DM, \mu_l, \mu_b, v_r)_i$.
+In this notation, $(l, b)$ are Galactic coordinates, $\DM$ is the distance modulus, $(\mu_l, \mu_b)$ are proper motions in the Galactic frame, and $v_r$ is the radial velocity.
+We assume that the sky coordinates for each star are known perfectly well (have zero uncertainty) and transform the data to a rotated, heliocentric coordinate system that is aligned with the stream and centered on the median sky position of the BHB stars in the densest part of the stream \cite[all BHB stars except the `fanned' stars: cand15, cand26, cand49, cand54 from][]{sesar15b}.
 We represent the longitude and latitude in these coordinates as $(\phi_1, \phi_2)$ and the rotation matrix to transform from Galactic to these coordinates is given in Appendix~\ref{sec:rotationmatrix}.
 We treat the stream longitude, $\phi_1$, as the perfectly-known, independent variable so that all other coordinates can be expressed as functions of this longitude (e.g., $\phi_2(\phi_1)$, ${\rm DM}(\phi_1)$, etc.).
+This methodology is similar to that used in \cite{koposov10} and \cite{sesar15a}.
 
-We include three nuisance parameters in our likelihood to account for the internal dispersion of the stream: $\bs{s} = (s_{\phi_2}, s_{\DM}, s_{v_r})$ (the proper motion uncertainties are sufficiently large that we can't resolve the velocity dispersion in these coordinates).\footnote{We assume that the dispersion in these coordinates is constant over the observed (short) section of the stream.
+\subsubsection{Likelihood}
+
+We include three nuisance parameters in our likelihood to account for the internal dispersion of the stream: in observed coordinates, these are the on-sky positional dispersion, $s_{\phi_2}$, a distance (modulus) dispersion, $s_{\DM}$, and a radial velocity dispersion, $s_{v_r}$ (the proper motion uncertainties are sufficiently large that we can't resolve the velocity dispersion in these coordinates).\footnote{We assume that the dispersion in these coordinates is constant over the observed (short) section of the stream.
 This may be a bad assumption.} We add two additional nuisance parameters for controlling the amount of time to integrate forwards, $t_f$, and backwards, $t_b$, from the given initial conditions, which ultimately controls the length of the section of orbit that is compared to the stream star data.
+For brevity in the equations below, we define $\bs{s} = (s_{\phi_2}, s_{\DM}, s_{v_r})$ and $\bs{\theta} = (\bs{w}_0, \Phi, t_b, t_f)$.
 
-We assume that each observed kinematic component is independent so that the likelihood of the data for a given star, $\bs{x}_i$, with uncertainties $\bs{\sigma}_i$ is given by the product over the likelihoods for each dimension of the data.
-Defining $\bs{\theta} = (\Phi, t_b, t_f)$,
-\begin{multline}
-	p(\bs{x}_i \given \bs{\sigma}_i, \bs{s}, \bs{\theta}) = p(\phi_{2,i} \given \phi_{1,i}, s_{\phi_2},\bs{\theta}) \, p(\DM_i \given \phi_{1,i}, \sigma_{\DM,i}, s_\DM, \bs{\theta})\\ 
-	\times p(\mu_{l,i} \given \phi_{1,i}, \sigma_{\mu_{l},i}, \bs{\theta}) \, p(\mu_{b,i} \given \phi_{1,i}, \sigma_{\mu_{b},i}, \bs{\theta}) \, p(v_{r,i} \given \phi_{1,i}, \sigma_{v_r,i}, s_{v_r}, \bs{\theta})
-\end{multline}
-
-, however the progenitor has not been found.
-We therefore take the medan position and velocity of stream members  from previous work and sample from the error distribution to generate initial conditions (these values are given in Table 3 of S15).
-We assume the mean Galactic longitude and latitude are fixed but generate samples in distance modulus, proper motions, and line-of-sight velocity assuming all uncertainties are Gaussian and independent.
-For each sample, we compute the proper motion angle (on the sky, in Galactic coordinates) and reject any sample with an angle relative to the elongation of the stream $>15^\circ$.
-We convert all sampled heliocentric coordinates to Galactocentric, cartesian coordinates by fixing the Sun to $(x,y,z) = (-8,0,0)~{\rm kpc}$ with $(v_x,v_y,v_z) = (-11.1, 220+24, 7.25)~{\rm km}~{\rm s}^{-1}$ \citep{schonrich10, bovy12}.
-
-In what follows, we assume the Galactocentric position and velocity of the Sun are: 
+For a given set of initial conditions ($\bs{w}_0$), we compute a model orbit as follows: (1) transform the initial conditions to Galactocentric coordinates, where we take the position and velocity of the Sun to be
 \begin{align}
@@
 \end{align}
-\citep{8kpref??, schonrich10, bovy12}.
+\citep{8kpref??, schonrich10, bovy12}, (2) integrate the orbit forward and backward by $t_f$ and $t_b$, respectively, in the potential $\Phi$, (3) transform all orbit points (time-steps) back to observed coordinates, and (4) define interpolating functions for each coordinate as a function of stream longitude, $\phi_1$, using cubic splines---e.g., functions $\widetilde{\phi}_{2}(\phi_1)$, $\widetilde{\DM}(\phi_1)$, $\widetilde{\mu_l}(\phi_1)$, $\widetilde{\mu_b}(\phi_1)$, $\widetilde{v_r}(\phi_1)$.
+These functions let us compute the predicted values of each of these coordinates at the longitudes of each observed star, $\phi_{1,i}$.
+
+We assume that each observed kinematic component is independent so that the likelihood of the data for a given star, $\bs{x}_i$, with uncertainties, $\bs{\sigma}_i$, is given by the product over the likelihoods for each dimension of the data:
+\begin{multline}
+	p(\bs{x}_i \given \bs{\sigma}_i, \bs{s}, \bs{\theta}) = p(\phi_{2,i} \given \phi_{1,i}, s_{\phi_2},\bs{\theta}) \, p(\DM_i \given \phi_{1,i}, \sigma_{\DM,i}, s_\DM, \bs{\theta})\\ 
+	\times p(\mu_{l,i} \given \phi_{1,i}, \sigma_{\mu_{l},i}, \bs{\theta}) \, p(\mu_{b,i} \given \phi_{1,i}, \sigma_{\mu_{b},i}, \bs{\theta}) \, p(v_{r,i} \given \phi_{1,i}, \sigma_{v_r,i}, s_{v_r}, \bs{\theta}).
+\end{multline}
+The uncertainties in these observed coordinate components are assumed to be normally distributed away from the model values: using the notation
+\begin{align}
+	\norm(x \given \mu, \sigma^2) &= \frac{1}{\sqrt{2\pi \sigma^2}} \, \exp\left(-\frac{(x-\mu)^2}{2\sigma^2}\right)
+\end{align}
+the likelihoods are
+\begin{align}	
+	p(\phi_{2,i} \given \phi_{1,i}, s_{\phi_2},\bs{\theta}) &= \norm(\phi_{2,i} \given \widetilde{\phi}_{2}(\phi_{1,i}), s^2_{\phi_2})\\
+	p(\DM_i \given \phi_{1,i}, \sigma_{\DM,i}, s_\DM, \bs{\theta}) &= \norm(\DM_i \given \widetilde{\DM}(\phi_{1,i}), s^2_{\DM} + \sigma^2_{\DM,i})\\
+	p(\mu_{l,i} \given \phi_{1,i}, \sigma_{\mu_{l},i}, \bs{\theta}) &= \norm(\mu_{l,i} \given \widetilde{\mu_{l}}(\phi_{1,i}), \sigma^2_{\mu_{l,i}})\\
+	p(\mu_{b,i} \given \phi_{1,i}, \sigma_{\mu_{b},i}, \bs{\theta}) &= \norm(\mu_{b,i} \given \widetilde{\mu_{b}}(\phi_{1,i}), \sigma^2_{\mu_{b,i}})\\
+	p(v_{r,i} \given \phi_{1,i}, \sigma_{v_r,i}, s_{v_r}, \bs{\theta}) &= \norm(v_{r,i} \given \widetilde{v_r}(\phi_{1,i}), s^2_{v_r} + \sigma^2_{v_r,i}).
+\end{align}
+
+\subsubsection{Priors}
+
+For the intrinsic dispersion parameters, we use logarithmic (scale-invariant) priors such that $p(s) \propto s^{-1}$.
+For the integration time parameters,\footnote{We take the present-day to be $t=0$} we use uniform priors, $\mathcal{U}(a,b)$, over the range $a$--$b$:
+\begin{align}
+	p(t_f) &= \mathcal{U}(1,100)~{\rm Myr}\\
+	p(t_b) &= \mathcal{U}(-100,-1)~{\rm Myr}.
+\end{align}
+However, for computational efficiency, we place strong priors on the minimum and maximum longitudes of the model points, $(\phi_{1,{\rm min}},\phi_{1,{\rm max}})$ so that the model orbit does not integrate for longer than necessary.
+In particular, we set
+\begin{align}
+	p(\phi_{1,{\rm min}} \given \bs{\theta}) &= \norm(\phi_{1,{\rm min}} \given \min(\phi_{1,i}), s^2_{\phi_2})\\
+	p(\phi_{1,{\rm max}} \given \bs{\theta}) &= \norm(\phi_{1,{\rm max}} \given \max(\phi_{1,i}), s^2_{\phi_2}).
+\end{align}
+For the orbital initial conditions, we assume uniform priors in each cartesian position component over the range $(-200,200)~{\rm kpc}$.
+For the velocity, we assume a Gaussian prior in the magnitude of the total velocity, $v$, with a dispersion of $150~\kms$,
+\begin{equation}
+	\norm(v \given 0, (150~\kms)^2)
+\end{equation}
+We keep the potential, $\Phi$, fixed.
+
+\subsubsection{Initializing the MCMC}
+
+...
 

### 2015-11-18 ms.tex
methods section done

@@
 
-\subsection{Fitting orbits to the Ophiuchus stream}\label{sec:orbit-fit}
+\subsection{Fitting orbits to the Ophiuchus stream}\label{sec:orbitfit}
 
@@
 \end{align}
+We assume the data from each star is independent and identically distributed (i.i.d.) so that the full likelihood is the product over the likelihoods for each star:
+\begin{equation}
+	 p(\{\bs{x}_i\} \given \{\bs{\sigma}_i\}, \bs{s}, \bs{\theta}) = \prod_i p(\bs{x}_i \given \bs{\sigma}_i, \bs{s}, \bs{\theta}).\label{eq:likelihood}
+\end{equation}
 
@@
 For the intrinsic dispersion parameters, we use logarithmic (scale-invariant) priors such that $p(s) \propto s^{-1}$.
-For the integration time parameters,\footnote{We take the present-day to be $t=0$} we use uniform priors, $\mathcal{U}(a,b)$, over the range $a$--$b$:
+For the integration time parameters, we use uniform priors, $\mathcal{U}(a,b)$ (over the range $a$--$b$),
 \begin{align}
-	p(t_f) &= \mathcal{U}(1,100)~{\rm Myr}\\
+	p(t_f) &= \mathcal{U}(1,100)~{\rm Myr}\label{eq:prior1}\\
 	p(t_b) &= \mathcal{U}(-100,-1)~{\rm Myr}.
 \end{align}
-However, for computational efficiency, we place strong priors on the minimum and maximum longitudes of the model points, $(\phi_{1,{\rm min}},\phi_{1,{\rm max}})$ so that the model orbit does not integrate for longer than necessary.
+Note that present-day is $t=0$.
+For computational efficiency, we place strong priors on the minimum and maximum longitudes of the model points, $(\phi_{1,{\rm min}},\phi_{1,{\rm max}})$ so that the model orbit does not integrate for longer than necessary.
 In particular, we set
@@
 \end{align}
-For the orbital initial conditions, we assume uniform priors in each cartesian position component over the range $(-200,200)~{\rm kpc}$.
-For the velocity, we assume a Gaussian prior in the magnitude of the total velocity, $v$, with a dispersion of $150~\kms$,
+For the orbital initial condition components, we use uniform priors in each cartesian position component over the range $(-200,200)~{\rm kpc}$.
+For velocity, we use a Gaussian prior on the magnitude of the total velocity, $v$, with a dispersion of $150~\kms$,
 \begin{equation}
-	\norm(v \given 0, (150~\kms)^2)
+	\norm(v \given 0, (150~\kms)^2) \label{eq:prior2}
 \end{equation}
-We keep the potential, $\Phi$, fixed.
+We keep the potential, $\Phi$, fixed. In total, this model has 10 parameters (5 phase-space coordinates, 5 nuisance parameters).
 
-\subsubsection{Initializing the MCMC}
+\subsubsection{Sampling from the posterior distributions}
 
-...
+The full expression for the posterior probability, $p(\bs{s}, \bs{w}_0, t_b, t_f \given \{\bs{x}_i\}, \{\bs{\sigma}_i\}, \Phi)$, is the joint likelihood (Equation~\ref{eq:likelihood}) multiplied by all priors described above (Equations~\ref{eq:prior1}--\ref{eq:prior2}).
+We use an ensemble Markov Chain Monte Carlo (MCMC) algorithm \citep{goodman10} implemented in \python (\package{emcee}) to draw samples from this posterior distribution \citep{foremanmackey13}.
+The algorithm uses an ensemble of individual ``walkers'' to adapt to the geometry of the parameter-space being explored.
+In all cases, we use 80 walkers.
 
-\subsection{Detecting and characterizing chaos with frequency diffusion}\label{sec:chaos-indicator}
+To initialize these walkers, we first run an optimization routine to maximize the likelihood.
+In detail, we use the Powell algorithm implemented in \package{scipy} \citep{powell64, scipy} to minimize the negative, log-likelihood.
+To generate initial conditions for the walkers, we sample from Gaussian distributions centered on the maximum likelihood values.
+For the coordinates, we set the dispersions of these Gaussians to 1/1000 of the median uncertainties of the stars.
+For the nuisance parameters, we set the dispersions to 1/1000 of their maximum likelihood values.
 
-Regular orbits are quasi-periodic: any phase-space component of the orbit can be decomposed in terms of a Fourier series, e.g.,
-\begin{equation}
-	x(t) = \sum^\infty_k \, a_{x,k}\, e^{i\, \omega_k \, t}.
-\end{equation}
-The amplitudes, $\bs{a}_k$ are constants which determine the strength of each mode, and the frequencies in each Fourier mode, $\omega_k$, are linear, integer combinations of the fundamental frequencies of the orbit, $\bs{\Omega}$---that is,
-\begin{equation}
-	\omega_k = \bs{n}_k \cdot \bs{\Omega}
-\end{equation}
-where $\bs{n}_k$ is a integer vector with dimensionality equal to the number of degrees of freedom of the gravitational potential \citep[e.g.,][]{binney82, binneytremaine}.
+For each potential, we run the MCMC walkers for a burn-in period of 128 steps and then re-initialize the walkers from their positions at the end of this run.
+This erases any relics of the initialization procedure outlined above.
+After burn-in, we run the walkers for an additional 256 steps.
+%
+%\subsection{Detecting and characterizing chaos with frequency diffusion}\label{sec:chaos-indicator}
+%
+%Regular orbits are quasi-periodic: any phase-space component of the orbit can be decomposed in terms of a Fourier series, e.g.,
+%\begin{equation}
+%	x(t) = \sum^\infty_k \, a_{x,k}\, e^{i\, \omega_k \, t}.
+%\end{equation}
+%The amplitudes, $\bs{a}_k$ are constants which determine the strength of each mode, and the frequencies in each Fourier mode, $\omega_k$, are linear, integer combinations of the fundamental frequencies of the orbit, $\bs{\Omega}$---that is,
+%\begin{equation}
+%	\omega_k = \bs{n}_k \cdot \bs{\Omega}
+%\end{equation}
+%where $\bs{n}_k$ is a integer vector with dimensionality equal to the number of degrees of freedom of the gravitational potential \citep[e.g.,][]{binney82, binneytremaine}.
 The fundamental frequencies of a regular orbit are directly related to the orbital actions, a set of integrals of motion intricately related to the [...] of Hamiltonian and nonlinear dynamics \citep[e.g.,][]{??}.
@@
 A very accurate procedure for numerically computing the frequencies was developed by \citet{laskar93} and has been re-implemented in \python\ and released as open-source software \citep[\superfreq;][]{superfreq}.
-
-Chaotic orbits have no actions and therefore no fundamental frequencies.
+%
+%Chaotic orbits have no actions and therefore no fundamental frequencies.
 The frequencies and actions drift erratically in time with rates that depend on the structure of resonances surrounding such orbits.
@@
 Numerically-integrated trajectories can be identified as chaotic orbits by the detection of frequency diffusion \citep[e.g.,][]{laskar96, valluri98}.
-
-The accuracy of the frequency-finding method (\superfreq) allows for detecting very small changes in the frequencies of even mildly chaotic orbits.
+%
+%The accuracy of the frequency-finding method (\superfreq) allows for detecting very small changes in the frequencies of even mildly chaotic orbits.
 Measuring the rate of frequency diffusion is therefore a powerful tool for detecting and characterizing dynamical chaos \citep[e.g.,][]{many}.
 From measuring the frequencies in many consecutive, overlapping time-windows along a numerically integrated orbit, we define the frequency diffusion rate, $DERP$, as
-\begin{equation}
-	yo
-\end{equation}
-where the index $k$ refers to the window number and $\tau$ is the shift in time between two consecutive windows.
+%\begin{equation}
+%	yo
+%\end{equation}
+%where the index $k$ refers to the window number and $\tau$ is the shift in time between two consecutive windows.
 We note that this definition differs from previous, coarse approximations of the diffusion rate that simply differenced the frequencies computed in two consecutive time windows.
@@
 
-To generate mock stellar streams, we use the method presented in \citep{fardal15} (similar to the \streakline\ method presented in \citet{kuepper12}, but here the stars are released with some dispersion in position and velocity).
-For a given set of ``final conditions''---here, a position and velocity sampled from the observational error distribution from the orbit fit to the Ophiuchus stream (Section~\ref{sec:orbit-grids})---we integrate the orbit backwards in time for a given integration period, $T$, in each potential model.
-From the endpoint of the backwards-integration (e.g., the initial conditions), we begin integrating the orbit forward in time, but now at each time-step a star particle is released near each Lagrange point of the progenitor.
+To generate mock stellar streams, we use the method presented in \citep{fardal14}: star particles are `released' from a progenitor system near the Lagrange points with some dispersion in position and velocity set by the mass and orbit of the progenitor (this is similar to the \streakline\ method of \citet{kuepper12} but with finite dispersions).
+To generate progenitor orbits, we draw samples from the posterior probability distributions over orbital parameters from fitting orbits to the stream star members (Section~\ref{sec:orbit-fit}).
+For a given progenitor orbit---the 6D position of the orbit today---we integrate the orbit backwards in time for a given integration period.
+From the endpoint of the backwards-integration (e.g., the past position), we begin integrating the orbit forward in time, but now at each time-step a star particle is released near each of the Lagrange points of the progenitor.
 The position of the Lagrange points and the scale of the dispersion in position and velocity are set by the progenitor mass, $m$.
-The position of the test-particle is sampled from an isotropic Gaussian centered on each of the Lagrange points and the velocity of the particle is set equal to the velocity of the progenitor with a small Gaussian scatter in the radial direction.
-\todo{why these choices?} This method has been shown to reproduce the morphologies of N-body simulations of globular cluster streams, but requires far less computing time because it relies only on integrating test-particle orbits.
+The star particles are drawn from Gaussians centered on the Lagrange points (in position) and the progenitor (in velocity) and the full parametrization of the release distribution is given in \cite{fardal14}.
+This method has been shown to reproduce the morphologies of $N$-body simulations of stellar streams, but requires far less computing time because it relies only on integrating test-particle orbits.
 
-\section{Results}
+\section{Results I: Orbit fits?}
 
@@
 
-\subsection{Reproducing the observed properties of the stream}\label{sec:results2}
+\section{Results 2: Models for the Ophiuchus stream}
 

### 2015-11-18 ms.tex
more refs yo

@@
 
-The Ophiuchus stream \citep{bernard14, sesar15} is a recently discovered stellar tidal stream that sits above the Galactic bulge at a Galactocentric radius and height $(R,z) \approx (1, 4.3)~{\rm kpc}$.
-All observational evidence suggests that the stream is a totally disrupted globular cluster: The stream stars have (1) a small positional dispersion orthogonal to the extended direction of the stream ($\approx$10 pc); (2) a small velocity dispersion $\approx$0.4 ${\rm km}~{\rm s}^{-1}$; (3) an old stellar population ($\approx$12 Gyr) estimated from isochrone fitting; and (4) no detectable over-density along the stream that could be  the progenitor system \citep[][hereafter S15a]{sesar15}.
+The Ophiuchus stream \citep{bernard14, sesar15a} is a recently discovered stellar tidal stream that sits above the Galactic bulge at a Galactocentric radius and height $(R,z) \approx (1, 4.3)~{\rm kpc}$.
+All observational evidence suggests that the stream is a totally disrupted globular cluster: The stream stars have (1) a small positional dispersion orthogonal to the extended direction of the stream ($\approx$10 pc); (2) a small velocity dispersion $\approx$0.4 ${\rm km}~{\rm s}^{-1}$; (3) an old stellar population ($\approx$12 Gyr) estimated from isochrone fitting; and (4) no detectable over-density along the stream that could be  the progenitor system \citep[][hereafter S15a]{sesar15a}.
 
@@
 Given the proximity of the stream to the center of the Galaxy, the time-dependent, triaxial potential of the Galactic bar must be taken into account when modeling the orbit of the progenitor system.
-The presence of a bar-like perturbation to the potential will change the possible orbit of the stream progenitor and will drastically change the resonance structure of orbits in the inner galaxy \citep{athanassoula, portail15b MANY}, potentially introducing a significant number of chaotic orbits in this region \citep{weinberg15}.
+The presence of a bar-like perturbation to the potential will change the possible orbit of the stream progenitor and will drastically change the resonance structure of orbits in the inner galaxy \citep{athanassoula, portail15b, zotos12}, potentially introducing a significant number of chaotic orbits in this region \citep{weinberg15}.
 

### 2015-11-23 ms.tex
more figures

@@
 %\title{Tapping the keg: the Galactic bar peculiarity of the Ophiuchus stream and the Galactic bar}
-\title{Spending too much time at the Galactic bar: did the Ophiuchus stream form on a chaotic orbit?}
+\title{Spending too much time at the Galactic bar: Chaotic evolution of the Ophiuchus stream}
 \author{Adrian M. Price-Whelan\altaffilmark{\colum,\adrn},
@@
 This dynamical age is at odds with the old stellar population.
-More recently, a significant observational effort has followed-up nearly all blue horizontal branch (BHB) star candidates within XX deg of the stream to obtain radial velocities to identify more stream members \citep{sesar15b}.
+More recently, a significant observational effort has followed-up nearly all blue horizontal branch (BHB) star candidates within XX deg of the stream to obtain radial velocities to identify more stream members \citep[][hereafter S15b]{sesar15b}.
 The stream has a very distinct and large line-of-sight velocity ($\approx$290 \kms) and is therefore easily detected above the background halo population.
@@
 In this work, we study whether stream-fanning---chaotic or simply from density evolution in a triaxial, time-dependent potential---can explain the observed properties of the Ophiuchus stream.
+In particular, we consider a model successful if it qualitatively reproduces:
+\begin{enumerate}
+	\item the apparent shortness and fast density truncation of the stream;
+	\item the increased positional dispersion of the four new candidate members from S15b;
+	\item the much larger velocity dispersion of the S15b stars.
+\end{enumerate}
 
@@
 
-Our goal is to (1) assess whether the Galactic bar can produce chaotic orbits in the inner Galaxy in the vicinity of the Ophiuchus stream and (2) if chaotic density evolution of tidal debris stripped from the progenitor of this stream can explain the apparent shortness of the stream and low-density, high-dispersion ends.
+Our goal is to (1) assess whether the Galactic bar can produce chaotic orbits in the vicinity of the Ophiuchus stream and (2) determine if chaotic density evolution of tidal debris stripped from the progenitor of this stream can explain the apparent shortness of the stream and low-density, high-dispersion ends.
 In this section, we describe the potential models we use to represent the galaxy and describe the methods we use to detect and quantify the strength of chaos for orbits.
@@
 Figure~\ref{fig:surface-density-barred} shows contours of constant surface density for a face-on (left) and edge-on (right) view of this potential model with the bar angle set to $20^\circ$ \citep[compare to, e.g., Figure 3 in][]{portail15}.
+We consider a grid of nine parameter combinations of bar angle and pattern speed.
+These model names and parameter values are given in Table~\ref{tbl:bar-specific}.
 
@@
 		Bar & $M_{\rm bar}$ & $1.8 \times 10^{10}~\msun$ \\
-		& $\alpha$ & $20^\circ$\\
-		& $\Omega$ & $60~{\rm km}~{\rm s}^{-1}~{\rm kpc}^{-1}$ \\
 		\bottomrule
@@
 	\caption{The disk potential scale lengths ($a$, $b$) were adopted following \citep{bovy15-galpy} to match the exponential scale length of the disk \citep{bovyrix13} and local dark-matter density \citep[e.g.,][]{bovytremaine12}.
-The bar mass is taken from recent 3D density modeling of red clump stars in the Galactic bulge \citep{portail15}.
-The bar angle, $\alpha$, and pattern speed, $\Omega$, are taken taken from \cite{dwek95} and \cite{wang12}, respectively.
 Halo flattening is taken from recent work that suggests the inner Milky Way halo is slightly oblate \citep{koposov10, kuepper15}.
 The halo mass scale is set by specifying the circular velocity at the scale radius, $v_c$, and the scale velocity in Equation~\ref{eq:flatnfw} is given by $v_h^2 = v_c^2 / (\ln2 - 1/2)$.
+The bar mass is taken from recent 3D density modeling of red clump stars in the Galactic bulge \citep{portail15}.
+The other bar parameters are listed in Table~\ref{tbl:bar-specific} next to the corresponding model name.
 \label{tbl:potential-params-barred}}
+\end{center}
+\end{table*}
+
+\begin{table*}[ht]
+\begin{center}
+	\begin{tabular}{ c | c | c }
+	         \toprule
+	         Name & $\alpha$ [deg] & $\Omega_p$ [${\rm km}~{\rm s}^{-1}~{\rm kpc}^{-1}$] \\\toprule
+		bar1 & 20 & 40\\
+		bar2 & 20 & 50\\
+		bar3 & 20 & 60\\
+		bar4 & 25 & 40\\
+		bar5 & 25 & 50\\
+		bar6 & 25 & 60\\
+		bar7 & 30 & 40\\
+		bar8 & 30 & 50\\
+		bar9 & 30 & 60\\
+		\bottomrule
+		\end{tabular}
+	\caption{Present-day bar angle ($\alpha$) and pattern speed ($\Omega_p$) for the nine parameter pairs considered in this work.
+These values span the range of recent measurements from a variety of techniques \citep{dwek95,wang12,wang13, MOREMOREMORE}.
+\label{tbl:bar-specific}}
 \end{center}
@@
 
-In each of the potentials described above, we fit orbits to observed kinematic properties of BHB stars that are confirmed members of the Ophiuchus stream \citep{sesar15a, sesar15b}.
+In each of the potentials described above, we fit orbits to the measured kinematics of BHB stars that are high-likelihood members of the Ophiuchus stream \citep{sesar15a, sesar15b}.
 Our goal is to infer the posterior probability distributions over orbital initial conditions, $\bs{w}_0=(l, b, \DM, \mu_l, \mu_b, v_r)_0$, given a potential, $\Phi$, and kinematic data for each $i$ stream star, $\bs{x}_i=(l, b, \DM, \mu_l, \mu_b, v_r)_i$.
@@
 The full expression for the posterior probability, $p(\bs{s}, \bs{w}_0, t_b, t_f \given \{\bs{x}_i\}, \{\bs{\sigma}_i\}, \Phi)$, is the joint likelihood (Equation~\ref{eq:likelihood}) multiplied by all priors described above (Equations~\ref{eq:prior1}--\ref{eq:prior2}).
-We use an ensemble Markov Chain Monte Carlo (MCMC) algorithm \citep{goodman10} implemented in \python (\package{emcee}) to draw samples from this posterior distribution \citep{foremanmackey13}.
+We use an ensemble Markov Chain Monte Carlo (MCMC) algorithm \citep{goodman10} implemented in \python\ (\package{emcee}) to draw samples from this posterior distribution \citep{foremanmackey13}.
 The algorithm uses an ensemble of individual ``walkers'' to adapt to the geometry of the parameter-space being explored.
@@
 To initialize these walkers, we first run an optimization routine to maximize the likelihood.
-In detail, we use the Powell algorithm implemented in \package{scipy} \citep{powell64, scipy} to minimize the negative, log-likelihood.
+In detail, we use the Powell algorithm implemented in \package{Scipy} \citep{powell64, scipy} to minimize the negative, log-likelihood.
 To generate initial conditions for the walkers, we sample from Gaussian distributions centered on the maximum likelihood values.
@@
 
-\section{Results I: Orbit fits?}
+\section{Results I: Orbit fits}\label{sec:results1}
 
-Using the orbit initial conditions described above (Section~\ref{sec:orbit-grids}), we integrate orbits and compute the frequency diffusion rate for the static potential and for XX barred models with varied initial bar angle and pattern speed.
+Figures~\ref{fig:orbitfits1}--\ref{fig:orbitfits2} summarize our results from fitting orbits to the BHB stream stars.
+Shown in each panel are the Ophiuchus stream star members (black points) and samples from the  posterior probability over orbital parameters (blue lines).
+The end-to-end integration time of the orbit over the observed extent of the stream is only $\approx$6 Myr, so the derived orbits are extremely similar in all potentials (the time-dependence of the bar potential is not significant over such short timescales).
+The orbital periods are typically $\approx$150--250 Myr.
+
+Though the orbital parameter values in observed coordinates are very similar between each potential model, the resulting orbits are quite different.
+Figure~\ref{fig:orbits-yz} shows projections of the ``mean'' orbits from each potential model; for each posterior distribution from orbit fitting, we take the mean values of the orbital parameters and convert to Galactocentric coordinates (e.g., Table~\ref{tbl:param-means}).
+For each orbit, we compute the maximum Lyapunov exponent (MLE, $\lambda$) to assess whether the orbit is chaotic.
+Though not predictive for the timescale over which chaos is important for tidal debris, the MLE is still an appropriate indicator of chaos \citep{apw15-chaos}.
+Figure~\ref{fig:lyapunov} shows numerical estimates of the Lyapunov time ($t_\lambda = 1/\lambda$) for the mean orbits in each potential model: in the static potential the orbit appears regular (black, diagonal line), but in all barred potentials the MLE estimates converge to finite values in the range $\approx$2--8 orbital periods.
+\todo{Should I compute Lyapunov exponents for samples from the posterior, not just the mean orbit?}
+
+The mean orbits in the barred potentials are all strongly chaotic.
+We have tried computing the frequency diffusion rate for these orbits but have found that, over two consecutive integration windows, the frequency recovery fails because the frequency spectrum changes dramatically.
+Over timescales of $\approx$10--20 orbital periods.
+
+% \section{Results ??: Characterizing chaos ..?}
+%Using the orbit initial conditions described above (Section~\ref{sec:orbit-grids}), we integrate orbits and compute the frequency diffusion rate for the static potential and for XX barred models with varied initial bar angle and pattern speed.
 We find that [...]
 
-\section{Results 2: Models for the Ophiuchus stream}
+% Full table with everything - boring
+%\begin{table*}[ht]
+%\footnotesize
+%\begin{center}
+%	\begin{tabular}{ccccccccc}
+%	\toprule
+%	name & $\phi_2$ [deg] & $d$ [kpc] & $\mu_l$ [mas yr$^{-1}$] & $\mu_b$ [mas yr$^{-1}$] & $v_r$ [km s$^{-1}$] & $s_{\phi_2}$ [deg] & $s_{d}$ [kpc] & $s_{v_r}$ [km s$^{-1}$]\\\midrule
+%static & $-0.03\pm0.05$ & $8.35\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.1\pm0.9$ & $0.20\pm0.04$ & $0.31\pm0.10$ & $2.9\pm0.8$\\
+%bar1 & $-0.03\pm0.05$ & $8.40\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.0\pm0.9$ & $0.21\pm0.05$ & $0.31\pm0.14$ & $3.2\pm0.8$\\
+%bar2 & $-0.03\pm0.05$ & $8.40\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $288.9\pm1.1$ & $0.20\pm0.04$ & $0.33\pm0.09$ & $3.3\pm0.9$\\
+%bar3 & $-0.04\pm0.06$ & $8.40\pm0.06$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.0\pm1.0$ & $0.21\pm0.05$ & $0.35\pm0.09$ & $3.2\pm0.8$\\
+%bar4 & $-0.03\pm0.06$ & $8.41\pm0.06$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.1\pm1.0$ & $0.21\pm0.05$ & $0.36\pm0.10$ & $3.4\pm0.9$\\
+%bar5 & $-0.03\pm0.05$ & $8.41\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.1\pm0.9$ & $0.21\pm0.04$ & $0.35\pm0.09$ & $3.3\pm0.8$\\
+%bar6 & $-0.03\pm0.05$ & $8.40\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.0\pm1.0$ & $0.21\pm0.05$ & $0.35\pm0.10$ & $3.2\pm0.8$\\
+%bar7 & $-0.04\pm0.05$ & $8.41\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.0\pm1.0$ & $0.21\pm0.05$ & $0.33\pm0.12$ & $3.4\pm0.9$\\
+%bar8 & $-0.03\pm0.05$ & $8.42\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.1\pm1.0$ & $0.20\pm0.04$ & $0.34\pm0.09$ & $3.4\pm0.9$\\
+%bar9 & $-0.02\pm0.05$ & $8.41\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $288.9\pm1.0$ & $0.21\pm0.04$ & $0.35\pm0.09$ & $3.3\pm0.8$\\
+%	\bottomrule
+%	\end{tabular}
+%	\caption{Estimated mean and standard deviation of the marginal posterior distributions over each parameter in our orbit fit likelihood.
+All values are nearly identical because the potentials are very similar over the short length of the stream.
+This table is mainly included as a sanity check---we have made samples from the full posterior distribution available with this article and provide code to transform to and from stream coordinates (see http://adrian.pw/ophiuchus/coordinates for more information).\label{tbl:param-means} }
+%\end{center}
+%\end{table*}
 
-\section{Conclusions}\label{sec:conclusions}
+\begin{table*}[ht]
+\footnotesize
+\begin{center}
+	\begin{tabular}{cccccc}
+	\toprule
+%	name & $\phi_2$ [deg] & $d$ [kpc] & $\mu_l$ [mas yr$^{-1}$] & $\mu_b$ [mas yr$^{-1}$] & $v_r$ [km s$^{-1}$] & $s_{\phi_2}$ [deg] & $s_{d}$ [kpc] & $s_{v_r}$ [km s$^{-1}$]\\\midrule
+%static & $-0.03\pm0.05$ & $8.35\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.1\pm0.9$ & $0.20\pm0.04$ & $0.31\pm0.10$ & $2.9\pm0.8$\\
+%bar1--9 & $-0.03\pm0.05$ & $8.40\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.0\pm1.0$ & $0.21\pm0.05$ & $0.31\pm0.14$ & $3.2\pm0.9$\\
+	name & $\phi_2$ [deg] & $d$ [kpc] & $\mu_l$ [mas yr$^{-1}$] & $\mu_b$ [mas yr$^{-1}$] & $v_r$ [km s$^{-1}$]\\\midrule
+	static & $-0.03\pm0.05$ & $8.35\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.1\pm0.9$\\
+	bar1--9 & $-0.03\pm0.05$ & $8.40\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.0\pm1.0$\\
+	\bottomrule
+	\end{tabular}
 
-We have found that a qualitative but observationally-motivated potential model for the Galactic bar generates a significant number of chaotic orbits in the vicinity of the recently-discovered Ophiuchus stream.
-By sampling many possible phase-space 6D positions and velocities for the stream centroid from the observational error distribution, we have shown that most possible orbits are strongly chaotic in the time-dependent, barred potential model for the Milky Way, but are mostly regular, eccentric orbits in a static, axisymmetric potential.
-We have shown that---using \streakline\ models of the stream density---the apparent shortness of the stream and the existence of nearby, high-dispersion stars can be explained by chaotic density evolution of the stripped stars.
-This is the first time chaos has been used to explain the morphology and [...] of a stellar stream.
-This work demonstrates the importance of the Galactic bar on the potential of the inner Milky Way and has important implications for future modeling of streams near in this region.
-This result also motivates deep imaging around the Ophiuchus stream to measure the density contrast of the ``fanned'' debris compared to the main stream and to find more candidate stars associated with chaotic tidal debris.
+	\begin{tabular}{ccc}
+	\toprule
+%	name & $\phi_2$ [deg] & $d$ [kpc] & $\mu_l$ [mas yr$^{-1}$] & $\mu_b$ [mas yr$^{-1}$] & $v_r$ [km s$^{-1}$] & $s_{\phi_2}$ [deg] & $s_{d}$ [kpc] & $s_{v_r}$ [km s$^{-1}$]\\\midrule
+%static & $-0.03\pm0.05$ & $8.35\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.1\pm0.9$ & $0.20\pm0.04$ & $0.31\pm0.10$ & $2.9\pm0.8$\\
+%bar1--9 & $-0.03\pm0.05$ & $8.40\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.0\pm1.0$ & $0.21\pm0.05$ & $0.31\pm0.14$ & $3.2\pm0.9$\\
+	$s_{\phi_2}$ [deg] & $s_{d}$ [kpc] & $s_{v_r}$ [km s$^{-1}$]\\\midrule
+	$0.20\pm0.04$ & $0.31\pm0.10$ & $2.9\pm0.8$\\
+	$0.21\pm0.05$ & $0.31\pm0.14$ & $3.2\pm0.9$\\
+	\bottomrule
+	\end{tabular}
+	\caption{Estimated mean and standard deviation of samples from the marginal posterior distributions over each parameter in our orbit fit model.
+For the barred potentials, all mean values are the same because the time-dependence of the bar doesn't impact the orbit fit over the short length of the stream.
+We have made samples from the full posterior distribution available with this article and provide code to transform to and from stream coordinates (see http://adrian.pw/ophiuchus/coordinates for more information).\label{tbl:param-means} }
+\end{center}
+\end{table*}
+
+\section{Results 2: Stream models for the Ophiuchus stream}
+
+\section{Discussion}\label{sec:discussion}
+
+[bar not well understood but maybe this stream can help distintuish models with more data -- cite discrepancy in measruements (antoja14, wang12, wang13)]
 
@@
 
+\section{Conclusions}\label{sec:conclusions}
+
+We have found that a qualitative but observationally-motivated potential model for the Galactic bar generates a significant number of chaotic orbits in the vicinity of the recently-discovered Ophiuchus stream.
+By sampling many possible phase-space 6D positions and velocities for the stream centroid from the observational error distribution, we have shown that most possible orbits are strongly chaotic in the time-dependent, barred potential model for the Milky Way, but are mostly regular, eccentric orbits in a static, axisymmetric potential.
+We have shown that---using \streakline\ models of the stream density---the apparent shortness of the stream and the existence of nearby, high-dispersion stars can be explained by chaotic density evolution of the stripped stars.
+This is the first time chaos has been used to explain the morphology and [...] of a stellar stream.
+This work demonstrates the importance of the Galactic bar on the potential of the inner Milky Way and has important implications for future modeling of streams near in this region.
+This result also motivates deep imaging around the Ophiuchus stream to measure the density contrast of the ``fanned'' debris compared to the main stream and to find more candidate stars associated with chaotic tidal debris.
+
 \acknowledgements
@@
 \appendix
-\section{Stream coordinates} \label{sec:rotationmatrix}
-\todo{Add this!}
+\section{Transformation from Galactic to Ophiuchus stream coordinates} \label{sec:rotationmatrix}
+The transformation matrix is approximately represented as
+\begin{equation*}
+\left( \begin{array}{c}
+x \\
+y \\
+z \end{array} \right)_{\rm Oph} \approx
+\left( \begin{array}{ccc}
+0.84922096554 & 0.07001279040 & 0.52337554476\\
+-0.27043653641 & -0.79364259852 &  0.54497294023\\
+0.45352820359 & -0.60434231606 & -0.65504391727
+\end{array} \right) \,
+\left( \begin{array}{c}
+x \\
+y \\
+z \end{array} \right)_{\rm Gal}
+\end{equation*}
+but the precise transformation and coordinate frame is implemented in \python\ using the \package{Astropy} coordinates package.\footnote{See \url{http://adrian.pw/ophiuchus/coordinates} for more information}  This code is hosted on \project{GitHub}.\footnote{\url{https://github.com/adrn/ophiuchus}}
 
@@
 
+% ---------------------------------------------------------------------------------
+\clearpage
+\begin{figure}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/orbitfits-0}
+\caption{ Results from fitting orbits to BHB stars associated with the Ophiuchus stream in each potential model (see also Figure~\ref{fig:orbitfits2}).
+Data are shown as black points with grey error bars.
+Error bars may sometimes be smaller than the point size.
+Lines (blue) show sections of orbits integrated forward and backwards from initial conditions drawn from the posterior samples generated by MCMC (See Section~\ref{sec:orbitfit}).
+Note that the four higher dispersion stars (the four stars with highest longitude) were not used when computing the likelihood and are only shown for completeness.
+}
+\label{fig:orbitfits1}
+\end{center}
+\end{figure}
+
+\clearpage
+\begin{figure}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/orbitfits-1}
+\caption{Same as Figure~\ref{fig:orbitfits2} for the five other potential models.}
+\label{fig:orbitfits2}
+\end{center}
+\end{figure}
+
+% ---------------------------------------------------------------------------------
+\clearpage
+\begin{figure}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/orbit-yz}
+\caption{  TODO }
+\label{fig:orbits-yz}
+\end{center}
+\end{figure}
+
+% ---------------------------------------------------------------------------------
+\clearpage
+\begin{figure}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/lyapunov}
+\caption{ Estimates of the Lyapunov time, $t_\lambda=1/\lambda$, for the mean orbits in each potential model (Section~\ref{sec:results1}).
+All time units are given in units of orbital periods.
+The orbits are integrated from the mean values of orbital parameters from orbit fitting (Section~\ref{sec:orbitfit}).
+In the static, axisymmetric potential model, an orbit fit to the Ophiuchus stream is regular.
+In all barred potential models (bar1--9), the mean orbit is chaotic with a Lyapunov time that is less than 10 orbital periods.
+}
+\label{fig:lyapunov}
+\end{center}
+\end{figure}
+
 \end{document}

### 2015-11-23 ms.tex
remove old title

@@
 
-%\title{Tapping the keg: the Galactic bar peculiarity of the Ophiuchus stream and the Galactic bar}
 \title{Spending too much time at the Galactic bar: Chaotic evolution of the Ophiuchus stream}
@@
 
-We have found that a qualitative but observationally-motivated potential model for the Galactic bar generates a significant number of chaotic orbits in the vicinity of the recently-discovered Ophiuchus stream.
-By sampling many possible phase-space 6D positions and velocities for the stream centroid from the observational error distribution, we have shown that most possible orbits are strongly chaotic in the time-dependent, barred potential model for the Milky Way, but are mostly regular, eccentric orbits in a static, axisymmetric potential.
-We have shown that---using \streakline\ models of the stream density---the apparent shortness of the stream and the existence of nearby, high-dispersion stars can be explained by chaotic density evolution of the stripped stars.
-This is the first time chaos has been used to explain the morphology and [...] of a stellar stream.
-This work demonstrates the importance of the Galactic bar on the potential of the inner Milky Way and has important implications for future modeling of streams near in this region.
-This result also motivates deep imaging around the Ophiuchus stream to measure the density contrast of the ``fanned'' debris compared to the main stream and to find more candidate stars associated with chaotic tidal debris.
+We have found that, with a qualitative but observationally-motivated potential model for the Galactic bar, the orbits of the Ophiuchus stream stars are likely sensitive to the time-dependence and shape of the bar potential.
+For modeling the stream itself, it is therefore crucial to include this component of the Galactic potential.
+By fitting orbits to kinematic data for members of the stream in Milky Way-like potential models, we have found that orbits in the vicinity of the Ophiuchus stream are strongly chaotic for a range of bar parameters (pattern speeds and present-day angles).
+Using mock stellar stream models generated assuming a globular cluster-mass progenitor object, we have shown that the apparent shortness of the stream and the existence of nearby stars with very high velocity dispersion are plausibly explained by chaotic density evolution of the stars stripped from the progenitor object.
+
+This is the first time chaos has been used to explain the morphology of a stellar stream and the first observational evidence for the importance of chaos in the Galactic halo.
+It also highlights the importance of including the Galactic bar in dynamical modeling of the Milky Way's inner halo and has important implications for future modeling of streams near in this region.
+With more Ophiuchus stream members, density and velocity information over a larger region near the stream, and better models for the internal structure of the Galactic bar, careful modeling of this stream could lead to tight constraints on the time-dependent properties of the bar.
 

### 2015-11-26 ms.tex
working on intro

@@
 The Ophiuchus stream \citep{bernard14, sesar15a} is a recently discovered stellar tidal stream that sits above the Galactic bulge at a Galactocentric radius and height $(R,z) \approx (1, 4.3)~{\rm kpc}$.
-All observational evidence suggests that the stream is a totally disrupted globular cluster: The stream stars have (1) a small positional dispersion orthogonal to the extended direction of the stream ($\approx$10 pc); (2) a small velocity dispersion $\approx$0.4 ${\rm km}~{\rm s}^{-1}$; (3) an old stellar population ($\approx$12 Gyr) estimated from isochrone fitting; and (4) no detectable over-density along the stream that could be the progenitor system \citep[][hereafter S15a]{sesar15a}.
+All observational evidence suggests that the stream is a totally disrupted globular cluster: The stream stars have (1) a small positional dispersion orthogonal to the extended direction of the stream ($\approx$10 pc); (2) no detectable over-density along the stream that could be the progenitor system; (3) a small velocity dispersion $\approx$0.4 ${\rm km}~{\rm s}^{-1}$; and (4) an old stellar population ($\approx$12 Gyr) estimated from isochrone fitting \citep[][hereafter S15a]{sesar15a}.
 
-The deprojected length of the visible part of the stream ($\approx$1.5 kpc) is short given the age of its stellar population.
+There are a number of peculiarities about the observed kinematics of probable member stars.
+For example, the de-projected length of the visible part of the stream is short given the age of its stellar population ($\approx$1.5 kpc).
 S15a fit an orbit to the kinematics of the stream stars in a static, axisymmetric model for the gravitational field of the inner Galaxy and ran N-body simulations of globular clusters on this orbit.
-S15a find that---on this orbit---the observed portion of the stream must have been formed in the last $\lesssim$400 Myr for the stream to remain as short as it is observed.
+S15a find that---on this orbit---the portion of the stream visible as an over-density in main-sequence stars must have been formed in the last $\lesssim$400 Myr for the stream to remain as short as it is observed.
 This dynamical age is at odds with the old stellar population.
-More recently, a significant observational effort has followed-up nearly all blue horizontal branch (BHB) star candidates within XX deg of the stream to obtain radial velocities to identify more stream members \citep[][hereafter S15b]{sesar15b}.
+Another puzzle is the existence of blue horizontal branch (BHB) stars close to the stream (within XX deg) with similar radial velocities, but with a large dispersion in both sky position and velocity \citep[][hereafter S15b]{sesar15b}.
 The stream has a very distinct and large line-of-sight velocity ($\approx$290 \kms) and is therefore easily detected above the background halo population.
-Of the \todo{XX} stars followed up, six have velocities significantly discrepant with the halo population.
-Of these six stars, two lie within the previously detected extent of the stream and have velocities consistent with the previously identified stream members.
-The remaining four stars lie close to an extrapolation of the stream, have velocities much larger than halo stars in this region ($v_{\rm los}>230$ \kms), but have a velocity dispersion $\approx$75 times the measured internal velocity dispersion of the stream stars and positional spread much larger than the orthogonal positional dispersion of the main part of the stream.
-These low-density, high-dispersion features were not modeled in S15a and N-body simulations on the orbit fit from the previous work do not  reproduce these features.
+Four stars have been detected with line-of-sight velocities $>230~\kms$ that lie close to an extrapolation of the stream on the sky, but have a velocity dispersion $\approx$75 times larger than the measured internal velocity dispersion of the stream stars.
+These stars hint at the existence of associated low-density, high-dispersion features that were not modeled in S15a and are not predicted by the $N$-body simulations from this prior work.
 
-It is not surprising that the orbit fit in the static, axisymmetric potential model used in S15a leads to a dynamical puzzle and failed to predict the new features: it is now well-known that the Galactic bulge contains a triaxial, bar-like structure several kpc in size \citep[e.g.,][]{blitzXX, wegg13, MANY}.
-Given the proximity of the stream to the center of the Galaxy, the time-dependent, triaxial potential of the Galactic bar must be taken into account when modeling the orbit of the progenitor system.
-The presence of a bar-like perturbation to the potential will change the possible orbit of the stream progenitor and will drastically change the resonance structure of orbits in the inner galaxy \citep{athanassoula, portail15b, zotos12}, potentially introducing a significant number of chaotic orbits in this region \citep{weinberg15}.
+The orbit fit and $N$-body simulations in S15a used a static, axisymmetric potential to represent the Milky Way potential, but it is now well-known that the Galactic bulge contains a triaxial, bar-like structure several kpc in size \citep[e.g.,][]{blitzXX, wegg13, MANY}.
+Given the proximity of the stream to the center of the Galaxy, the time-dependent, triaxial potential of the Galactic bar must be taken into account when modeling the orbit of the Ophiuchus stream.
+The presence of a bar-like perturbation to the potential will change the orbit of the stream progenitor and the orbit structure in the inner galaxy \citep{athanassoula, portail15b, zotos12}.
+Bar-like features can also introduce a significant number of chaotic orbits in their vicinity\citep{weinberg15}.
 
@@
 The mean orbits in the barred potentials are all strongly chaotic.
-We have tried computing the frequency diffusion rate for these orbits but have found that, over two consecutive integration windows, the frequency recovery fails because the frequency spectrum changes dramatically.
-Over timescales of $\approx$10--20 orbital periods.
-
-% \section{Results ??: Characterizing chaos ..?}
-%Using the orbit initial conditions described above (Section~\ref{sec:orbit-grids}), we integrate orbits and compute the frequency diffusion rate for the static potential and for XX barred models with varied initial bar angle and pattern speed.
-We find that [...]
-
-% Full table with everything - boring
-%\begin{table*}[ht]
-%\footnotesize
-%\begin{center}
-%	\begin{tabular}{ccccccccc}
-%	\toprule
-%	name & $\phi_2$ [deg] & $d$ [kpc] & $\mu_l$ [mas yr$^{-1}$] & $\mu_b$ [mas yr$^{-1}$] & $v_r$ [km s$^{-1}$] & $s_{\phi_2}$ [deg] & $s_{d}$ [kpc] & $s_{v_r}$ [km s$^{-1}$]\\\midrule
-%static & $-0.03\pm0.05$ & $8.35\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.1\pm0.9$ & $0.20\pm0.04$ & $0.31\pm0.10$ & $2.9\pm0.8$\\
-%bar1 & $-0.03\pm0.05$ & $8.40\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.0\pm0.9$ & $0.21\pm0.05$ & $0.31\pm0.14$ & $3.2\pm0.8$\\
-%bar2 & $-0.03\pm0.05$ & $8.40\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $288.9\pm1.1$ & $0.20\pm0.04$ & $0.33\pm0.09$ & $3.3\pm0.9$\\
-%bar3 & $-0.04\pm0.06$ & $8.40\pm0.06$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.0\pm1.0$ & $0.21\pm0.05$ & $0.35\pm0.09$ & $3.2\pm0.8$\\
-%bar4 & $-0.03\pm0.06$ & $8.41\pm0.06$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.1\pm1.0$ & $0.21\pm0.05$ & $0.36\pm0.10$ & $3.4\pm0.9$\\
-%bar5 & $-0.03\pm0.05$ & $8.41\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.1\pm0.9$ & $0.21\pm0.04$ & $0.35\pm0.09$ & $3.3\pm0.8$\\
-%bar6 & $-0.03\pm0.05$ & $8.40\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.0\pm1.0$ & $0.21\pm0.05$ & $0.35\pm0.10$ & $3.2\pm0.8$\\
-%bar7 & $-0.04\pm0.05$ & $8.41\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.0\pm1.0$ & $0.21\pm0.05$ & $0.33\pm0.12$ & $3.4\pm0.9$\\
-%bar8 & $-0.03\pm0.05$ & $8.42\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.1\pm1.0$ & $0.20\pm0.04$ & $0.34\pm0.09$ & $3.4\pm0.9$\\
-%bar9 & $-0.02\pm0.05$ & $8.41\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $288.9\pm1.0$ & $0.21\pm0.04$ & $0.35\pm0.09$ & $3.3\pm0.8$\\
-%	\bottomrule
-%	\end{tabular}
-%	\caption{Estimated mean and standard deviation of the marginal posterior distributions over each parameter in our orbit fit likelihood.
-All values are nearly identical because the potentials are very similar over the short length of the stream.
-This table is mainly included as a sanity check---we have made samples from the full posterior distribution available with this article and provide code to transform to and from stream coordinates (see http://adrian.pw/ophiuchus/coordinates for more information).\label{tbl:param-means} }
-%\end{center}
-%\end{table*}
+We have tried computing the frequency diffusion rate for these orbits but have found that, over two consecutive integration windows, the frequency recovery fails because the frequency spectrum changes dramatically over timescales of $\approx$10--20 orbital periods.
 
@@
 
+The model streams presented here do not reproduce all observed features of the Ophiuchus stream.
+Instead, these results illustrate that chaotic evolution of tidal debris can plausibly explain the peculiar features of the stream.
+If the cluster progenitor was on a regular orbit, it would have to have disrupted entirely within the last $\approx$300 Myr (Figure~\ref{TODO}) in order to explain the shortness and density profile.
+In addition, the four most recently identified BHB stars with similar distances and line-of-sight velocities would have to be (highly unlikely) chance alignments of halo stars.
+If instead the progenitor were on a chaotic orbit (because of the influence of the Galactic bar), (1) stars stripped early will have ``fanned'' out and would thus be harder to observe and (2) the nearby, high-velocity-dispersion BHB stars can be naturally explained by this chaotic stream-fanning.
+We consider the second scenario to be more plausible: our understanding of the formation and evolution of the Ophiuchus stream is that the progenitor object has been orbiting and steadily losing stars over the last several Gigayears, but only the stars stripped from the most recent few pericentric passages remain coherent enough to be detected as a stream-like over-density in the PS1 data.
+
 [bar not well understood but maybe this stream can help distintuish models with more data -- cite discrepancy in measruements (antoja14, wang12, wang13)]
+
+[Gaia, Don't do distance cut that Brani did, chemistry]
+
+[explored in future work: what are implications of the existence of this structure for total population]
 

### 2015-12-11 ms.tex
small changes
words about modified fardal method

@@
 
-There are a number of peculiarities about the observed kinematics of probable member stars.
+There are a number of peculiarities about the observed kinematics of probable member stars of the Ophiuchus stream.
 For example, the de-projected length of the visible part of the stream is short given the age of its stellar population ($\approx$1.5 kpc).
@@
 	\item the increased positional dispersion of the four new candidate members from S15b;
-	\item the much larger velocity dispersion of the S15b stars.
+	\item the large velocity dispersion of the S15b stars.
 \end{enumerate}
@@
 The full expression for the posterior probability, $p(\bs{s}, \bs{w}_0, t_b, t_f \given \{\bs{x}_i\}, \{\bs{\sigma}_i\}, \Phi)$, is the joint likelihood (Equation~\ref{eq:likelihood}) multiplied by all priors described above (Equations~\ref{eq:prior1}--\ref{eq:prior2}).
-We use an ensemble Markov Chain Monte Carlo (MCMC) algorithm \citep{goodman10} implemented in \python\ (\package{emcee}) to draw samples from this posterior distribution \citep{foremanmackey13}.
+We use an ensemble Markov Chain Monte Carlo (MCMC) algorithm \citep{goodman10} implemented in \python\ (\package{emcee}) to generate samples from this posterior distribution \citep{foremanmackey13}.
 The algorithm uses an ensemble of individual ``walkers'' to adapt to the geometry of the parameter-space being explored.
-In all cases, we use 80 walkers.
+In all cases, we use 80 walkers (8 times the number of parameters).
 
-To initialize these walkers, we first run an optimization routine to maximize the likelihood.
-In detail, we use the Powell algorithm implemented in \package{Scipy} \citep{powell64, scipy} to minimize the negative, log-likelihood.
+To initialize these walkers, we first run an optimization routine to maximize the likelihood: We use the Powell algorithm implemented in \package{Scipy} \citep{powell64, scipy} to minimize the negative, log-likelihood.
 To generate initial conditions for the walkers, we sample from Gaussian distributions centered on the maximum likelihood values.
@@
 
-For each potential, we run the MCMC walkers for a burn-in period of 128 steps and then re-initialize the walkers from their positions at the end of this run.
+For each potential, we run the MCMC walkers for a burn-in period of 512 steps and then re-initialize the walkers from their positions at the end of this run.
 This erases any relics of the initialization procedure outlined above.
-After burn-in, we run the walkers for an additional 256 steps.
+After burn-in, we run the walkers for an additional 512 steps.
+For each parameter, we compute the autocorrelation times, $\tau$, of the Markov chains and thin the chains by taking every $2\tau$ sample.
+This reduces the number of samples, but ensures that our posterior samples are effectively independent.
+
 %
@@
 
-To generate mock stellar streams, we use the method presented in \citep{fardal14}: star particles are `released' from a progenitor system near the Lagrange points with some dispersion in position and velocity set by the mass and orbit of the progenitor (this is similar to the \streakline\ method of \citet{kuepper12} but with finite dispersions).
-To generate progenitor orbits, we draw samples from the posterior probability distributions over orbital parameters from fitting orbits to the stream star members (Section~\ref{sec:orbit-fit}).
+To generate mock stellar streams, we use the method presented in \citep{fardal14}: star particles are `released' from a progenitor system near the Lagrange points with some dispersion in position and velocity set by the mass and orbit of the progenitor.
+To generate progenitor orbits, we draw samples from the posterior probability distributions over orbital parameters from fitting orbits to the stream star members (Section~\ref{sec:orbitfit}).
 For a given progenitor orbit---the 6D position of the orbit today---we integrate the orbit backwards in time for a given integration period.
@@
 This method has been shown to reproduce the morphologies of $N$-body simulations of stellar streams, but requires far less computing time because it relies only on integrating test-particle orbits.
+
+We make one modification to this method based on the idea that the Ophiuchus stream progenitor has been fully-disrupted.
+We add an additional parameter to the stream generation routine to specify the time of disruption, $\tau_d$.
+At this time, we set the offset of the Lagrange point to 0: In terms of the parametrization in \citep{fardal14}, we set $k_r = 0$ and $k_{vt}=0$ but preserve the dispersion in the release radius and velocity.
+Any star released after $\tau_d$ is released with a small dispersion around the progenitor orbit but no offset.
 
@@
 \section{Results 2: Stream models for the Ophiuchus stream}
+
+- increased positional dispersion may be just projection
+
+- if orbit is correct, at current time we're looking roughly onto the instantaneous orbital plane.
+if thats the case, the straightness of the stream is interesting because that means the thing must have totally disrupted (no sign of L pts.)
+
+-
 

### 2015-12-12 ms.tex
some minor edits

@@
 This dynamical age is at odds with the old stellar population.
-Another puzzle is the existence of blue horizontal branch (BHB) stars close to the stream (within XX deg) with similar radial velocities, but with a large dispersion in both sky position and velocity \citep[][hereafter S15b]{sesar15b}.
+Another puzzle is the existence of blue horizontal branch (BHB) stars close to the stream (within a few degrees) with similar radial velocities, but with a large dispersion in both sky position and velocity \citep[][hereafter S15b]{sesar15b}.
 The stream has a very distinct and large line-of-sight velocity ($\approx$290 \kms) and is therefore easily detected above the background halo population.
@@
 Along certain chaotic orbits, the stream stars will spread much faster in 3D position than from ordinary phase-mixing and, depending on the orbital phase at which the stream is observed, may develop large, low-density ``fans'' of stars at the ends of a stream \citep{apw15-chaos}.
-In this work, we study whether stream-fanning---chaotic or simply from density evolution in a triaxial, time-dependent potential---can explain the observed properties of the Ophiuchus stream.
-In particular, we consider a model successful if it qualitatively reproduces:
+In this work, we study whether stream-fanning---chaotic or simply from density evolution in a triaxial, time-dependent potential---could plausibly explain the observed properties of the Ophiuchus stream.
+In particular, we consider whether such models can reproduce:
 \begin{enumerate}
@@
 The key feature of the potential that we would like to capture is the time-dependence and triaxiality of the Galactic bar.
-Recent work has used stellar number counts of \todo{XX} stars in the Galactic bulge to constrain dynamical models of the bar \citep{portail15}.
-Measurements of the total mass of the bar feature from this study are largely consistent with past work \citep[e.g.,][]{XX}, however the measured pattern speed and present bar angle are significantly discrepant and this difference is not fully understood.
+Recent work has used stellar number counts of Red Clump giant stars in the Galactic bulge to constrain dynamical models of the bar \citep{portail15}.
+Measurements of the total mass of the bar feature from this study are largely consistent with past work \citep[e.g.,][]{wang12, XX}, however the measured pattern speed and present bar angle are significantly discrepant and this difference is not fully understood.
 We construct a parametrized potential model consisting of a triaxial, time-dependent (rotating) bulge component added to simple models for the disk and halo of the Milky Way.
@@
 The mass of the halo is fixed and the mass of the disk and bar are varied in order to qualitatively reproduce the flatness and amplitude of the circular velocity curve of the Milky Way \citep{bovy12}.
-Figure~\ref{fig:circ-vel-barred} shows the circular velocity along the line connecting the Sun to the Galactic center in this model (the Galactic $x$ axis).
-Figure~\ref{fig:surface-density-barred} shows contours of constant surface density for a face-on (left) and edge-on (right) view of this potential model with the bar angle set to $20^\circ$ \citep[compare to, e.g., Figure 3 in][]{portail15}.
+Figure~\ref{fig:potentials}, top left shows the circular velocity along the line connecting the Sun to the Galactic center in this model (the Galactic $x$ axis).
+Figure~\ref{fig:potentials}, top right shows contours of constant surface density for a face-on (left) and edge-on (right) view of this potential model with the bar angle set to $20^\circ$ \citep[compare to, e.g., Figure 3 in][]{portail15}.
 We consider a grid of nine parameter combinations of bar angle and pattern speed.
@@
 Parameters for this potential model are given in Table~\ref{tbl:potential-params-static}.
-Figure~\ref{fig:circ-vel-barred} shows the circular velocity along the line connecting the Sun to the Galactic center in this model (the Galactic $x$ axis).
-Figure~\ref{fig:surface-density-static} shows contours of constant surface density for a face-on (left) and edge-on (right) view of this potential model.
+Figure~\ref{fig:potentials}, bottom left shows the circular velocity along the line connecting the Sun to the Galactic center in this model (the Galactic $x$ axis).
+Figure~\ref{fig:potentials}, bottom right shows contours of constant surface density for a face-on (left) and edge-on (right) view of this potential model.
 
@@
 
-\section{Results I: Orbit fits}\label{sec:results1}
+\section{Results I: Orbit fits and chaos}\label{sec:results1}
 
-Figures~\ref{fig:orbitfits1}--\ref{fig:orbitfits2} summarize our results from fitting orbits to the BHB stream stars.
+Figure~\ref{fig:orbitfits} summarizes our results from fitting orbits to the BHB stream stars.
 Shown in each panel are the Ophiuchus stream star members (black points) and samples from the  posterior probability over orbital parameters (blue lines).
-The end-to-end integration time of the orbit over the observed extent of the stream is only $\approx$6 Myr, so the derived orbits are extremely similar in all potentials (the time-dependence of the bar potential is not significant over such short timescales).
+We only show one of the barred potentials: The end-to-end integration time of the orbit over the observed extent of the stream is only $\approx$6 Myr, so the derived orbits are extremely similar in all potentials (the time-dependence of the bar potential is not significant over such short timescales).
 The orbital periods are typically $\approx$150--250 Myr.
@@
 Though the orbital parameter values in observed coordinates are very similar between each potential model, the resulting orbits are quite different.
-Figure~\ref{fig:orbits-yz} shows projections of the ``mean'' orbits from each potential model; for each posterior distribution from orbit fitting, we take the mean values of the orbital parameters and convert to Galactocentric coordinates (e.g., Table~\ref{tbl:param-means}).
+Figure~\ref{fig:orbits-yz} shows projections of the ``mean'' orbits in each potential model; for the posterior samples in each potential, we take the mean values of the orbital parameters and convert to Galactocentric coordinates (e.g., Table~\ref{tbl:param-means}).
 For each orbit, we compute the maximum Lyapunov exponent (MLE, $\lambda$) to assess whether the orbit is chaotic.
@@
 Figure~\ref{fig:lyapunov} shows numerical estimates of the Lyapunov time ($t_\lambda = 1/\lambda$) for the mean orbits in each potential model: in the static potential the orbit appears regular (black, diagonal line), but in all barred potentials the MLE estimates converge to finite values in the range $\approx$2--8 orbital periods.
-\todo{Should I compute Lyapunov exponents for samples from the posterior, not just the mean orbit?}
+\todo{Should I compute Lyapunov exponents for samples from the posterior, not just the mean orbit?
+What fraction of the orbits sampled from the posterior and what are distribution of Lyapunov times?}
 
@@
 
--
+- only run on mean streams, show all -- discuss that morphology is very sensitive to the bar properties
+
+- some of the models have large dispersion, some have etc...
+
+- with better data, perhaps we can constrain the bar?
 
@@
 
-[bar not well understood but maybe this stream can help distintuish models with more data -- cite discrepancy in measruements (antoja14, wang12, wang13)]
+[bar not well understood but maybe this stream can help distinguish models with more data -- cite discrepancy in measruements (antoja14, wang12, wang13)]
 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/barred-circ-vel}
-\caption{Circular velocity curve along the Sun-Galactic center line for the barred MW potential model.
+\includegraphics[width=\textwidth]{figures/potentials-four}
+\caption{{\it Left column:} Circular velocity curves along the Sun-Galactic center line for a representative barred MW potential model (top left) and for the static MW potential model (bottom left).
+Solid black line shows total (sum of all components), lines below show a decomposition by potential component.
+Vertical grey bar shows approximate position of the Sun, horizontal grey bar shows roughly the range in measured circular velocity of the Sun.
+{\it Right column:} Contours of constant surface density for a barred MW potential (top right) and the static MW potential model (bottom right).
+Four contours are drawn per decade in surface density between $10^7$ and $10^{12}~\msun~{\rm kpc}^{-3}$.
+Note the perturbation from the bar potential within Galactocentric radius $r \lesssim 4~{\rm kpc}$.
+The Sun's position is indicated by the `$\odot$' symbol.
+}
+\label{fig:potentials}
+\end{center}
+\end{figure}
+
+% ---------------------------------------------------------------------------------
+%\clearpage
+%\begin{figure}[p]
+%\begin{center}
+%\includegraphics[width=\textwidth]{figures/barred-circ-vel}
+%\caption{Circular velocity curve along the Sun-Galactic center line for the barred MW potential model.
 Solid black line shows total (sum of all components), lines below show a decomposition by potential component.
 Vertical grey bar shows approximate position of the Sun, horizontal grey bar shows roughly the range in measured circular velocity of the Sun.}
-\label{fig:circ-vel-barred}
-\end{center}
-\end{figure}
+%\label{fig:circ-vel-barred}
+%\end{center}
+%\end{figure}
+%
+%% ---------------------------------------------------------------------------------
+%\clearpage
+%\begin{figure}[p]
+%\begin{center}
+%\includegraphics[width=\textwidth]{figures/barred-surface-density-contour}
+%\caption{Contours of constant surface density for two projections of the barred MW potential model.
+For both face-on view (left panel) and edge-on view (right panel), four contours are drawn per decade in surface density between $10^7$ and $10^{12}~\msun~{\rm kpc}^{-3}$.
+Note the perturbation from the bar potential within Galactocentric radius $r \lesssim 4~{\rm kpc}$.
+The Sun's position is indicated by the `$\odot$' symbol.}
+%\label{fig:surface-density-barred}
+%\end{center}
+%\end{figure}
+%
+%% ---------------------------------------------------------------------------------
+%\clearpage
+%\begin{figure}[p]
+%\begin{center}
+%\includegraphics[width=\textwidth]{figures/static-circ-vel}
+%\caption{Same as Figure~\ref{fig:circ-vel-barred} for the static, axisymmetric MW potential model. }
+%\label{fig:circ-vel-static}
+%\end{center}
+%\end{figure}
+%
+%% ---------------------------------------------------------------------------------
+%\clearpage
+%\begin{figure}[p]
+%\begin{center}
+%\includegraphics[width=\textwidth]{figures/static-surface-density-contour}
+%\caption{Same as Figure~\ref{fig:surface-density-barred} for the static, axisymmetric MW potential model. }
+%\label{fig:surface-density-static}
+%\end{center}
+%\end{figure}
 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/barred-surface-density-contour}
-\caption{Contours of constant surface density for two projections of the barred MW potential model.
-For both face-on view (left panel) and edge-on view (right panel), four contours are drawn per decade in surface density between $10^7$ and $10^{12}~\msun~{\rm kpc}^{-3}$.
-Note the perturbation from the bar potential within Galactocentric radius $r \lesssim 4~{\rm kpc}$.
-The Sun's position is indicated by the `$\odot$' symbol.}
-\label{fig:surface-density-barred}
-\end{center}
-\end{figure}
-
-% ---------------------------------------------------------------------------------
-\clearpage
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/static-circ-vel}
-\caption{Same as Figure~\ref{fig:circ-vel-barred} for the static, axisymmetric MW potential model. }
-\label{fig:circ-vel-static}
-\end{center}
-\end{figure}
-
-% ---------------------------------------------------------------------------------
-\clearpage
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/static-surface-density-contour}
-\caption{Same as Figure~\ref{fig:surface-density-barred} for the static, axisymmetric MW potential model. }
-\label{fig:surface-density-static}
-\end{center}
-\end{figure}
-
-% ---------------------------------------------------------------------------------
-\clearpage
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/orbitfits-0}
-\caption{ Results from fitting orbits to BHB stars associated with the Ophiuchus stream in each potential model (see also Figure~\ref{fig:orbitfits2}).
+\includegraphics[width=\textwidth]{figures/orbitfits}
+\caption{ Results from fitting orbits to BHB stars associated with the Ophiuchus stream in the static and barred potential models.
 Data are shown as black points with grey error bars.
@@
 }
-\label{fig:orbitfits1}
-\end{center}
-\end{figure}
-
-\clearpage
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/orbitfits-1}
-\caption{Same as Figure~\ref{fig:orbitfits2} for the five other potential models.}
-\label{fig:orbitfits2}
+\label{fig:orbitfits}
 \end{center}

### 2015-12-15 ms.tex
figures and captions and more, oh my

@@
-% \documentclass{emulateapj}
+%\documentclass{emulateapj}
 \documentclass[letterpaper,12pt,preprint]{aastex}
@@
 \newcommand{\todo}[1]{{\color{red} TODO: #1}}
+
+\newcommand{\lyapexp}{\lambda_{\rm max}}
+\newcommand{\lyapt}{t_\lambda}
 
@@
 
+\todo{Move a lot of this to appendix?}
+
 We include three nuisance parameters in our likelihood to account for the internal dispersion of the stream: in observed coordinates, these are the on-sky positional dispersion, $s_{\phi_2}$, a distance (modulus) dispersion, $s_{\DM}$, and a radial velocity dispersion, $s_{v_r}$ (the proper motion uncertainties are sufficiently large that we can't resolve the velocity dispersion in these coordinates).\footnote{We assume that the dispersion in these coordinates is constant over the observed (short) section of the stream.
@@
 \end{align}
-We assume the data from each star is independent and identically distributed (i.i.d.) so that the full likelihood is the product over the likelihoods for each star:
+We assume the data from each star is independent and identically distributed (i.i.d.) so that the full likelihood is the product over the likelihoods for all $N$ stars:
 \begin{equation}
-	 p(\{\bs{x}_i\} \given \{\bs{\sigma}_i\}, \bs{s}, \bs{\theta}) = \prod_i p(\bs{x}_i \given \bs{\sigma}_i, \bs{s}, \bs{\theta}).\label{eq:likelihood}
+	 p(\{\bs{x}_i\} \given \{\bs{\sigma}_i\}, \bs{s}, \bs{\theta}) = \prod_i^N p(\bs{x}_i \given \bs{\sigma}_i, \bs{s}, \bs{\theta}).\label{eq:likelihood}
 \end{equation}
@@
 
-To generate mock stellar streams, we use the method presented in \citep{fardal14}: star particles are `released' from a progenitor system near the Lagrange points with some dispersion in position and velocity set by the mass and orbit of the progenitor.
-To generate progenitor orbits, we draw samples from the posterior probability distributions over orbital parameters from fitting orbits to the stream star members (Section~\ref{sec:orbitfit}).
+To generate mock stellar streams, we use the method presented in \citep{fardal14}: star particles are `released' from a progenitor system near the Lagrange points with a dispersion in position and velocity that is set by the mass and orbit of the progenitor.
+We draw samples from the posterior probability distributions over orbital parameters from fitting orbits to the stream star members and use these as the progenitor orbital parameters (Section~\ref{sec:orbitfit}).
 For a given progenitor orbit---the 6D position of the orbit today---we integrate the orbit backwards in time for a given integration period.
@@
 The position of the Lagrange points and the scale of the dispersion in position and velocity are set by the progenitor mass, $m$.
-The star particles are drawn from Gaussians centered on the Lagrange points (in position) and the progenitor (in velocity) and the full parametrization of the release distribution is given in \cite{fardal14}.
+\todo{Slightly more detail here about Fardal's parametrization} The star particles are drawn from Gaussians centered on the Lagrange points (in position) and the progenitor (in velocity) and the full parametrization of the release distribution is given in \cite{fardal14}.
 This method has been shown to reproduce the morphologies of $N$-body simulations of stellar streams, but requires far less computing time because it relies only on integrating test-particle orbits.
@@
 Figure~\ref{fig:orbitfits} summarizes our results from fitting orbits to the BHB stream stars.
-Shown in each panel are the Ophiuchus stream star members (black points) and samples from the  posterior probability over orbital parameters (blue lines).
+Shown in each panel are the Ophiuchus stream star members (black points) and orbits integrated from samples from the posterior probability over orbital parameters (blue lines).
 We only show one of the barred potentials: The end-to-end integration time of the orbit over the observed extent of the stream is only $\approx$6 Myr, so the derived orbits are extremely similar in all potentials (the time-dependence of the bar potential is not significant over such short timescales).
-The orbital periods are typically $\approx$150--250 Myr.
+The orbital periods are typically $\approx$170--200 Myr with pericenters $r_p \approx 4~{\rm kpc}$ and apocenters $r_a \approx 12$--$15~{\rm kpc}$.
+Though the coordinate and velocity parameter values in observed coordinates are very similar between each potential model, the resulting orbits are quite different.
+For the posterior samples in each potential, we take the mean values of the coordinate and velocity parameters and convert to Galactocentric coordinates (e.g., Table~\ref{tbl:param-means}).
+Figure~\ref{fig:orbits-yz} shows projections of these ``mean'' orbits in each potential model.
 
-Though the orbital parameter values in observed coordinates are very similar between each potential model, the resulting orbits are quite different.
-Figure~\ref{fig:orbits-yz} shows projections of the ``mean'' orbits in each potential model; for the posterior samples in each potential, we take the mean values of the orbital parameters and convert to Galactocentric coordinates (e.g., Table~\ref{tbl:param-means}).
-For each orbit, we compute the maximum Lyapunov exponent (MLE, $\lambda$) to assess whether the orbit is chaotic.
-Though not predictive for the timescale over which chaos is important for tidal debris, the MLE is still an appropriate indicator of chaos \citep{apw15-chaos}.
-Figure~\ref{fig:lyapunov} shows numerical estimates of the Lyapunov time ($t_\lambda = 1/\lambda$) for the mean orbits in each potential model: in the static potential the orbit appears regular (black, diagonal line), but in all barred potentials the MLE estimates converge to finite values in the range $\approx$2--8 orbital periods.
-\todo{Should I compute Lyapunov exponents for samples from the posterior, not just the mean orbit?
-What fraction of the orbits sampled from the posterior and what are distribution of Lyapunov times?}
+For the posterior samples in each potential, we also compute the maximum Lyapunov exponent (MLE, $\lyapexp$) and corresponding Lyapunov \emph{time} ($\lyapt = 1/\lyapexp$) to assess whether the orbit is chaotic.
+For strongly chaotic orbits, the Lyapunov time is still an appropriate indicator of chaos and of the timescale over which chaos is important for tidal debris \citep{apw15-chaos}.
+Figure~\ref{fig:lyapunov} shows numerical estimates of the Lyapunov time for the mean orbits in each potential model: In the static potential the orbit appears regular (black, diagonal line), but in all barred potentials the MLE estimates converge to finite values in the range $\approx$2--8 orbital periods.
+Figure~\ref{XX} shows distributions of Lyapunov times for orbits drawn from the posterior distributions from orbit fitting in each potential model.
+\todo{What else to say?} \todo{What fraction of the orbits sampled from the posterior are chaotic and what are distribution of Lyapunov times?
+}
+\todo{PLOT MEAN ORBIT VALUE AS VERTICAL LINE}
 
-The mean orbits in the barred potentials are all strongly chaotic.
-We have tried computing the frequency diffusion rate for these orbits but have found that, over two consecutive integration windows, the frequency recovery fails because the frequency spectrum changes dramatically over timescales of $\approx$10--20 orbital periods.
+The orbits in the barred potentials are all strongly chaotic.
+We have tried computing the frequency diffusion rate for these orbits but have found that, over consecutive integration windows, the frequency recovery fails or is unreliable because the frequency spectrum changes dramatically over timescales of $\approx$10--20 orbital periods.
 
@@
 
-- increased positional dispersion may be just projection
+We generate mock streams along each of the mean orbits from the orbit fitting procedure (defined above).
+We use the method outlined above (Section \ref{sec:mocks}) to generate the streams and set the free parameters as follows: (1) we evolve the progenitors for 6 Gyr along each orbit, (2) we release star particles every 0.5 Myr (uniformly in time) to densely sample the final density distribution, (3) we set the progenitor mass to $m=10^4~\msun$, and (4) we set the disruption time of the progenitor equal to the last time at which a pericenter and disk crossing coincide (in each case, this is at $t \approx -200~{\rm Myr}$).
 
-- if orbit is correct, at current time we're looking roughly onto the instantaneous orbital plane.
-if thats the case, the straightness of the stream is interesting because that means the thing must have totally disrupted (no sign of L pts.)
+Figures~\ref{fig:mockstream0}--\ref{fig:mockstream1} show the final particle positions and line-of-sight velocities in heliocentric coordinates for these mock streams (grey points).
+Vertical, dashed lines show the approximate extent of the part of the stream visible in main-sequence stars (excluding the BHB stars from S15b).
+We have applied the same distance-modulus cut to the points in these figures as used to select the BHB stars identified as stream members by S15b (see Section~2 in S15b).
+There are a few interesting features to note:
+\begin{enumerate}
+	\item Even in the static potential (Figure~\ref{fig:mockstream0}, leftmost column), there is an increased dispersion in sky position for the stream stars towards higher Galactic longitudes, $l$.
+This is a projection effect: The portion of the stream at larger $l$ is closer and almost pointing towards the Sun, as demonstrated in Figure~\ref{fig:mockstreamxyz}.
+Figure~\ref{fig:mockstreamxyz} shows $x$-$z$ projections of the star particles in Galactocentric Cartesian coordinates (grey) along with the position of the Sun (symbol at $(x,z)=(-8,0)~{\rm kpc}$) and the corners of the Heliocentric, sky-position plots of Figures~\ref{fig:mockstream0}--\ref{fig:mockstream1}.
+	\item The density of the mock stream in the static potential decreases slowly rather than abruptly as is observed.
+This is more clearly seen in terms of the \todo{linear density of the stream}, shown in Figure~\ref{fig:lineardensity}.
+....
+\todo{Figure showing some brightness density vs longitude?}
+	\item For the barred potentials, the stream morphology is very sensitive to the properties of the bar (especially the pattern speed) and to the initial conditions of the orbit.
+In addition to generating mock streams on the mean orbits, we have run similar simulations for random draws from the posterior distributions over orbit fits.
+We have found that the morphology can vary significantly between different samples (because these are strongly chaotic orbits), but the overall characteristics remain similar: along more strongly chaotic orbits, the debris ``fans'' more and the apparent dense part of the model streams is shorter (see, for example, Figure~\ref{fig:mockstreamxyz}, panels bar2 vs.
+bar3).
+\end{enumerate}
 
-- only run on mean streams, show all -- discuss that morphology is very sensitive to the bar properties
+To generate each mock stream, the star particles are released uniformly in time.
+The stream star particles at present day---e.g.
+the ``model'' points---are then a sampling of the stream model (phase-space) density field given a uniform stripping time distribution.
+We can then ask, given the data---a noisy sampling of the true density field---what stripping time distribution would reproduce the observed density field (e.g., the positions and velocities of the BHB stars).
+This corresponds to computing a weight for each $k$ model star particle with phase-space position $\bs{z}_k$ and release time, $t_{s,k}$:
+\begin{align}
+	w_k = \frac{1}{N} \, \sum^N_i \, p(\bs{z}_k \given \bs{x}_i, \bs{\sigma}_i, \bs{s})
+\end{align}
+(see Section~\ref{sec:orbitfit} for variable definitions).
 
-- some of the models have large dispersion, some have etc...
+If we assume that the data are a noisy sampling of the true (phase-space) density field,
 
-- with better data, perhaps we can constrain the bar?
+Each model point We do this by computing the \todo{what is it called?} for each model point, $k$:  \citep{hogg10}.
+
+As a more quantitative / alternative exploration of the density ....... we consider the mass-loss histories...
 
@@
 Instead, these results illustrate that chaotic evolution of tidal debris can plausibly explain the peculiar features of the stream.
-If the cluster progenitor was on a regular orbit, it would have to have disrupted entirely within the last $\approx$300 Myr (Figure~\ref{TODO}) in order to explain the shortness and density profile.
+If the cluster progenitor was on a regular orbit, it would have to have disrupted entirely within the last $\approx$300 Myr in order to explain the shortness and density profile.
 In addition, the four most recently identified BHB stars with similar distances and line-of-sight velocities would have to be (highly unlikely) chance alignments of halo stars.
-If instead the progenitor were on a chaotic orbit (because of the influence of the Galactic bar), (1) stars stripped early will have ``fanned'' out and would thus be harder to observe and (2) the nearby, high-velocity-dispersion BHB stars can be naturally explained by this chaotic stream-fanning.
+If instead the progenitor were on a chaotic orbit (because of the influence of the Galactic bar): (1) stars stripped early will have ``fanned'' out and would thus be harder to observe and (2) the nearby, high-velocity-dispersion BHB stars can be naturally explained by this chaotic stream-fanning.
 We consider the second scenario to be more plausible: our understanding of the formation and evolution of the Ophiuchus stream is that the progenitor object has been orbiting and steadily losing stars over the last several Gigayears, but only the stars stripped from the most recent few pericentric passages remain coherent enough to be detected as a stream-like over-density in the PS1 data.
+
+- with better data, perhaps we can constrain the bar?
 
@@
 
+\todo{Still to early to say for sure, but a plausible explanation...}
+
 \section{Conclusions}\label{sec:conclusions}
 
-We have found that, with a qualitative but observationally-motivated potential model for the Galactic bar, the orbits of the Ophiuchus stream stars are likely sensitive to the time-dependence and shape of the bar potential.
-For modeling the stream itself, it is therefore crucial to include this component of the Galactic potential.
+We have shown that, with a qualitative but observationally-motivated potential model for the Galactic bar, the orbits of the Ophiuchus stream stars are likely sensitive to the time-dependence and shape of the bar potential.
+For modeling the stream density itself, it is therefore crucial to include this component of the Galactic potential.
 By fitting orbits to kinematic data for members of the stream in Milky Way-like potential models, we have found that orbits in the vicinity of the Ophiuchus stream are strongly chaotic for a range of bar parameters (pattern speeds and present-day angles).
@@
 It also highlights the importance of including the Galactic bar in dynamical modeling of the Milky Way's inner halo and has important implications for future modeling of streams near in this region.
-With more Ophiuchus stream members, density and velocity information over a larger region near the stream, and better models for the internal structure of the Galactic bar, careful modeling of this stream could lead to tight constraints on the time-dependent properties of the bar.
+With more Ophiuchus stream members, density and velocity information over a larger region near the stream, and better models for the internal structure of the Galactic bar, careful modeling of this stream could lead to tight constraints on the structure and evolution of the bar.
 
@@
 \includegraphics[width=\textwidth]{figures/lyapunov}
-\caption{ Estimates of the Lyapunov time, $t_\lambda=1/\lambda$, for the mean orbits in each potential model (Section~\ref{sec:results1}).
-All time units are given in units of orbital periods.
+\caption{ Estimates of the Lyapunov time, $\lyapt=1/\lyapexp$, for the mean orbits in each potential model (Section~\ref{sec:results1}).
 The orbits are integrated from the mean values of orbital parameters from orbit fitting (Section~\ref{sec:orbitfit}).
-In the static, axisymmetric potential model, an orbit fit to the Ophiuchus stream is regular.
-In all barred potential models (bar1--9), the mean orbit is chaotic with a Lyapunov time that is less than 10 orbital periods.
-}
+In the static, axisymmetric potential model, an orbit fit to the Ophiuchus stream is regular (e.g., the Lyapunov time keeps increasing).
+In all barred potential models (bar1--9), the mean orbit is chaotic with a Lyapunov time between 0.5--1 Gyr.}
 \label{fig:lyapunov}
@@
 
+% ---------------------------------------------------------------------------------
+\clearpage
+\begin{figure}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/lyapunov-hist}
+\caption{ Histograms of estimates of the Lyapunov time, $\lyapt=1/\lyapexp$, for posterior samples from the orbit fitting procedure (Section~\ref{sec:orbitfit}).
+Vertical lines show the Lyapunov time value of the mean orbit in that potential model.
+Orbits are typically more chaotic (have shorter Lyapunov times) for higher pattern speeds.
+There is little dependence on bar angle.}
+\label{fig:lyapunov-hist}
+\end{center}
+\end{figure}
+
+% ---------------------------------------------------------------------------------
+\clearpage
+\begin{figure}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/mockstream0}
+\caption{ Sky position, distance, and line-of-sight velocity in Heliocentric, Galactic coordinates for star particles (grey points) from mock streams generated on the mean orbits in each potential.
+Vertical, dashed lines show the approximate extent of the densest part of the stream visible in main-sequence stars \citep[the segment originally detected in ][]{bernard14}.}
+\label{fig:mockstream0}
+\end{center}
+\end{figure}
+
+\clearpage
+\begin{figure}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/mockstream1}
+\caption{ Same as Figure~\ref{fig:mockstream0} for the other five barred potentials. }
+\label{fig:mockstream1}
+\end{center}
+\end{figure}
+
+% ---------------------------------------------------------------------------------
+\clearpage
+\begin{figure}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/mockstream-xyz}
+\caption{ Star particles (grey points) from mock streams generated on the mean orbits in each potential model shown in projections of Galatocentric, Cartesian coordinates.
+The position of the Sun is shown as the symbol at $(x,z)=(-8,0)~{\rm kpc}$.
+The corners of the sky position and distance plots of Figures~\ref{fig:mockstream0}--\ref{fig:mockstream1} are shown transformed to these coordinates as dark markers near $(x,z)=(-4,3)~{\rm kpc}$, $(x,z)=(0,4)~{\rm kpc}$.
+This demonstrates that the stream is nearly aligned our viewing angle.
+}
+\label{fig:mockstreamxyz}
+\end{center}
+\end{figure}
+
 \end{document}

### 2015-12-23 ms.tex
great simplification of many words and new figures

@@
 
-\title{Spending too much time at the Galactic bar: Chaotic evolution of the Ophiuchus stream}
+\title{Spending too much time at the Galactic bar: Chaotic stream-fanning and the Ophiuchus stream}
 \author{Adrian M. Price-Whelan\altaffilmark{\colum,\adrn},
@@
 % Context
-The Ophiuchus stream was recently discovered as an over-density of main sequence stars in Pan-STARRS (PS1) photometry and was later detected in BHB stars using line-of-sight velocities: XX of the YY BHB stars around the sky position of the stream appear kinematically distinct from the background halo population.
-A small number of stars (XX) near the ends of the stream have similarly distinct velocities but with much larger velocity and spatial dispersions compared to the short segment of the stream first identified in the PS1 data.
-In recent work, we have shown that streams formed on chaotic orbits may display similar kinematic signatures---short segments of recently-disrupted stellar debris flanked by high-dispersion, low-density ``stream-fans'' at the ends of the stream.
-The stream's proximity to the triaxial, time-dependent potential of the Galactic bar and the peculiar kinematics and morphology of the stream suggest that the stream may have formed on a chaotic orbit.
+The Ophiuchus stream was recently discovered as an over-density of main sequence stars in Pan-STARRS (PS1) photometry and is peculiar in it's short apparent length given the age of its stellar population.
+The stream was later detected in BHB stars by identifying members using line-of-sight velocities: Twenty Blue Horizontal Branch (BHB) stars in the vicinity of the stream are kinematically distinct from the background halo stars and have distances consistent with the previously detected segment from PS1.
+Four of these BHB stars near the end of the stream have similarly distinct velocities but have much larger velocity and spatial dispersions compared to the PS1 stream segment.
+In recent work, we have shown that streams formed on chaotic orbits may display similar peculiar morphologies: short segments of coherent stellar debris flanked by high-dispersion, low-density ``stream-fans'' at the ends of the stream.
 % Aims
-Here we study whether the Galactic bar can lead to chaotic orbits near (in phase-space) to the Ophiuchus stream. 
+The stream's proximity to the triaxial, time-dependent potential of the Galactic bar and the peculiar kinematics and morphology of the stream suggest that it may have formed on a chaotic orbit.
+Here we study whether the presence of a triaxial, time-dependent bar can lead to chaotic orbits near (in phase-space) to the Ophiuchus stream. 
 % Methods
-We construct a potential model for the Milky Way bar based on recent measurements of the density of XXX stars.
+[We fit orbits to the debris in one static and several barred potential models ...]
 We then assess the level of chaos for orbits like that of the Ophiuchus stream and generate mock streams along these orbits that qualitatively reproduce the observed kinematics of the stream.
 % Results
-We find that the apparent shortness and nearby low-density, high-velocity-dispersion ...
+With the presence of a bar, many orbits near ... to that of  are chaotic ...
+We find that the apparent shortness and nearby low-density, high-velocity-dispersion can be qualitatively explained if the stream formed along a strongly chaotic orbit.
 % Conclusions
@@
 \end{align}
-\citep{8kpref??, schonrich10, bovy12}, (2) integrate the orbit forward and backward by $t_f$ and $t_b$, respectively, in the potential $\Phi$, (3) transform all orbit points (time-steps) back to observed coordinates, and (4) define interpolating functions for each coordinate as a function of stream longitude, $\phi_1$, using cubic splines---e.g., functions $\widetilde{\phi}_{2}(\phi_1)$, $\widetilde{\DM}(\phi_1)$, $\widetilde{\mu_l}(\phi_1)$, $\widetilde{\mu_b}(\phi_1)$, $\widetilde{v_r}(\phi_1)$.
+\citep{8kpc 	ref??, schonrich10, bovy12}, (2) integrate the orbit forward and backward by $t_f$ and $t_b$, respectively, in the potential $\Phi$, (3) transform all orbit points (time-steps) back to observed coordinates, and (4) define interpolating functions for each coordinate as a function of stream longitude, $\phi_1$, using cubic splines---e.g., functions $\widetilde{\phi}_{2}(\phi_1)$, $\widetilde{\DM}(\phi_1)$, $\widetilde{\mu_l}(\phi_1)$, $\widetilde{\mu_b}(\phi_1)$, $\widetilde{v_r}(\phi_1)$.
 These functions let us compute the predicted values of each of these coordinates at the longitudes of each observed star, $\phi_{1,i}$.
@@
 The position of the Lagrange points and the scale of the dispersion in position and velocity are set by the progenitor mass, $m$.
-\todo{Slightly more detail here about Fardal's parametrization} The star particles are drawn from Gaussians centered on the Lagrange points (in position) and the progenitor (in velocity) and the full parametrization of the release distribution is given in \cite{fardal14}.
+The star particles are drawn from Gaussians centered on the Lagrange points (in position) and the progenitor (in velocity) and the full parametrization of the release distribution is given in \cite{fardal14}.
 This method has been shown to reproduce the morphologies of $N$-body simulations of stellar streams, but requires far less computing time because it relies only on integrating test-particle orbits.
+\todo{Does this need more detail  about Fardal's method?}
 
@@
 
-For the posterior samples in each potential, we also compute the maximum Lyapunov exponent (MLE, $\lyapexp$) and corresponding Lyapunov \emph{time} ($\lyapt = 1/\lyapexp$) to assess whether the orbit is chaotic.
+For the posterior samples in each potential, we also compute the maximum Lyapunov exponent (MLE, $\lyapexp$) and corresponding Lyapunov \emph{time} ($\lyapt = 1/\lyapexp$) to assess whether each orbit is chaotic.
 For strongly chaotic orbits, the Lyapunov time is still an appropriate indicator of chaos and of the timescale over which chaos is important for tidal debris \citep{apw15-chaos}.
-Figure~\ref{fig:lyapunov} shows numerical estimates of the Lyapunov time for the mean orbits in each potential model: In the static potential the orbit appears regular (black, diagonal line), but in all barred potentials the MLE estimates converge to finite values in the range $\approx$2--8 orbital periods.
-Figure~\ref{XX} shows distributions of Lyapunov times for orbits drawn from the posterior distributions from orbit fitting in each potential model.
-\todo{What else to say?} \todo{What fraction of the orbits sampled from the posterior are chaotic and what are distribution of Lyapunov times?
-}
-\todo{PLOT MEAN ORBIT VALUE AS VERTICAL LINE}
+Figure~\ref{fig:lyapunov-hist} shows distributions of Lyapunov times for orbits drawn from the posterior distributions from orbit fitting in each potential model.
+All orbits in the static potential have Lyapunov times $\lyapt > 20~{\rm Gyr}$ and we consider them to be regular (no panel is shown for these orbits).
+All orbits sampled from each barred potential are strongly chaotic with Lyapunov times that range from $\lyapt \approx 400--1100~{\rm Myr}$.
+It is clear from these panels that the orbits around Ophiuchus are generally more strongly chaotic (have lower Lyapunov times) for larger pattern speeds.
 
-The orbits in the barred potentials are all strongly chaotic.
-We have tried computing the frequency diffusion rate for these orbits but have found that, over consecutive integration windows, the frequency recovery fails or is unreliable because the frequency spectrum changes dramatically over timescales of $\approx$10--20 orbital periods.
+The orbits sampled from the orbit-fit posteriors in the barred potentials are all strongly chaotic.
+We have tried computing the frequency diffusion rate for these orbits as an independent check of their chaotic timescale but have found that, over consecutive integration windows, the frequency recovery fails or is unreliable because the frequency spectrum changes dramatically over timescales of $\approx$10--20 orbital periods.
 
@@
 	name & $\phi_2$ [deg] & $d$ [kpc] & $\mu_l$ [mas yr$^{-1}$] & $\mu_b$ [mas yr$^{-1}$] & $v_r$ [km s$^{-1}$]\\\midrule
-	static & $-0.03\pm0.05$ & $8.35\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.1\pm0.9$\\
-	bar1--9 & $-0.03\pm0.05$ & $8.40\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.0\pm1.0$\\
+	static & $-0.03\pm0.05$ & $8.3\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.1\pm0.9$\\
+	bar1--9 & $-0.03\pm0.05$ & $8.35\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.0\pm1.0$\\
 	\bottomrule
@@
 	\item The density of the mock stream in the static potential decreases slowly rather than abruptly as is observed.
-This is more clearly seen in terms of the \todo{linear density of the stream}, shown in Figure~\ref{fig:lineardensity}.
+This is more clearly seen in terms of the \todo{linear density of the stream??}, shown in Figure~\ref{fig:lineardensity}.
 ....
-\todo{Figure showing some brightness density vs longitude?}
+\todo{Figure showing some brightness density vs longitude?
+Waiting on Branimir!}
 	\item For the barred potentials, the stream morphology is very sensitive to the properties of the bar (especially the pattern speed) and to the initial conditions of the orbit.
@@
 The stream star particles at present day---e.g.
-the ``model'' points---are then a sampling of the stream model (phase-space) density field given a uniform stripping time distribution.
+the ``model'' points---are therefore a sampling of the stream model (phase-space) density field given a uniform stripping time distribution.
 We can then ask, given the data---a noisy sampling of the true density field---what stripping time distribution would reproduce the observed density field (e.g., the positions and velocities of the BHB stars).
@@
 (see Section~\ref{sec:orbitfit} for variable definitions).
+For the static potential all of the BHB stars associated with the stream must have been released very recently, so we expect that the inferred stripping time distribution for the mock stream in the static potential should be heavily skewed towards recent times.
+Indeed, the top left panel in Figure~\ref{fig:massloss} shows the stripping time distribution for the mock stream in the static potential: all stars would have to have been disrupted in the last $\lesssim 600~{\rm Myr}$ in order to produce the short segment of the stream observed.
+This is consistent with the findings and $N$-body simulations in S15a.
 
-If we assume that the data are a noisy sampling of the true (phase-space) density field,
-
-Each model point We do this by computing the \todo{what is it called?} for each model point, $k$:  \citep{hogg10}.
-
-As a more quantitative / alternative exploration of the density ....... we consider the mass-loss histories...
+\todo{This is wrong -- maybe axe? maybe bar3 is showing resonant stuff that the fardal, sanders paper talks about...}
+For the mock streams in the barred potentials, the stripping time distributions are generally multi-modal and in some cases extend several Gigayears into the past, as we would expect for a disrupting globular cluster.
+In particular, the panel labeled bar3 shows a case in which there could be many ``mass-loss'' events over the history of the cluster's orbit, but the final debris distribution would still remain short.
+This is because the star particles released earlier will diffuse away and therefore won't get penalized by the weighting scheme described above.
 
@@
 Instead, these results illustrate that chaotic evolution of tidal debris can plausibly explain the peculiar features of the stream.
-If the cluster progenitor was on a regular orbit, it would have to have disrupted entirely within the last $\approx$300 Myr in order to explain the shortness and density profile.
+If the cluster progenitor was on a regular orbit, it would have to have disrupted entirely within the last $\approx$300--600 Myr in order to explain the shortness and density profile.
 In addition, the four most recently identified BHB stars with similar distances and line-of-sight velocities would have to be (highly unlikely) chance alignments of halo stars.
@@
 We consider the second scenario to be more plausible: our understanding of the formation and evolution of the Ophiuchus stream is that the progenitor object has been orbiting and steadily losing stars over the last several Gigayears, but only the stars stripped from the most recent few pericentric passages remain coherent enough to be detected as a stream-like over-density in the PS1 data.
+%[Gaia, Don't do distance cut that Brani did, chemistry]
 
-- with better data, perhaps we can constrain the bar?
+It is still too early to say for sure that chaotic stream-fanning is occurring for the Ophiuchus stream; deeper follow-up imaging and spectroscopy over a larger area region around the stream will be needed to test the predictions of this work.
+With more information about the density distribution of stars in the stream and better proper motion measurements we would be able to (1) help distinguish models for the Milky Way bar independent from current methods and (2) begin to model the survivability of globular clusters orbiting in the central Milky Way.
 
-[bar not well understood but maybe this stream can help distinguish models with more data -- cite discrepancy in measruements (antoja14, wang12, wang13)]
+\todo{Where to put this?}
 
-[Gaia, Don't do distance cut that Brani did, chemistry]
+A typical inner MW globular cluster with a mass $m_{\rm cl} \approx 10^4~\msun$ has a half-mass radius $r_h \approx 2~{\rm pc}$.
+We can ask, given an assumed mass density profile for the inner galaxy, what Galactocentric pericentric radius the Ophiuchus progenitor would have to reach such that the tidal radius of the cluster is equal to the half-mass radius.
 
-[explored in future work: what are implications of the existence of this structure for total population]
+Let's assume the density profile of the inner $\approx 1~{\rm kpc}$ of the MW is a Jaffe spheroid so that
+\begin{align}
+	\rho(r) &= \frac{\rho_0}{(r/a)^2 \, (1 + r/a)^2}\\
+	M(<r) &= 4\pi \, \rho_0 \, a^3 \, \frac{r/a}{1 + r/a}
+\end{align}
+with scale radius $a \approx 0.5~{\rm kpc}$ and mass within the scale radius $M(<a) \approx 10^{10}~\msun$ ($\rho_0 \approx \frac{10^{10}~\msun}{2\pi a^3} \approx 10^{10}~\msun~{\rm kpc}^{-3}$).
+The tidal radius is approximately
+\begin{align}
+	r_{\rm tide} \approx \left(\frac{m_{\rm cl}}{3M(<r)}\right)^{1/3} \, r
+\end{align}
+so that
 
-What is a successful model?
-- Sharp density drop
-	- BHB density falls off sooner? Is that just small-number statistics?
-	- Density profile along the stream / how sharp is the drop?
-	- Density contrast for BHB stars?
-- Increased positional dispersion
-	- The 4 new stars
-- Velocity dispersion of the 4 new stars
-	- velocity dispersion ~ 40 times internal velocity dispersion of stream stars
-- What can we say about the connection between chaotic timescale and the length of the stream? Is it the onset time for chaos, or are all orbits strongly chaotic?
-	- for a given orbit and model, paint the particles to say what the mass loss rate would have to be to match the obs.
-- "It's stupid to make it more complicated" -- Kathryn Johnston
-- from painting particles, Length of stream --> dispersion
+then
+\begin{align}
+	2~{\rm pc} &= \left(\frac{m_{\rm cl}}{3M(<r)}\right)^{1/3} \, r\\
+\end{align}
 
-\todo{Still to early to say for sure, but a plausible explanation...}
+if we set
+\begin{align}
+	r_{\rm tide} = r_h = 2~{\rm pc} 
+\end{align}
+
+\todo{Predictions and tests of this model!}
+
+\subsection{Future work: Modeling the Galactic bar}
+
+The current kinematic data for the Ophiuchus BHB stars suggests that the stream is sensitive to the potential of the bar, but with better measurements of the velocities and a larger sample of member stars we would be able to constrain the bar model parameters.
+Current measurements of the pattern speed, angle, and structure of the Milky Way's bulge and bar are largely discrepant \citep[e.g.,][]{wang12, wang13, antoja14, wegg??
+is wegg appropriate here}.
+Most of the methods that infer these quantities rely on modeling the density or kinematics of stars at low Galactic latitudes and must therefore handle challenges with completeness and dust extinction.
+The Ophiuchus stream offers a unique opportunity to independently measure these quantities by modeling the density and kinematics of stars associated with the stream.
+
+\subsection{Future work: The orbits and survivability of inner Milky Way globular clusters}
+
+If the Ophiuchus stream formed from a globular cluster on a chaotic orbit and we happen to be witnessing its final demise, what does this imply about the population of clusters that have already been fully disrupted?
+For a given model for the Galactic bar, we can study the orbit structure in the inner Milky Way and, in particular, study the fraction of strongly chaotic orbits that would lead to chaotic fanning or fast dispersal of tidal debris \citep{recent work by zotos or those greek peeps}.
+The existence of strongly chaotic orbits in this region would limit the expected number of cold stellar streams in the inner Galaxy.
+Comparing the observed numbers with expectations ...
+\todo{what to say exactly -- KVJ?}
 
@@
 \includegraphics[width=\textwidth]{figures/orbit-yz}
-\caption{  TODO }
+\caption{  Projections of orbits integrated from the mean orbital parameters estimated from the orbit fitting posterior distributions in each potential model.
+Orbits are integrated for 6 Gyr and shown in Galactocentric, Cartesian coordinates.
+Even though the mean orbital parameters have nearly identical values (e.g., the initial conditions are nearly identical in Heliocentric coordinates), the orbits in each potential model are  different in appearance.
+}
 \label{fig:orbits-yz}
@@
 % ---------------------------------------------------------------------------------
-\clearpage
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/lyapunov}
-\caption{ Estimates of the Lyapunov time, $\lyapt=1/\lyapexp$, for the mean orbits in each potential model (Section~\ref{sec:results1}).
+%\clearpage
+%\begin{figure}[p]
+%\begin{center}
+%\includegraphics[width=\textwidth]{figures/lyapunov}
+%\caption{ Estimates of the Lyapunov time, $\lyapt=1/\lyapexp$, for the mean orbits in each potential model (Section~\ref{sec:results1}).
 The orbits are integrated from the mean values of orbital parameters from orbit fitting (Section~\ref{sec:orbitfit}).
@@
 In all barred potential models (bar1--9), the mean orbit is chaotic with a Lyapunov time between 0.5--1 Gyr.}
-\label{fig:lyapunov}
-\end{center}
-\end{figure}
+%\label{fig:lyapunov}
+%\end{center}
+%\end{figure}
 
@@
 
+\clearpage
+\begin{figure}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/surface-density}
+\caption{ Surface density of mock star particles ... \todo{This is with the same distance cut as used for the BHB stars -- is that what we want to use?}}	
+\label{fig:surfacedensity}
+\end{center}
+\end{figure}
+
 % ---------------------------------------------------------------------------------
@@
 
+% ---------------------------------------------------------------------------------
+%\clearpage
+%\begin{figure}[p]
+%\begin{center}
+%\includegraphics[width=\textwidth]{figures/mass-loss}
+%\caption{ Computed stripping time distributions needed for each mock stream model in order to reproduce the observed distribution of BHB stars. }
+%\label{fig:massloss}
+%\end{center}
+%\end{figure}
+
 \end{document}

### 2015-12-25 ms.tex
before making two-column

@@
 
-\subsection{Fitting orbits to the Ophiuchus stream}\label{sec:orbitfit}
+\subsection{Fitting orbits to the Ophiuchus stream}
 
 In each of the potentials described above, we fit orbits to the measured kinematics of BHB stars that are high-likelihood members of the Ophiuchus stream \citep{sesar15a, sesar15b}.
-Our goal is to infer the posterior probability distributions over orbital initial conditions, $\bs{w}_0=(l, b, \DM, \mu_l, \mu_b, v_r)_0$, given a potential, $\Phi$, and kinematic data for each $i$ stream star, $\bs{x}_i=(l, b, \DM, \mu_l, \mu_b, v_r)_i$.
-In this notation, $(l, b)$ are Galactic coordinates, $\DM$ is the distance modulus, $(\mu_l, \mu_b)$ are proper motions in the Galactic frame, and $v_r$ is the radial velocity.
-We assume that the sky coordinates for each star are known perfectly well (have zero uncertainty) and transform the data to a rotated, heliocentric coordinate system that is aligned with the stream and centered on the median sky position of the BHB stars in the densest part of the stream \cite[all BHB stars except the `fanned' stars: cand15, cand26, cand49, cand54 from][]{sesar15b}.
-We represent the longitude and latitude in these coordinates as $(\phi_1, \phi_2)$ and the rotation matrix to transform from Galactic to these coordinates is given in Appendix~\ref{sec:rotationmatrix}.
-We treat the stream longitude, $\phi_1$, as the perfectly-known, independent variable so that all other coordinates can be expressed as functions of this longitude (e.g., $\phi_2(\phi_1)$, ${\rm DM}(\phi_1)$, etc.).
-This methodology is similar to that used in \cite{koposov10} and \cite{sesar15a}.
-
-\subsubsection{Likelihood}
-
-\todo{Move a lot of this to appendix?}
-
-We include three nuisance parameters in our likelihood to account for the internal dispersion of the stream: in observed coordinates, these are the on-sky positional dispersion, $s_{\phi_2}$, a distance (modulus) dispersion, $s_{\DM}$, and a radial velocity dispersion, $s_{v_r}$ (the proper motion uncertainties are sufficiently large that we can't resolve the velocity dispersion in these coordinates).\footnote{We assume that the dispersion in these coordinates is constant over the observed (short) section of the stream.
-This may be a bad assumption.} We add two additional nuisance parameters for controlling the amount of time to integrate forwards, $t_f$, and backwards, $t_b$, from the given initial conditions, which ultimately controls the length of the section of orbit that is compared to the stream star data.
-For brevity in the equations below, we define $\bs{s} = (s_{\phi_2}, s_{\DM}, s_{v_r})$ and $\bs{\theta} = (\bs{w}_0, \Phi, t_b, t_f)$.
-
-For a given set of initial conditions ($\bs{w}_0$), we compute a model orbit as follows: (1) transform the initial conditions to Galactocentric coordinates, where we take the position and velocity of the Sun to be
-\begin{align}
-	(x,y,z)_\odot &= (-8,0,0)~{\rm kpc}\\
-	(v_x,v_y,v_z)_\odot &= (-11.1, 220+24, 7.25)~{\rm km}~{\rm s}^{-1}
-\end{align}
-\citep{8kpc 	ref??, schonrich10, bovy12}, (2) integrate the orbit forward and backward by $t_f$ and $t_b$, respectively, in the potential $\Phi$, (3) transform all orbit points (time-steps) back to observed coordinates, and (4) define interpolating functions for each coordinate as a function of stream longitude, $\phi_1$, using cubic splines---e.g., functions $\widetilde{\phi}_{2}(\phi_1)$, $\widetilde{\DM}(\phi_1)$, $\widetilde{\mu_l}(\phi_1)$, $\widetilde{\mu_b}(\phi_1)$, $\widetilde{v_r}(\phi_1)$.
-These functions let us compute the predicted values of each of these coordinates at the longitudes of each observed star, $\phi_{1,i}$.
-
-We assume that each observed kinematic component is independent so that the likelihood of the data for a given star, $\bs{x}_i$, with uncertainties, $\bs{\sigma}_i$, is given by the product over the likelihoods for each dimension of the data:
-\begin{multline}
-	p(\bs{x}_i \given \bs{\sigma}_i, \bs{s}, \bs{\theta}) = p(\phi_{2,i} \given \phi_{1,i}, s_{\phi_2},\bs{\theta}) \, p(\DM_i \given \phi_{1,i}, \sigma_{\DM,i}, s_\DM, \bs{\theta})\\ 
-	\times p(\mu_{l,i} \given \phi_{1,i}, \sigma_{\mu_{l},i}, \bs{\theta}) \, p(\mu_{b,i} \given \phi_{1,i}, \sigma_{\mu_{b},i}, \bs{\theta}) \, p(v_{r,i} \given \phi_{1,i}, \sigma_{v_r,i}, s_{v_r}, \bs{\theta}).
-\end{multline}
-The uncertainties in these observed coordinate components are assumed to be normally distributed away from the model values: using the notation
-\begin{align}
-	\norm(x \given \mu, \sigma^2) &= \frac{1}{\sqrt{2\pi \sigma^2}} \, \exp\left(-\frac{(x-\mu)^2}{2\sigma^2}\right)
-\end{align}
-the likelihoods are
-\begin{align}	
-	p(\phi_{2,i} \given \phi_{1,i}, s_{\phi_2},\bs{\theta}) &= \norm(\phi_{2,i} \given \widetilde{\phi}_{2}(\phi_{1,i}), s^2_{\phi_2})\\
-	p(\DM_i \given \phi_{1,i}, \sigma_{\DM,i}, s_\DM, \bs{\theta}) &= \norm(\DM_i \given \widetilde{\DM}(\phi_{1,i}), s^2_{\DM} + \sigma^2_{\DM,i})\\
-	p(\mu_{l,i} \given \phi_{1,i}, \sigma_{\mu_{l},i}, \bs{\theta}) &= \norm(\mu_{l,i} \given \widetilde{\mu_{l}}(\phi_{1,i}), \sigma^2_{\mu_{l,i}})\\
-	p(\mu_{b,i} \given \phi_{1,i}, \sigma_{\mu_{b},i}, \bs{\theta}) &= \norm(\mu_{b,i} \given \widetilde{\mu_{b}}(\phi_{1,i}), \sigma^2_{\mu_{b,i}})\\
-	p(v_{r,i} \given \phi_{1,i}, \sigma_{v_r,i}, s_{v_r}, \bs{\theta}) &= \norm(v_{r,i} \given \widetilde{v_r}(\phi_{1,i}), s^2_{v_r} + \sigma^2_{v_r,i}).
-\end{align}
-We assume the data from each star is independent and identically distributed (i.i.d.) so that the full likelihood is the product over the likelihoods for all $N$ stars:
-\begin{equation}
-	 p(\{\bs{x}_i\} \given \{\bs{\sigma}_i\}, \bs{s}, \bs{\theta}) = \prod_i^N p(\bs{x}_i \given \bs{\sigma}_i, \bs{s}, \bs{\theta}).\label{eq:likelihood}
-\end{equation}
-
-\subsubsection{Priors}
-
-For the intrinsic dispersion parameters, we use logarithmic (scale-invariant) priors such that $p(s) \propto s^{-1}$.
-For the integration time parameters, we use uniform priors, $\mathcal{U}(a,b)$ (over the range $a$--$b$),
-\begin{align}
-	p(t_f) &= \mathcal{U}(1,100)~{\rm Myr}\label{eq:prior1}\\
-	p(t_b) &= \mathcal{U}(-100,-1)~{\rm Myr}.
-\end{align}
-Note that present-day is $t=0$.
-For computational efficiency, we place strong priors on the minimum and maximum longitudes of the model points, $(\phi_{1,{\rm min}},\phi_{1,{\rm max}})$ so that the model orbit does not integrate for longer than necessary.
-In particular, we set
-\begin{align}
-	p(\phi_{1,{\rm min}} \given \bs{\theta}) &= \norm(\phi_{1,{\rm min}} \given \min(\phi_{1,i}), s^2_{\phi_2})\\
-	p(\phi_{1,{\rm max}} \given \bs{\theta}) &= \norm(\phi_{1,{\rm max}} \given \max(\phi_{1,i}), s^2_{\phi_2}).
-\end{align}
-For the orbital initial condition components, we use uniform priors in each cartesian position component over the range $(-200,200)~{\rm kpc}$.
-For velocity, we use a Gaussian prior on the magnitude of the total velocity, $v$, with a dispersion of $150~\kms$,
-\begin{equation}
-	\norm(v \given 0, (150~\kms)^2) \label{eq:prior2}
-\end{equation}
-We keep the potential, $\Phi$, fixed. In total, this model has 10 parameters (5 phase-space coordinates, 5 nuisance parameters).
-
-\subsubsection{Sampling from the posterior distributions}
-
-The full expression for the posterior probability, $p(\bs{s}, \bs{w}_0, t_b, t_f \given \{\bs{x}_i\}, \{\bs{\sigma}_i\}, \Phi)$, is the joint likelihood (Equation~\ref{eq:likelihood}) multiplied by all priors described above (Equations~\ref{eq:prior1}--\ref{eq:prior2}).
-We use an ensemble Markov Chain Monte Carlo (MCMC) algorithm \citep{goodman10} implemented in \python\ (\package{emcee}) to generate samples from this posterior distribution \citep{foremanmackey13}.
+The details of this procedure and a definition of the likelihood function we use are presented in Appendix~\ref{sec:orbitfit}.
+We use an ensemble Markov Chain Monte Carlo (MCMC) algorithm \citep{goodman10} implemented in \python\ (\package{emcee}) to generate samples from the posterior distribution over the parameters in our orbit-fitting model \citep{foremanmackey13}.
 The algorithm uses an ensemble of individual ``walkers'' to adapt to the geometry of the parameter-space being explored.
@@
 
-We generate mock streams along each of the mean orbits from the orbit fitting procedure (defined above).
+We randomly sample 256 orbits from the orbital parameter posterior distributions and generate mock streams along each orbit.
 We use the method outlined above (Section \ref{sec:mocks}) to generate the streams and set the free parameters as follows: (1) we evolve the progenitors for 6 Gyr along each orbit, (2) we release star particles every 0.5 Myr (uniformly in time) to densely sample the final density distribution, (3) we set the progenitor mass to $m=10^4~\msun$, and (4) we set the disruption time of the progenitor equal to the last time at which a pericenter and disk crossing coincide (in each case, this is at $t \approx -200~{\rm Myr}$).
+After the disruption time, we continue releasing the star particles uniformly in time rather than releasing a ``burst'' of particles at once.
+We therefore expect that the morphology and value of the density of the most recently disrupted debris will not exactly match the data.
 
-Figures~\ref{fig:mockstream0}--\ref{fig:mockstream1} show the final particle positions and line-of-sight velocities in heliocentric coordinates for these mock streams (grey points).
+For each generated mock stream, we compute the likelihood of the data given the star particles by estimating the phase-space model density using a kernel density estimate with a Gaussian kernel \citep[see, e.g.,][]{bonaca14}.
+For each $i$ data point and each $k$ model point with coordinates $\bs{y}_k$, we compute the likelihood by converting the model point position and velocity into Heliocentric coordinates and evaluate
+\begin{align}
+	p(\bs{x}_i \given \bs{y}_k, \bs{\sigma}_i, \bs{h}) = \mathcal{N}(\bs{x}_i \given \bs{y}_k, \bs{\sigma}_i^2 +\bs{h}^2)
+\end{align}
+where $\mathcal{N}(x\given \mu,\sigma^2)$ is the normal distribution with mean $\mu$ and variance $\sigma^2$ and $\bs{h}$ represents the diagonal of the bandwidth matrix, $\bs{\rm H}$, used for the density estimate, $\bs{h} = {\rm diag}(\bs{\rm H})$.
+We fix the bandwidth parameters as follows
+\begin{equation}
+	\bs{h} = \left(
+	\begin{array}{c}
+	0.2~{\rm deg}\\
+	0.2~{\rm deg}\\
+	0.2~{\rm kpc}\\
+	0.2~{\rm mas}~{\rm yr}^{-1}\\
+	0.2~{\rm mas}~{\rm yr}^{-1}\\
+	0.2~\kms
+	\end{array}
+	\right)
+\end{equation} %(max$\mathcal{L}$)
+for sky position in Galactic coordinates ($l$, $b$), distance, proper motions ($\mu_l$, $\mu_b$), and radial velocity. The full likelihood for all $N$ data points given $K$ model points is 
+\begin{equation}
+	\mathcal{L} = \prod_i^N \frac{1}{K} \sum_k^K p(\bs{x}_i \given \bs{y}_k, \bs{\sigma}_i, \bs{h}).
+\end{equation}
+
+Figures~\ref{fig:mockstream0}--\ref{fig:mockstream1} show the final particle positions and line-of-sight velocities in heliocentric coordinates for the maximum likelihood mock streams (grey points) in each potential.
 Vertical, dashed lines show the approximate extent of the part of the stream visible in main-sequence stars (excluding the BHB stars from S15b).
-We have applied the same distance-modulus cut to the points in these figures as used to select the BHB stars identified as stream members by S15b (see Section~2 in S15b).
-There are a few interesting features to note:
+There are a few interesting features to note from these panels:
 \begin{enumerate}
-	\item Even in the static potential (Figure~\ref{fig:mockstream0}, leftmost column), there is an increased dispersion in sky position for the stream stars towards higher Galactic longitudes, $l$.
-This is a projection effect: The portion of the stream at larger $l$ is closer and almost pointing towards the Sun, as demonstrated in Figure~\ref{fig:mockstreamxyz}.
-Figure~\ref{fig:mockstreamxyz} shows $x$-$z$ projections of the star particles in Galactocentric Cartesian coordinates (grey) along with the position of the Sun (symbol at $(x,z)=(-8,0)~{\rm kpc}$) and the corners of the Heliocentric, sky-position plots of Figures~\ref{fig:mockstream0}--\ref{fig:mockstream1}.
+	\item Even in the static potential (Figure~\ref{fig:mockstream0}, leftmost column), there is a slight decrease in the density for the model stream towards higher Galactic longitudes, $l$.
+This is a projection effect: The portion of the stream at larger $l$ is closer points almost directly towards the Sun so that the debris covers a larger area.
+This is more clearly seen in Figure~\ref{fig:mockstreamxyz}, which shows $x$-$z$ projections of the star particles in Galactocentric Cartesian coordinates (grey) along with the position of the Sun (symbol at $(x,z)=(-8,0)~{\rm kpc}$) and the ``window'' of the Heliocentric, sky-position plots of Figures~\ref{fig:mockstream0}--\ref{fig:mockstream1} (shown as blue lines).
 	\item The density of the mock stream in the static potential decreases slowly rather than abruptly as is observed.
-This is more clearly seen in terms of the \todo{linear density of the stream??}, shown in Figure~\ref{fig:lineardensity}.
+This is more clearly seen in terms of the \todo{linear density along the stream, shown in Figure~\ref{fig:lineardensity}}.
 ....
-\todo{Figure showing some brightness density vs longitude?
-Waiting on Branimir!}
+\todo{Waiting on Branimir!}
 	\item For the barred potentials, the stream morphology is very sensitive to the properties of the bar (especially the pattern speed) and to the initial conditions of the orbit.
@@
 
-To generate each mock stream, the star particles are released uniformly in time.
-The stream star particles at present day---e.g.
-the ``model'' points---are therefore a sampling of the stream model (phase-space) density field given a uniform stripping time distribution.
-We can then ask, given the data---a noisy sampling of the true density field---what stripping time distribution would reproduce the observed density field (e.g., the positions and velocities of the BHB stars).
-This corresponds to computing a weight for each $k$ model star particle with phase-space position $\bs{z}_k$ and release time, $t_{s,k}$:
-\begin{align}
-	w_k = \frac{1}{N} \, \sum^N_i \, p(\bs{z}_k \given \bs{x}_i, \bs{\sigma}_i, \bs{s})
-\end{align}
-(see Section~\ref{sec:orbitfit} for variable definitions).
-For the static potential all of the BHB stars associated with the stream must have been released very recently, so we expect that the inferred stripping time distribution for the mock stream in the static potential should be heavily skewed towards recent times.
-Indeed, the top left panel in Figure~\ref{fig:massloss} shows the stripping time distribution for the mock stream in the static potential: all stars would have to have been disrupted in the last $\lesssim 600~{\rm Myr}$ in order to produce the short segment of the stream observed.
-This is consistent with the findings and $N$-body simulations in S15a.
-
-\todo{This is wrong -- maybe axe? maybe bar3 is showing resonant stuff that the fardal, sanders paper talks about...}
-For the mock streams in the barred potentials, the stripping time distributions are generally multi-modal and in some cases extend several Gigayears into the past, as we would expect for a disrupting globular cluster.
-In particular, the panel labeled bar3 shows a case in which there could be many ``mass-loss'' events over the history of the cluster's orbit, but the final debris distribution would still remain short.
-This is because the star particles released earlier will diffuse away and therefore won't get penalized by the weighting scheme described above.
+%We have applied the same distance-modulus cut to the points in these figures as used to select the BHB stars identified as stream members by S15b (see Section~2 in S15b).
 
@@
 It is still too early to say for sure that chaotic stream-fanning is occurring for the Ophiuchus stream; deeper follow-up imaging and spectroscopy over a larger area region around the stream will be needed to test the predictions of this work.
-With more information about the density distribution of stars in the stream and better proper motion measurements we would be able to (1) help distinguish models for the Milky Way bar independent from current methods and (2) begin to model the survivability of globular clusters orbiting in the central Milky Way.
+With more information about the density distribution of stars in the stream and better proper motion measurements we would be able to (1) help distinguish models for the Milky Way bar independent from current methods and (2) begin to model the survivability of globular clusters orbiting in the central Milky Way \citep[e.g.,][]{gnedin97}.
 
-\todo{Where to put this?}
-
-A typical inner MW globular cluster with a mass $m_{\rm cl} \approx 10^4~\msun$ has a half-mass radius $r_h \approx 2~{\rm pc}$.
+%\todo{Where to put this?}
+%
+%also, cite gnedin97
+%
+%A typical inner MW globular cluster with a mass $m_{\rm cl} \approx 10^4~\msun$ has a half-mass radius $r_h \approx 2~{\rm pc}$.
 We can ask, given an assumed mass density profile for the inner galaxy, what Galactocentric pericentric radius the Ophiuchus progenitor would have to reach such that the tidal radius of the cluster is equal to the half-mass radius.
-
-Let's assume the density profile of the inner $\approx 1~{\rm kpc}$ of the MW is a Jaffe spheroid so that
-\begin{align}
-	\rho(r) &= \frac{\rho_0}{(r/a)^2 \, (1 + r/a)^2}\\
-	M(<r) &= 4\pi \, \rho_0 \, a^3 \, \frac{r/a}{1 + r/a}
-\end{align}
-with scale radius $a \approx 0.5~{\rm kpc}$ and mass within the scale radius $M(<a) \approx 10^{10}~\msun$ ($\rho_0 \approx \frac{10^{10}~\msun}{2\pi a^3} \approx 10^{10}~\msun~{\rm kpc}^{-3}$).
+%
+%Let's assume the density profile of the inner $\approx 1~{\rm kpc}$ of the MW is a Jaffe spheroid so that
+%\begin{align}
+%	\rho(r) &= \frac{\rho_0}{(r/a)^2 \, (1 + r/a)^2}\\
+%	M(<r) &= 4\pi \, \rho_0 \, a^3 \, \frac{r/a}{1 + r/a}
+%\end{align}
+%with scale radius $a \approx 0.5~{\rm kpc}$ and mass within the scale radius $M(<a) \approx 10^{10}~\msun$ ($\rho_0 \approx \frac{10^{10}~\msun}{2\pi a^3} \approx 10^{10}~\msun~{\rm kpc}^{-3}$).
 The tidal radius is approximately
-\begin{align}
-	r_{\rm tide} \approx \left(\frac{m_{\rm cl}}{3M(<r)}\right)^{1/3} \, r
-\end{align}
-so that
-
-then
-\begin{align}
-	2~{\rm pc} &= \left(\frac{m_{\rm cl}}{3M(<r)}\right)^{1/3} \, r\\
-\end{align}
-
-if we set
-\begin{align}
-	r_{\rm tide} = r_h = 2~{\rm pc} 
-\end{align}
-
-\todo{Predictions and tests of this model!}
+%\begin{align}
+%	r_{\rm tide} \approx \left(\frac{m_{\rm cl}}{3M(<r)}\right)^{1/3} \, r
+%\end{align}
+%so that 
+%
+%then
+%\begin{align}
+%	2~{\rm pc} &= \left(\frac{m_{\rm cl}}{3M(<r)}\right)^{1/3} \, r\\
+%\end{align}
+%
+%
+%if we set
+%\begin{align}
+%	r_{\rm tide} = r_h = 2~{\rm pc} 
+%\end{align}
 
@@
 Comparing the observed numbers with expectations ...
-\todo{what to say exactly -- KVJ?}
+\todo{KVJ add words?}
 
 \section{Conclusions}\label{sec:conclusions}
+
+\todo{Predictions and tests of this model!}
 
@@
 
+\section{Fitting orbits to stellar streams}\label{sec:orbitfit}
+
+Our goal is to infer the posterior probability distributions over orbital initial conditions, $\bs{w}_0=(l, b, \DM, \mu_l, \mu_b, v_r)_0$, given a potential, $\Phi$, and kinematic data for each $i$ stream star, $\bs{x}_i=(l, b, \DM, \mu_l, \mu_b, v_r)_i$.
+In this notation, $(l, b)$ are Galactic coordinates, $\DM$ is the distance modulus, $(\mu_l, \mu_b)$ are proper motions in the Galactic frame, and $v_r$ is the radial velocity.
+We assume that the sky coordinates for each star are known perfectly well (have zero uncertainty) and transform the data to a rotated, heliocentric coordinate system that is aligned with the stream and centered on the median sky position of the BHB stars in the densest part of the stream \cite[all BHB stars except the `fanned' stars: cand15, cand26, cand49, cand54 from][]{sesar15b}.
+We represent the longitude and latitude in these coordinates as $(\phi_1, \phi_2)$ and the rotation matrix to transform from Galactic to these coordinates is given in Appendix~\ref{sec:rotationmatrix}.
+We treat the stream longitude, $\phi_1$, as the perfectly-known, independent variable so that all other coordinates can be expressed as functions of this longitude (e.g., $\phi_2(\phi_1)$, ${\rm DM}(\phi_1)$, etc.).
+This methodology is similar to that used in \cite{koposov10} and \cite{sesar15a}.
+
+\subsection{Likelihood}
+
+We include three nuisance parameters in our likelihood to account for the internal dispersion of the stream: in observed coordinates, these are the on-sky positional dispersion, $s_{\phi_2}$, a distance (modulus) dispersion, $s_{\DM}$, and a radial velocity dispersion, $s_{v_r}$ (the proper motion uncertainties are sufficiently large that we can't resolve the velocity dispersion in these coordinates).\footnote{We assume that the dispersion in these coordinates is constant over the observed (short) section of the stream.
+This may be a bad assumption.} We add two additional nuisance parameters for controlling the amount of time to integrate forwards, $t_f$, and backwards, $t_b$, from the given initial conditions, which ultimately controls the length of the section of orbit that is compared to the stream star data.
+For brevity in the equations below, we define $\bs{s} = (s_{\phi_2}, s_{\DM}, s_{v_r})$ and $\bs{\theta} = (\bs{w}_0, \Phi, t_b, t_f)$.
+
+For a given set of initial conditions ($\bs{w}_0$), we compute a model orbit as follows: (1) transform the initial conditions to Galactocentric coordinates, where we take the position and velocity of the Sun to be
+\begin{align}
+	(x,y,z)_\odot &= (-8,0,0)~{\rm kpc}\\
+	(v_x,v_y,v_z)_\odot &= (-11.1, 220+24, 7.25)~{\rm km}~{\rm s}^{-1}
+\end{align}
+\citep{8kpc 	ref??, schonrich10, bovy12}, (2) integrate the orbit forward and backward by $t_f$ and $t_b$, respectively, in the potential $\Phi$, (3) transform all orbit points (time-steps) back to observed coordinates, and (4) define interpolating functions for each coordinate as a function of stream longitude, $\phi_1$, using cubic splines---e.g., functions $\widetilde{\phi}_{2}(\phi_1)$, $\widetilde{\DM}(\phi_1)$, $\widetilde{\mu_l}(\phi_1)$, $\widetilde{\mu_b}(\phi_1)$, $\widetilde{v_r}(\phi_1)$.
+These functions let us compute the predicted values of each of these coordinates at the longitudes of each observed star, $\phi_{1,i}$.
+
+We assume that each observed kinematic component is independent so that the likelihood of the data for a given star, $\bs{x}_i$, with uncertainties, $\bs{\sigma}_i$, is given by the product over the likelihoods for each dimension of the data:
+\begin{multline}
+	p(\bs{x}_i \given \bs{\sigma}_i, \bs{s}, \bs{\theta}) = p(\phi_{2,i} \given \phi_{1,i}, s_{\phi_2},\bs{\theta}) \, p(\DM_i \given \phi_{1,i}, \sigma_{\DM,i}, s_\DM, \bs{\theta})\\ 
+	\times p(\mu_{l,i} \given \phi_{1,i}, \sigma_{\mu_{l},i}, \bs{\theta}) \, p(\mu_{b,i} \given \phi_{1,i}, \sigma_{\mu_{b},i}, \bs{\theta}) \, p(v_{r,i} \given \phi_{1,i}, \sigma_{v_r,i}, s_{v_r}, \bs{\theta}).
+\end{multline}
+The uncertainties in these observed coordinate components are assumed to be normally distributed away from the model values: using the notation
+\begin{align}
+	\norm(x \given \mu, \sigma^2) &= \frac{1}{\sqrt{2\pi \sigma^2}} \, \exp\left(-\frac{(x-\mu)^2}{2\sigma^2}\right)
+\end{align}
+the likelihoods are
+\begin{align}	
+	p(\phi_{2,i} \given \phi_{1,i}, s_{\phi_2},\bs{\theta}) &= \norm(\phi_{2,i} \given \widetilde{\phi}_{2}(\phi_{1,i}), s^2_{\phi_2})\\
+	p(\DM_i \given \phi_{1,i}, \sigma_{\DM,i}, s_\DM, \bs{\theta}) &= \norm(\DM_i \given \widetilde{\DM}(\phi_{1,i}), s^2_{\DM} + \sigma^2_{\DM,i})\\
+	p(\mu_{l,i} \given \phi_{1,i}, \sigma_{\mu_{l},i}, \bs{\theta}) &= \norm(\mu_{l,i} \given \widetilde{\mu_{l}}(\phi_{1,i}), \sigma^2_{\mu_{l,i}})\\
+	p(\mu_{b,i} \given \phi_{1,i}, \sigma_{\mu_{b},i}, \bs{\theta}) &= \norm(\mu_{b,i} \given \widetilde{\mu_{b}}(\phi_{1,i}), \sigma^2_{\mu_{b,i}})\\
+	p(v_{r,i} \given \phi_{1,i}, \sigma_{v_r,i}, s_{v_r}, \bs{\theta}) &= \norm(v_{r,i} \given \widetilde{v_r}(\phi_{1,i}), s^2_{v_r} + \sigma^2_{v_r,i}).
+\end{align}
+We assume the data from each star is independent and identically distributed (i.i.d.) so that the full likelihood is the product over the likelihoods for all $N$ stars:
+\begin{equation}
+	 p(\{\bs{x}_i\} \given \{\bs{\sigma}_i\}, \bs{s}, \bs{\theta}) = \prod_i^N p(\bs{x}_i \given \bs{\sigma}_i, \bs{s}, \bs{\theta}).\label{eq:likelihood}
+\end{equation}
+
+\subsection{Priors}
+
+For the intrinsic dispersion parameters, we use logarithmic (scale-invariant) priors such that $p(s) \propto s^{-1}$.
+For the integration time parameters, we use uniform priors, $\mathcal{U}(a,b)$ (over the range $a$--$b$),
+\begin{align}
+	p(t_f) &= \mathcal{U}(1,100)~{\rm Myr}\label{eq:prior1}\\
+	p(t_b) &= \mathcal{U}(-100,-1)~{\rm Myr}.
+\end{align}
+Note that present-day is $t=0$.
+For computational efficiency, we place strong priors on the minimum and maximum longitudes of the model points, $(\phi_{1,{\rm min}},\phi_{1,{\rm max}})$ so that the model orbit does not integrate for longer than necessary.
+In particular, we set
+\begin{align}
+	p(\phi_{1,{\rm min}} \given \bs{\theta}) &= \norm(\phi_{1,{\rm min}} \given \min(\phi_{1,i}), s^2_{\phi_2})\\
+	p(\phi_{1,{\rm max}} \given \bs{\theta}) &= \norm(\phi_{1,{\rm max}} \given \max(\phi_{1,i}), s^2_{\phi_2}).
+\end{align}
+For the orbital initial condition components, we use uniform priors in each cartesian position component over the range $(-200,200)~{\rm kpc}$.
+For velocity, we use a Gaussian prior on the magnitude of the total velocity, $v$, with a dispersion of $150~\kms$,
+\begin{equation}
+	\norm(v \given 0, (150~\kms)^2) \label{eq:prior2}
+\end{equation}
+We keep the potential, $\Phi$, fixed. In total, this model has 10 parameters (5 phase-space coordinates, 5 nuisance parameters).
+
+The full expression for the posterior probability, $p(\bs{s}, \bs{w}_0, t_b, t_f \given \{\bs{x}_i\}, \{\bs{\sigma}_i\}, \Phi)$, is the joint likelihood (Equation~\ref{eq:likelihood}) multiplied by all priors described above (Equations~\ref{eq:prior1}--\ref{eq:prior2}).
+
 % ---------------------------------------------------------------------------------
@@
 \caption{{\it Left column:} Circular velocity curves along the Sun-Galactic center line for a representative barred MW potential model (top left) and for the static MW potential model (bottom left).
-Solid black line shows total (sum of all components), lines below show a decomposition by potential component.
+Solid black line shows total (sum of all components), lines below show a decomposition by potential components.
 Vertical grey bar shows approximate position of the Sun, horizontal grey bar shows roughly the range in measured circular velocity of the Sun.
@@
 \includegraphics[width=\textwidth]{figures/surface-density}
-\caption{ Surface density of mock star particles ... \todo{This is with the same distance cut as used for the BHB stars -- is that what we want to use?}}	
+\caption{ Surface density of mock stream star particles in each potential.
+Contours are spaced logarithmically from $10^{-2}$ to $10$ particles per sq.
+deg---that is, each color represents a factor of 10 difference in surface density.
+In the static potential (top left), the density remains high along the center of the stream, but for some of the barred potentials the density drops sharply because of chaotic stream-fanning.}
 \label{fig:surfacedensity}

### 2015-12-26 ms.tex
apj format

@@
-%\documentclass{emulateapj}
-\documentclass[letterpaper,12pt,preprint]{aastex}
+\documentclass[numberedappendix]{emulateapj}
+%\documentclass[letterpaper,12pt,preprint]{aastex}
 
@@
 
+% ---------------------------------------------------------------------------------
+\begin{figure*}[!tbp]
+\begin{center}
+\includegraphics[width=0.7\textwidth]{figures/potentials-four}
+\caption{{\it Left column:} Circular velocity curves along the Sun-Galactic center line for a representative barred MW potential model (top left) and for the static MW potential model (bottom left).
+Solid black line shows total (sum of all components), lines below show a decomposition by potential components.
+Vertical grey bar shows approximate position of the Sun, horizontal grey bar shows roughly the range in measured circular velocity of the Sun.
+{\it Right column:} Contours of constant surface density for a barred MW potential (top right) and the static MW potential model (bottom right).
+Four contours are drawn per decade in surface density between $10^7$ and $10^{12}~\msun~{\rm kpc}^{-3}$.
+Note the perturbation from the bar potential within Galactocentric radius $r \lesssim 4~{\rm kpc}$.
+The Sun's position is indicated by the `$\odot$' symbol.
+}
+\label{fig:potentials}
+\end{center}
+\end{figure*}
+
 The orbit fit and $N$-body simulations in S15a used a static, axisymmetric potential to represent the Milky Way potential, but it is now well-known that the Galactic bulge contains a triaxial, bar-like structure several kpc in size \citep[e.g.,][]{blitzXX, wegg13, MANY}.
@@
 \subsubsection{Barred potential}
-We use a flattened Navarro-Frenk-White potential to represent the dark matter halo \citep[e.g.,][]{kuepper15} parametrized as
+We use a spherical Navarro-Frenk-White potential to represent the dark matter halo \citep[e.g.,][]{kuepper15} parametrized as
 \begin{align}
-	\Phi(r) &= -v_h^2\,\frac{\ln{(1 + r/r_s)}}{r/r_s}\label{eq:flatnfw}\\
-	r^2 &= x^2 + y^2 + \frac{z^2}{q_z^2}
+	\Phi(r) &= -v_h^2\,\frac{\ln{(1 + r/r_s)}}{r/r_s}\label{eq:nfw}
 \end{align}
@@
 
-\begin{table*}[ht]
+\begin{table}[ht]
 \begin{center}
@@
 	         Component & Parameter & Value \\\toprule
-		Disk & $M_{\rm disk}$ & $5 \times 10^{10}~\msun$ \\
+		Disk & $M_{\rm disk}$ & $4 \times 10^{10}~\msun$ \\
 		& $a$ & 3~{\rm kpc}\\
 		& $b$ & 0.28~{\rm kpc} \\\midrule
+		Spheroid & $M_{\rm sph}$ & $5 \times 10^{9}~\msun$ \\
+		& $c$ & 0.2 \\\midrule
 	         Halo & $v_c$ & 185.8~\kms\\
 		& $r_s$ & 30~kpc \\
-		& $q_z$ & 0.9 \\\midrule
 		Bar & $M_{\rm bar}$ & $1.8 \times 10^{10}~\msun$ \\
@@
 	\caption{The disk potential scale lengths ($a$, $b$) were adopted following \citep{bovy15-galpy} to match the exponential scale length of the disk \citep{bovyrix13} and local dark-matter density \citep[e.g.,][]{bovytremaine12}.
-Halo flattening is taken from recent work that suggests the inner Milky Way halo is slightly oblate \citep{koposov10, kuepper15}.
-The halo mass scale is set by specifying the circular velocity at the scale radius, $v_c$, and the scale velocity in Equation~\ref{eq:flatnfw} is given by $v_h^2 = v_c^2 / (\ln2 - 1/2)$.
+The halo mass scale is set by specifying the circular velocity at the scale radius, $v_c$, and the scale velocity in Equation~\ref{eq:nfw} is given by $v_h^2 = v_c^2 / (\ln2 - 1/2)$.
 The bar mass is taken from recent 3D density modeling of red clump stars in the Galactic bulge \citep{portail15}.
@@
 \end{center}
-\end{table*}
+\end{table}
 
-\begin{table*}[ht]
+\begin{table}[ht]
 \begin{center}
@@
 \end{center}
-\end{table*}
+\end{table}
 
@@
 	         Component & Parameter & Value \\\toprule
-		Disk & $M_{\rm disk}$ & $6.5 \times 10^{10}~\msun$ \\
+		Disk & $M_{\rm disk}$ & $6 \times 10^{10}~\msun$ \\
 		& $a$ & 3~{\rm kpc}\\
@@
 	         Halo & $v_c$ & 185.8~\kms\\
-		& $r_s$ & 30~kpc \\
-		& $q_z$ & 0.9 \\\midrule
-		Spheroid & $M_{\rm sph}$ & $10^{10}~\msun$ \\
-		& $c$ & 0.2 \\
+		& $r_s$ & 30~kpc \\\midrule
+		Spheroid & $M_{\rm sph}$ & $1.2 \times 10^{10}~\msun$ \\
+		& $c$ & 0.3 \\
 		\bottomrule
@@
 
+% ---------------------------------------------------------------------------------
+\begin{figure}[!tbp]
+\begin{center}
+\includegraphics[width=0.5\textwidth]{figures/orbitfits}
+\caption{ Results from fitting orbits to BHB stars associated with the Ophiuchus stream in the static and barred potential models.
+Data are shown as black points with grey error bars.
+Error bars may sometimes be smaller than the point size.
+Lines (blue) show sections of orbits integrated forward and backwards from initial conditions drawn from the posterior samples generated by MCMC (See Section~\ref{sec:orbitfit}).
+Note that the four higher dispersion stars (the four stars with highest longitude) were not used when computing the likelihood and are only shown for completeness.
+}
+\label{fig:orbitfits}
+\end{center}
+\end{figure}
+
 We make one modification to this method based on the idea that the Ophiuchus stream progenitor has been fully-disrupted.
@@
 
+% ---------------------------------------------------------------------------------
+\begin{figure}[!tbp]
+\begin{center}
+\includegraphics[width=0.5\textwidth]{figures/orbit-yz}
+\caption{  Projections of orbits integrated from the mean orbital parameters estimated from the orbit fitting posterior distributions in each potential model.
+Orbits are integrated for 6 Gyr and shown in Galactocentric, Cartesian coordinates.
+Even though the mean orbital parameters have nearly identical values (e.g., the initial conditions are nearly identical in Heliocentric coordinates), the orbits in each potential model are  different in appearance.
+}
+\label{fig:orbits-yz}
+\end{center}
+\end{figure}
+
 The orbits sampled from the orbit-fit posteriors in the barred potentials are all strongly chaotic.
 We have tried computing the frequency diffusion rate for these orbits as an independent check of their chaotic timescale but have found that, over consecutive integration windows, the frequency recovery fails or is unreliable because the frequency spectrum changes dramatically over timescales of $\approx$10--20 orbital periods.
+
+% ---------------------------------------------------------------------------------
+\begin{figure}[!tbp]
+\begin{center}
+\includegraphics[width=0.5\textwidth]{figures/lyapunov-hist}
+\caption{ Histograms of estimates of the Lyapunov time, $\lyapt=1/\lyapexp$, for posterior samples from the orbit fitting procedure (Section~\ref{sec:orbitfit}).
+Vertical lines show the Lyapunov time value of the mean orbit in that potential model.
+Orbits are typically more chaotic (have shorter Lyapunov times) for higher pattern speeds.
+There is little dependence on bar angle.}
+\label{fig:lyapunov-hist}
+\end{center}
+\end{figure}
+
+\section{Results 2: Stream models for the Ophiuchus stream}
+
+We randomly sample 256 orbits from the orbital parameter posterior distributions and generate mock streams along each orbit.
+We use the method outlined above (Section \ref{sec:mocks}) to generate the streams and set the free parameters as follows: (1) we evolve the progenitors for 6 Gyr along each orbit, (2) we release star particles every 0.5 Myr (uniformly in time) to densely sample the final density distribution, (3) we set the progenitor mass to $m=10^4~\msun$, and (4) we set the disruption time of the progenitor equal to the last time at which a pericenter and disk crossing coincide (in each case, this is at $t \approx -200~{\rm Myr}$).
+After the disruption time, we continue releasing the star particles uniformly in time rather than releasing a ``burst'' of particles at once.
+We therefore expect that the morphology and value of the density of the most recently disrupted debris will not exactly match the data.
 
@@
 \end{table*}
-
-\section{Results 2: Stream models for the Ophiuchus stream}
-
-We randomly sample 256 orbits from the orbital parameter posterior distributions and generate mock streams along each orbit.
-We use the method outlined above (Section \ref{sec:mocks}) to generate the streams and set the free parameters as follows: (1) we evolve the progenitors for 6 Gyr along each orbit, (2) we release star particles every 0.5 Myr (uniformly in time) to densely sample the final density distribution, (3) we set the progenitor mass to $m=10^4~\msun$, and (4) we set the disruption time of the progenitor equal to the last time at which a pericenter and disk crossing coincide (in each case, this is at $t \approx -200~{\rm Myr}$).
-After the disruption time, we continue releasing the star particles uniformly in time rather than releasing a ``burst'' of particles at once.
-We therefore expect that the morphology and value of the density of the most recently disrupted debris will not exactly match the data.
 
@@
 APW acknowledges the staff at the MPIA for...
-The authors wish to acknowledge Victor Debattista, ...
+The authors wish to acknowledge Victor Debattista, Melissa Ness, ...?
 This research made use of Astropy, a community-developed core Python package for Astronomy \citep{astropy13}.
 This work additionally relied on Columbia University's \emph{Yeti} compute cluster, and we acknowledge the Columbia HPC support staff for assistance. \\
+
+% ---------------------------------------------------------------------------------
+
+\begin{figure*}[p]
+\begin{center}
+\includegraphics[width=0.8\textwidth]{figures/mockstream0}
+\caption{ Sky position, distance, and line-of-sight velocity in Heliocentric, Galactic coordinates for star particles (grey points) from mock streams generated on the mean orbits in each potential.
+Vertical, dashed lines show the approximate extent of the densest part of the stream visible in main-sequence stars \citep[the segment originally detected in ][]{bernard14}.}
+\label{fig:mockstream0}
+\end{center}
+\end{figure*}
+
+\begin{figure*}[p]
+\begin{center}
+\includegraphics[width=0.8\textwidth]{figures/mockstream1}
+\caption{ Same as Figure~\ref{fig:mockstream0} for the other five barred potentials. }
+\label{fig:mockstream1}
+\end{center}
+\end{figure*}
+
+\begin{figure*}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/surface-density}
+\caption{ Surface density of mock stream star particles in each potential.
+Contours are spaced logarithmically from $10^{-2}$ to $10$ particles per sq.
+deg---that is, each color represents a factor of 10 difference in surface density.
+In the static potential (top left), the density remains high along the center of the stream, but for some of the barred potentials the density drops sharply because of chaotic stream-fanning.}
+\label{fig:surfacedensity}
+\end{center}
+\end{figure*}
+
+% ---------------------------------------------------------------------------------
+
+\begin{figure*}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/mockstream-xyz}
+\caption{ Star particles (grey points) from mock streams generated on the mean orbits in each potential model shown in projections of Galatocentric, Cartesian coordinates.
+The position of the Sun is shown as the symbol at $(x,z)=(-8,0)~{\rm kpc}$.
+The corners of the sky position and distance plots of Figures~\ref{fig:mockstream0}--\ref{fig:mockstream1} are shown transformed to these coordinates as dark markers near $(x,z)=(-4,3)~{\rm kpc}$, $(x,z)=(0,4)~{\rm kpc}$.
+This demonstrates that the stream is nearly aligned our viewing angle.
+}
+\label{fig:mockstreamxyz}
+\end{center}
+\end{figure*}
+
+% ---------------------------------------------------------------------------------
 
@@
 % ---------------------------------------------------------------------------------
-\clearpage
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/potentials-four}
-\caption{{\it Left column:} Circular velocity curves along the Sun-Galactic center line for a representative barred MW potential model (top left) and for the static MW potential model (bottom left).
-Solid black line shows total (sum of all components), lines below show a decomposition by potential components.
-Vertical grey bar shows approximate position of the Sun, horizontal grey bar shows roughly the range in measured circular velocity of the Sun.
-{\it Right column:} Contours of constant surface density for a barred MW potential (top right) and the static MW potential model (bottom right).
-Four contours are drawn per decade in surface density between $10^7$ and $10^{12}~\msun~{\rm kpc}^{-3}$.
-Note the perturbation from the bar potential within Galactocentric radius $r \lesssim 4~{\rm kpc}$.
-The Sun's position is indicated by the `$\odot$' symbol.
-}
-\label{fig:potentials}
-\end{center}
-\end{figure}
-
-% ---------------------------------------------------------------------------------
 %\clearpage
@@
 % ---------------------------------------------------------------------------------
-\clearpage
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/orbitfits}
-\caption{ Results from fitting orbits to BHB stars associated with the Ophiuchus stream in the static and barred potential models.
-Data are shown as black points with grey error bars.
-Error bars may sometimes be smaller than the point size.
-Lines (blue) show sections of orbits integrated forward and backwards from initial conditions drawn from the posterior samples generated by MCMC (See Section~\ref{sec:orbitfit}).
-Note that the four higher dispersion stars (the four stars with highest longitude) were not used when computing the likelihood and are only shown for completeness.
-}
-\label{fig:orbitfits}
-\end{center}
-\end{figure}
-
-% ---------------------------------------------------------------------------------
-\clearpage
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/orbit-yz}
-\caption{  Projections of orbits integrated from the mean orbital parameters estimated from the orbit fitting posterior distributions in each potential model.
-Orbits are integrated for 6 Gyr and shown in Galactocentric, Cartesian coordinates.
-Even though the mean orbital parameters have nearly identical values (e.g., the initial conditions are nearly identical in Heliocentric coordinates), the orbits in each potential model are  different in appearance.
-}
-\label{fig:orbits-yz}
-\end{center}
-\end{figure}
-
-% ---------------------------------------------------------------------------------
 %\clearpage
@@
 % ---------------------------------------------------------------------------------
-\clearpage
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/lyapunov-hist}
-\caption{ Histograms of estimates of the Lyapunov time, $\lyapt=1/\lyapexp$, for posterior samples from the orbit fitting procedure (Section~\ref{sec:orbitfit}).
-Vertical lines show the Lyapunov time value of the mean orbit in that potential model.
-Orbits are typically more chaotic (have shorter Lyapunov times) for higher pattern speeds.
-There is little dependence on bar angle.}
-\label{fig:lyapunov-hist}
-\end{center}
-\end{figure}
-
-% ---------------------------------------------------------------------------------
-\clearpage
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/mockstream0}
-\caption{ Sky position, distance, and line-of-sight velocity in Heliocentric, Galactic coordinates for star particles (grey points) from mock streams generated on the mean orbits in each potential.
-Vertical, dashed lines show the approximate extent of the densest part of the stream visible in main-sequence stars \citep[the segment originally detected in ][]{bernard14}.}
-\label{fig:mockstream0}
-\end{center}
-\end{figure}
-
-\clearpage
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/mockstream1}
-\caption{ Same as Figure~\ref{fig:mockstream0} for the other five barred potentials. }
-\label{fig:mockstream1}
-\end{center}
-\end{figure}
-
-\clearpage
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/surface-density}
-\caption{ Surface density of mock stream star particles in each potential.
-Contours are spaced logarithmically from $10^{-2}$ to $10$ particles per sq.
-deg---that is, each color represents a factor of 10 difference in surface density.
-In the static potential (top left), the density remains high along the center of the stream, but for some of the barred potentials the density drops sharply because of chaotic stream-fanning.}
-\label{fig:surfacedensity}
-\end{center}
-\end{figure}
-
-% ---------------------------------------------------------------------------------
-\clearpage
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/mockstream-xyz}
-\caption{ Star particles (grey points) from mock streams generated on the mean orbits in each potential model shown in projections of Galatocentric, Cartesian coordinates.
-The position of the Sun is shown as the symbol at $(x,z)=(-8,0)~{\rm kpc}$.
-The corners of the sky position and distance plots of Figures~\ref{fig:mockstream0}--\ref{fig:mockstream1} are shown transformed to these coordinates as dark markers near $(x,z)=(-4,3)~{\rm kpc}$, $(x,z)=(0,4)~{\rm kpc}$.
-This demonstrates that the stream is nearly aligned our viewing angle.
-}
-\label{fig:mockstreamxyz}
-\end{center}
-\end{figure}
-
-% ---------------------------------------------------------------------------------
 %\clearpage

### 2015-12-31 ms.tex
comments from kathryn

@@
-\documentclass[numberedappendix]{emulateapj}
-%\documentclass[letterpaper,12pt,preprint]{aastex}
+%\documentclass[numberedappendix]{emulateapj}
+\documentclass[letterpaper,12pt,preprint]{aastex}
 
@@
 
+\todo{Still working on the abstract!}
 % Context
 The Ophiuchus stream was recently discovered as an over-density of main sequence stars in Pan-STARRS (PS1) photometry and is peculiar in it's short apparent length given the age of its stellar population.
-The stream was later detected in BHB stars by identifying members using line-of-sight velocities: Twenty Blue Horizontal Branch (BHB) stars in the vicinity of the stream are kinematically distinct from the background halo stars and have distances consistent with the previously detected segment from PS1.
-Four of these BHB stars near the end of the stream have similarly distinct velocities but have much larger velocity and spatial dispersions compared to the PS1 stream segment.
+The stream was later detected in Blue Horizontal Branch (BHB) stars by identifying members using line-of-sight velocities: Twenty BHB stars in the vicinity of the stream are kinematically distinct from the background halo stars and have distances consistent with the previously detected segment from PS1.
+Four of these BHB stars beyond the apparent end of PS1 stars have similarly distinct velocities but have much larger velocity and spatial dispersions. % compared to the PS1 stream segment.
 In recent work, we have shown that streams formed on chaotic orbits may display similar peculiar morphologies: short segments of coherent stellar debris flanked by high-dispersion, low-density ``stream-fans'' at the ends of the stream.
@@
 
-The Ophiuchus stream \citep{bernard14, sesar15a} is a recently discovered stellar tidal stream that sits above the Galactic bulge at a Galactocentric radius and height $(R,z) \approx (1, 4.3)~{\rm kpc}$.
-All observational evidence suggests that the stream is a totally disrupted globular cluster: The stream stars have (1) a small positional dispersion orthogonal to the extended direction of the stream ($\approx$10 pc); (2) no detectable over-density along the stream that could be the progenitor system; (3) a small velocity dispersion $\approx$0.4 ${\rm km}~{\rm s}^{-1}$; and (4) an old stellar population ($\approx$12 Gyr) estimated from isochrone fitting \citep[][hereafter S15a]{sesar15a}.
+The Ophiuchus stream \citep{bernard14, sesar15a} is a recently discovered stellar tidal stream that sits above the Galactic bulge at a Galactocentric radius and height $(R,z) \approx (1.5, 4.3)~{\rm kpc}$.
+All observational evidence suggests that the stream is a totally disrupted globular cluster: The stream stars have (1) a small positional dispersion orthogonal to the extended direction of the stream (width $\approx$10 pc, length $\approx$1.5 kpc); (2) no detectable over-density along the stream that could be the progenitor system; (3) a small velocity dispersion $\approx$0.4 ${\rm km}~{\rm s}^{-1}$; and (4) an old stellar population ($\approx$12 Gyr) estimated from isochrone fitting \citep[][hereafter S15a]{sesar15a}.
 
@@
 S15a find that---on this orbit---the portion of the stream visible as an over-density in main-sequence stars must have been formed in the last $\lesssim$400 Myr for the stream to remain as short as it is observed.
-This dynamical age is at odds with the old stellar population.
+This dynamical age is at odds with the old stellar population: The stars appear to have formed $\approx$12 Gyr ago, but the cluster apparently fully disrupted at once in the last 400 Myrs.
 Another puzzle is the existence of blue horizontal branch (BHB) stars close to the stream (within a few degrees) with similar radial velocities, but with a large dispersion in both sky position and velocity \citep[][hereafter S15b]{sesar15b}.
 The stream has a very distinct and large line-of-sight velocity ($\approx$290 \kms) and is therefore easily detected above the background halo population.
-Four stars have been detected with line-of-sight velocities $>230~\kms$ that lie close to an extrapolation of the stream on the sky, but have a velocity dispersion $\approx$75 times larger than the measured internal velocity dispersion of the stream stars.
+Four BHB stars have been detected with line-of-sight velocities $>230~\kms$ that lie close to an extrapolation of the stream on the sky, but have a velocity dispersion $\approx$75 times larger than the measured internal velocity dispersion of the stream stars.
+The velocities of these stars are highly inconsistent with belonging to the background halo population (see Section 4, S15b).
 These stars hint at the existence of associated low-density, high-dispersion features that were not modeled in S15a and are not predicted by the $N$-body simulations from this prior work.
@@
 
-The orbit fit and $N$-body simulations in S15a used a static, axisymmetric potential to represent the Milky Way potential, but it is now well-known that the Galactic bulge contains a triaxial, bar-like structure several kpc in size \citep[e.g.,][]{blitzXX, wegg13, MANY}.
+The orbit fit and $N$-body simulations in S15a used a static, axisymmetric potential to represent the Milky Way potential, but it is well-known that the Galactic bulge contains a triaxial, bar-like structure several kpc in size \citep[e.g.,][]{blitz91, weinberg92, dwek95, wegg13}.
 Given the proximity of the stream to the center of the Galaxy, the time-dependent, triaxial potential of the Galactic bar must be taken into account when modeling the orbit of the Ophiuchus stream.
-The presence of a bar-like perturbation to the potential will change the orbit of the stream progenitor and the orbit structure in the inner galaxy \citep{athanassoula, portail15b, zotos12}.
+The presence of a bar-like perturbation to the potential will change the orbit of the stream progenitor and the orbit structure in the inner galaxy \citep{portail15b, gajda15, zotos12}.
 Bar-like features can also introduce a significant number of chaotic orbits in their vicinity\citep{weinberg15}.
 
-Recent work has shown that chaos can dramatically alter the density evolution of tidal streams.
-Along certain chaotic orbits, the stream stars will spread much faster in 3D position than from ordinary phase-mixing and, depending on the orbital phase at which the stream is observed, may develop large, low-density ``fans'' of stars at the ends of a stream \citep{apw15-chaos}.
-In this work, we study whether stream-fanning---chaotic or simply from density evolution in a triaxial, time-dependent potential---could plausibly explain the observed properties of the Ophiuchus stream.
+Recent work has shown that chaos can dramatically alter the density evolution of tidal streams \citep[e.g.,][]{fardal14, apw15-chaos}.
+Along certain chaotic orbits, the stream stars will spread much faster in 3D position than from ordinary phase-mixing and, depending on the orbital phase at which the stream is observed, may develop large, low-density ``fans'' of stars at the ends of a stream \citep{pearson15, apw15-chaos}.
+As a first application of this theoretical understanding, we study whether stream-fanning---chaotic or simply from density evolution in a triaxial, time-dependent potential---could plausibly explain the observed properties of the Ophiuchus stream.
 In particular, we consider whether such models can reproduce:
@@
 \end{enumerate}
+Our goal is not to perfectly represent the observed data but rather to explore the plausibility of explaining the peculiarities of the stream using chaotic stream fanning.
 
@@
 
-Our goal is to (1) assess whether the Galactic bar can produce chaotic orbits in the vicinity of the Ophiuchus stream and (2) determine if chaotic density evolution of tidal debris stripped from the progenitor of this stream can explain the apparent shortness of the stream and low-density, high-dispersion ends.
+Our goal is to (1) assess whether the Galactic bar can produce chaotic orbits in the vicinity of the Ophiuchus stream and (2) determine if chaotic density evolution of tidal debris stripped from the progenitor of this stream can explain the apparent shortness of the stream and low-density, high-dispersion stars beyond the extent of main sequence stars observed in PS1.
 In this section, we describe the potential models we use to represent the galaxy and describe the methods we use to detect and quantify the strength of chaos for orbits.
@@
 Finally, we describe the method we use to generate mock stellar streams.
+
+In this work, we assume the Sun is at Galactocentric position $(x,y,z) = (-8.3,0,0)~{\rm kpc}$ \citep[e.g.,][]{schoenrich12} with velocity $(v_x,v_y,v_z) = (-11.1, 250, 7.25)~\kms$ \citep[e.g.,][]{schoenrich10, schoenrich12}.
 
@@
 We also define a static potential with a spherical bulge for comparison.
-These potential models are meant to be representative rather than definitive since the potential of the inner and outer Milky Way are not well known.
+
+These potential models are meant to be representative rather than definitive.
+The uncertainty in the Milky Way potential within Galactocentric radii of $r\lesssim 4~{\rm kpc}$ and outside of $r\gtrsim 15~{\rm kpc}$ are large enough that trying to match the exact density distribution of the Ophiuchus stream is not a useful exercise.
+Instead, we consider qualitatively different potentials that allow us to isolate and study the affect of chaotic stream-fanning of tidal debris in the vicinity of the stream.
+\todo{KVJ -- is this better?}
 
 \subsubsection{Barred potential}
-We use a spherical Navarro-Frenk-White potential to represent the dark matter halo \citep[e.g.,][]{kuepper15} parametrized as
+We use a spherical Navarro-Frenk-White potential to represent the dark matter halo \citep{navarro96} parametrized as
 \begin{align}
@@
 This method has been shown to reproduce the morphologies of $N$-body simulations of stellar streams, but requires far less computing time because it relies only on integrating test-particle orbits.
-\todo{Does this need more detail  about Fardal's method?}
 
@@
 \caption{ Results from fitting orbits to BHB stars associated with the Ophiuchus stream in the static and barred potential models.
-Data are shown as black points with grey error bars.
+Data used for computing the likelihood are shown as black points with grey error bars.
+The four ``fanned'' BHB stars from S15b excluded from the likelihood computation are shown as grey squares.
 Error bars may sometimes be smaller than the point size.
@@
 Any star released after $\tau_d$ is released with a small dispersion around the progenitor orbit but no offset.
+This mass-loss history is intended to mimic the expected gradual evaporation of a globular cluster over a tidally-limited boundary (i.e.
+driven by two-body relaxation and gravitational shocks over many Gigayears) with final disruption likely to occur once the tidal boundary is less than the core radius of the cluster.
+The physics of this disruption is not followed exactly but rather the disruption rate and final disruption time are set by matching the observed density along the stream.
+\todo{APW: check that last statement -- can we be more precise?}
 
@@
 
+Here we study whether the observed abrupt drop in density and possible fanned debris stars can be explained without assuming a sudden change in the mass-loss history of the cluster.
+In particular, we are interested in whether the mock streams formed around the strongly chaotic progenitor orbits in the barred potentials can explain these features while having been steadily disrupted over many Gigayears.
+
 We randomly sample 256 orbits from the orbital parameter posterior distributions and generate mock streams along each orbit.
-We use the method outlined above (Section \ref{sec:mocks}) to generate the streams and set the free parameters as follows: (1) we evolve the progenitors for 6 Gyr along each orbit, (2) we release star particles every 0.5 Myr (uniformly in time) to densely sample the final density distribution, (3) we set the progenitor mass to $m=10^4~\msun$, and (4) we set the disruption time of the progenitor equal to the last time at which a pericenter and disk crossing coincide (in each case, this is at $t \approx -200~{\rm Myr}$).
+We use the method outlined above (Section \ref{sec:mocks}) to generate the streams and set the free parameters as follows: (1) we evolve the progenitors for 6 Gyr along each orbit to explore stream models where the shortness of the stream is \emph{not} due to an instantaneous disruption 400 Myrs ago, (2) we release star particles every 0.5 Myr (uniformly in time) to densely sample the final density distribution, (3) we set the progenitor mass to $m=10^4~\msun$ (as was estimated by S15a), and (4) we set the disruption time of the progenitor equal to the last time at which a pericenter and disk crossing coincide (in each case, this is at $t \approx -200~{\rm Myr}$).
 After the disruption time, we continue releasing the star particles uniformly in time rather than releasing a ``burst'' of particles at once.
-We therefore expect that the morphology and value of the density of the most recently disrupted debris will not exactly match the data.
+We therefore expect that the density of the most recently disrupted debris will be systematically higher for the model streams as compared to the observed stream.
 
@@
 	name & $\phi_2$ [deg] & $d$ [kpc] & $\mu_l$ [mas yr$^{-1}$] & $\mu_b$ [mas yr$^{-1}$] & $v_r$ [km s$^{-1}$]\\\midrule
-	static & $-0.03\pm0.05$ & $8.3\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.1\pm0.9$\\
-	bar1--9 & $-0.03\pm0.05$ & $8.35\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.0\pm1.0$\\
+	static & $-0.03\pm0.05$ & $8.3\pm0.05$ & $-7.4\pm0.1$ & $0.9\pm0.1$ & $288.9\pm0.9$\\
+	bar1--9 & $-0.03\pm0.05$ & $8.35\pm0.05$ & $-7.4\pm0.1$ & $0.9\pm0.1$ & $289.0\pm1.0$\\
 	\bottomrule
@@
 
-For each generated mock stream, we compute the likelihood of the data given the star particles by estimating the phase-space model density using a kernel density estimate with a Gaussian kernel \citep[see, e.g.,][]{bonaca14}.
-For each $i$ data point and each $k$ model point with coordinates $\bs{y}_k$, we compute the likelihood by converting the model point position and velocity into Heliocentric coordinates and evaluate
+For each generated mock stream, we compute the likelihood of the data (now including all BHB stars from S15a and S15b, e.g., all points in Figure~\ref{fig:orbitfits}) given the star particles by estimating the phase-space model density using a kernel density estimate with a Gaussian kernel \citep[see, e.g.,][]{bonaca14}.
+For each $i$ data point $\bs{x}_i$ and each $k$ model point with coordinates $\bs{y}_k$, we compute the likelihood by converting the model point position and velocity into Heliocentric coordinates and evaluate
 \begin{align}
@@
 	0.2~{\rm kpc}\\
-	0.2~{\rm mas}~{\rm yr}^{-1}\\
-	0.2~{\rm mas}~{\rm yr}^{-1}\\
 	0.2~\kms
@@
 \end{equation} %(max$\mathcal{L}$)
-for sky position in Galactic coordinates ($l$, $b$), distance, proper motions ($\mu_l$, $\mu_b$), and radial velocity. The full likelihood for all $N$ data points given $K$ model points is 
+for sky position in Galactic coordinates ($l$, $b$), distance, and radial velocity. The full likelihood for all $N$ data points given $K$ model points is 
 \begin{equation}
@@
 This is a projection effect: The portion of the stream at larger $l$ is closer points almost directly towards the Sun so that the debris covers a larger area.
-This is more clearly seen in Figure~\ref{fig:mockstreamxyz}, which shows $x$-$z$ projections of the star particles in Galactocentric Cartesian coordinates (grey) along with the position of the Sun (symbol at $(x,z)=(-8,0)~{\rm kpc}$) and the ``window'' of the Heliocentric, sky-position plots of Figures~\ref{fig:mockstream0}--\ref{fig:mockstream1} (shown as blue lines).
+This is more clearly seen in Figure~\ref{fig:mockstreamxyz}, which shows $x$-$z$ projections of the star particles in Galactocentric Cartesian coordinates (grey) along with the position of the Sun (symbol at $(x,z)=(-8.3,0)~{\rm kpc}$) and the ``window'' of the Heliocentric, sky-position plots of Figures~\ref{fig:mockstream0}--\ref{fig:mockstream1} (shown as blue lines).
 	\item The density of the mock stream in the static potential decreases slowly rather than abruptly as is observed.
@@
 \todo{Waiting on Branimir!}
+	\item \todo{4 BHBs beyond end of stream never fit in pos/vel for static case, but do in some bar potentials?}
 	\item For the barred potentials, the stream morphology is very sensitive to the properties of the bar (especially the pattern speed) and to the initial conditions of the orbit.
@@
 
+\todo{Still working on what to say here - suggestions welcome!}
 If the Ophiuchus stream formed from a globular cluster on a chaotic orbit and we happen to be witnessing its final demise, what does this imply about the population of clusters that have already been fully disrupted?
@@
 Comparing the observed numbers with expectations ...
-\todo{KVJ add words?}
 
@@
 APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155.
-APW acknowledges the staff at the MPIA for...
-The authors wish to acknowledge Victor Debattista, Melissa Ness, ...?
+KVJ and APW acknowledge support from NSF grant AST-1312196. 
+This material is based on work supported by the National Aeronautics and Space Administration under Grant No.\, NNX15AK78G issued through the Astrophysics Theory Program. 
+APW acknowledges the staff at the MPIA for their support and assistance.
+The authors wish to acknowledge Victor Debattista and Melissa Ness for useful discussions.
 This research made use of Astropy, a community-developed core Python package for Astronomy \citep{astropy13}.
@@
 \caption{ Sky position, distance, and line-of-sight velocity in Heliocentric, Galactic coordinates for star particles (grey points) from mock streams generated on the mean orbits in each potential.
-Vertical, dashed lines show the approximate extent of the densest part of the stream visible in main-sequence stars \citep[the segment originally detected in ][]{bernard14}.}
+Vertical, dashed lines show the approximate extent of the densest part of the stream visible in main-sequence stars \citep[the segment originally detected in ][]{bernard14}.\todo{make top panels density maps instead}}
 \label{fig:mockstream0}

### 2016-01-04 ms.tex
ahhhh

@@
 Given the proximity of the stream to the center of the Galaxy, the time-dependent, triaxial potential of the Galactic bar must be taken into account when modeling the orbit of the Ophiuchus stream.
-The presence of a bar-like perturbation to the potential will change the orbit of the stream progenitor and the orbit structure in the inner galaxy \citep{portail15b, gajda15, zotos12}.
-Bar-like features can also introduce a significant number of chaotic orbits in their vicinity\citep{weinberg15}.
+The presence of a bar-like perturbation to the potential will change the orbit of the stream progenitor and the orbit structure in the inner galaxy \citep{zotos12, portail15b, gajda15}.
+Bar-like features can also introduce a significant number of chaotic orbits in their vicinity \citep{weinberg15}.
 
@@
 Recent work has used stellar number counts of Red Clump giant stars in the Galactic bulge to constrain dynamical models of the bar \citep{portail15}.
-Measurements of the total mass of the bar feature from this study are largely consistent with past work \citep[e.g.,][]{wang12, XX}, however the measured pattern speed and present bar angle are significantly discrepant and this difference is not fully understood.
+Measurements of the total mass of the bar feature from this study are largely consistent with past work \citep[e.g.,][]{wang12}, however the measured pattern speed and present bar angle are significantly discrepant and this difference is not fully understood.
 We construct a parametrized potential model consisting of a triaxial, time-dependent (rotating) bulge component added to simple models for the disk and halo of the Milky Way.
@@
 Instead, we consider qualitatively different potentials that allow us to isolate and study the affect of chaotic stream-fanning of tidal debris in the vicinity of the stream.
-\todo{KVJ -- is this better?}
 
@@
 	\caption{Present-day bar angle ($\alpha$) and pattern speed ($\Omega_p$) for the nine parameter pairs considered in this work.
-These values span the range of recent measurements from a variety of techniques \citep{dwek95,wang12,wang13, MOREMOREMORE}.
+These values span the range of recent measurements from a variety of techniques \citep{dwek95, wang12,wang13,wegg13}.
 \label{tbl:bar-specific}}
@@
 driven by two-body relaxation and gravitational shocks over many Gigayears) with final disruption likely to occur once the tidal boundary is less than the core radius of the cluster.
-The physics of this disruption is not followed exactly but rather the disruption rate and final disruption time are set by matching the observed density along the stream.
-\todo{APW: check that last statement -- can we be more precise?}
+The physics of this disruption is not followed exactly but rather the disruption rate and final disruption time are set by \todo{the hypothesis that the most recent shock ...}
 
@@
 All orbits in the static potential have Lyapunov times $\lyapt > 20~{\rm Gyr}$ and we consider them to be regular (no panel is shown for these orbits).
-All orbits sampled from each barred potential are strongly chaotic with Lyapunov times that range from $\lyapt \approx 400--1100~{\rm Myr}$.
+All orbits sampled from each barred potential are strongly chaotic with Lyapunov times that range from $\lyapt \approx 400$--$1100~{\rm Myr}$.
 It is clear from these panels that the orbits around Ophiuchus are generally more strongly chaotic (have lower Lyapunov times) for larger pattern speeds.
@@
 \todo{Waiting on Branimir!}
-	\item \todo{4 BHBs beyond end of stream never fit in pos/vel for static case, but do in some bar potentials?}
+	\item The four high-dispersion BHB stars beyond the end of the stream (from S15b) don't match in position and velocity with the particle distribution from even the maximum likelihood stream model in the static potential.
+However, in some barred potentials the chaotic evolution of the stream stars can lead to over-densities of stars with an increased positional dispersion and significantly discrepant velocities (e.g., bar8).
+	\item None of the stream models produce an appreciable density of stars with line-of-sight velocities near the S15b BHB star with the largest velocity ($\approx 320~\kms$).
+This star is either an interesting Ophiuchus member star or is associated with some other kinematic substructure.
 	\item For the barred potentials, the stream morphology is very sensitive to the properties of the bar (especially the pattern speed) and to the initial conditions of the orbit.
-In addition to generating mock streams on the mean orbits, we have run similar simulations for random draws from the posterior distributions over orbit fits.
-We have found that the morphology can vary significantly between different samples (because these are strongly chaotic orbits), but the overall characteristics remain similar: along more strongly chaotic orbits, the debris ``fans'' more and the apparent dense part of the model streams is shorter (see, for example, Figure~\ref{fig:mockstreamxyz}, panels bar2 vs.
-bar3).
+We have found that the morphology can vary significantly between nearby orbits in the same potential model (because these are strongly chaotic orbits), but the overall characteristics remain similar: along more strongly chaotic orbits, the debris ``fans'' more and the apparent dense part of the model streams is shorter (see, for example, Figure~\ref{fig:mockstreamxyz}, panels bar3 vs.
+bar8).
 \end{enumerate}
@@
 
-The current kinematic data for the Ophiuchus BHB stars suggests that the stream is sensitive to the potential of the bar, but with better measurements of the velocities and a larger sample of member stars we would be able to constrain the bar model parameters.
-Current measurements of the pattern speed, angle, and structure of the Milky Way's bulge and bar are largely discrepant \citep[e.g.,][]{wang12, wang13, antoja14, wegg??
-is wegg appropriate here}.
+The current kinematic data for the Ophiuchus BHB stars suggests that the stream is sensitive to the gravitational potential of the Milky Way bar---with better measurements of the velocities and a larger sample of member stars we will constrain parameters for the bar model.
+Current measurements of the pattern speed, angle, and structure of the Milky Way's bulge and bar are largely discrepant \citep[e.g.,][]{wang12, wang13, wegg13, antoja14}.
 Most of the methods that infer these quantities rely on modeling the density or kinematics of stars at low Galactic latitudes and must therefore handle challenges with completeness and dust extinction.
@@
 
-\todo{Still working on what to say here - suggestions welcome!}
 If the Ophiuchus stream formed from a globular cluster on a chaotic orbit and we happen to be witnessing its final demise, what does this imply about the population of clusters that have already been fully disrupted?
-For a given model for the Galactic bar, we can study the orbit structure in the inner Milky Way and, in particular, study the fraction of strongly chaotic orbits that would lead to chaotic fanning or fast dispersal of tidal debris \citep{recent work by zotos or those greek peeps}.
-The existence of strongly chaotic orbits in this region would limit the expected number of cold stellar streams in the inner Galaxy.
-Comparing the observed numbers with expectations ...
+For a given model for the Galactic bar, we will study the orbit structure in the inner Milky Way and, in particular, study the fraction of strongly chaotic orbits that would lead to chaotic fanning or fast dispersal of tidal debris \citep{recent work by zotos or those greek peeps}.
+The existence of strongly chaotic orbits in this region would limit the expected number of cold stellar streams in the inner Galaxy and enhance the rate of mixing of the debris.
+Gaia will ....
+compare amount of substructure with expectations .....
+\todo{Quote Gnedin and Ostriker, Murali and Weinberg}
 
@@
 \includegraphics[width=0.8\textwidth]{figures/mockstream0}
-\caption{ Sky position, distance, and line-of-sight velocity in Heliocentric, Galactic coordinates for star particles (grey points) from mock streams generated on the mean orbits in each potential.
-Vertical, dashed lines show the approximate extent of the densest part of the stream visible in main-sequence stars \citep[the segment originally detected in ][]{bernard14}.\todo{make top panels density maps instead}}
+\caption{ Sky position, distance, and line-of-sight velocity in Heliocentric, Galactic coordinates for star particles (contours or grey points) from the maximum-likelihood mock streams in each potential.
+Top panels show surface density of mock stream star particles in each potential.
+Contours are spaced logarithmically from $10^{-2}$ to $10$ particles per sq.
+deg---that is, each color represents a factor of 10 difference in surface density.
+In the static potential (top left), the density remains high along the center of the stream, but for some of the barred potentials the density drops sharply because of chaotic stream-fanning.
+Vertical, dashed lines show the approximate extent of the densest part of the stream visible in main-sequence stars \citep[the segment originally detected in ][]{bernard14}.}
 \label{fig:mockstream0}
@@
 
-\begin{figure*}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/surface-density}
-\caption{ Surface density of mock stream star particles in each potential.
+%\begin{figure*}[p]
+%\begin{center}
+%\includegraphics[width=\textwidth]{figures/surface-density}
+%\caption{ Surface density of mock stream star particles in each potential.
 Contours are spaced logarithmically from $10^{-2}$ to $10$ particles per sq.
@@
 In the static potential (top left), the density remains high along the center of the stream, but for some of the barred potentials the density drops sharply because of chaotic stream-fanning.}
-\label{fig:surfacedensity}
-\end{center}
-\end{figure*}
+%\label{fig:surfacedensity}
+%\end{center}
+%\end{figure*}
 
@@
 
-For a given set of initial conditions ($\bs{w}_0$), we compute a model orbit as follows: (1) transform the initial conditions to Galactocentric coordinates, where we take the position and velocity of the Sun to be
-\begin{align}
-	(x,y,z)_\odot &= (-8,0,0)~{\rm kpc}\\
-	(v_x,v_y,v_z)_\odot &= (-11.1, 220+24, 7.25)~{\rm km}~{\rm s}^{-1}
-\end{align}
-\citep{8kpc 	ref??, schonrich10, bovy12}, (2) integrate the orbit forward and backward by $t_f$ and $t_b$, respectively, in the potential $\Phi$, (3) transform all orbit points (time-steps) back to observed coordinates, and (4) define interpolating functions for each coordinate as a function of stream longitude, $\phi_1$, using cubic splines---e.g., functions $\widetilde{\phi}_{2}(\phi_1)$, $\widetilde{\DM}(\phi_1)$, $\widetilde{\mu_l}(\phi_1)$, $\widetilde{\mu_b}(\phi_1)$, $\widetilde{v_r}(\phi_1)$.
+For a given set of initial conditions ($\bs{w}_0$), we compute a model orbit as follows: (1) transform the initial conditions to Galactocentric coordinates, (2) integrate the orbit forward and backward by $t_f$ and $t_b$, respectively, in the potential $\Phi$, (3) transform all orbit points (time-steps) back to observed coordinates, and (4) define interpolating functions for each coordinate as a function of stream longitude, $\phi_1$, using cubic splines---e.g., functions $\widetilde{\phi}_{2}(\phi_1)$, $\widetilde{\DM}(\phi_1)$, $\widetilde{\mu_l}(\phi_1)$, $\widetilde{\mu_b}(\phi_1)$, $\widetilde{v_r}(\phi_1)$.
 These functions let us compute the predicted values of each of these coordinates at the longitudes of each observed star, $\phi_{1,i}$.

### 2016-01-14 ms.tex
comments from kathryn and nw figure

@@
 
-\title{Spending too much time at the Galactic bar: Chaotic stream-fanning and the Ophiuchus stream}
+\title{Spending too much time at the Galactic bar: Chaotic fanning of the Ophiuchus stream}
 \author{Adrian M. Price-Whelan\altaffilmark{\colum,\adrn},
@@
 driven by two-body relaxation and gravitational shocks over many Gigayears) with final disruption likely to occur once the tidal boundary is less than the core radius of the cluster.
-The physics of this disruption is not followed exactly but rather the disruption rate and final disruption time are set by \todo{the hypothesis that the most recent shock ...}
+The physics of this disruption is not followed exactly but rather the disruption rate and final disruption time are set by the hypothesis that the most recent combined pericenter and disk shock fully disrupted the cluster, but, critically, that the cluster has been losing debris over its entire orbital history.
 
@@
 	\item The density of the mock stream in the static potential decreases slowly rather than abruptly as is observed.
-This is more clearly seen in terms of the \todo{linear density along the stream, shown in Figure~\ref{fig:lineardensity}}.
-....
-\todo{Waiting on Branimir!}
+This is more clearly seen in terms of the density contrast between stream stars and background stars, visualized in Figure~\ref{fig:densitymaps}: This figure shows mock sky-density maps of stars generated by superimposing the maximum likelihood model streams over a noisy background of stars.
+The number of mock stream star particles used to generate the map has been normalized such that the total number of stars within the observed extent of the stream (between $5.85 > l > 3.81~{\rm deg}$) is equal to the number of stars attributed to the Ophiuchus stream in the PS1 data \citep[$N \approx 500$][]{bernard14}.
 	\item The four high-dispersion BHB stars beyond the end of the stream (from S15b) don't match in position and velocity with the particle distribution from even the maximum likelihood stream model in the static potential.
@@
 
+\todo{Add sentences about Hattori ophiuchus paper}
+
 The model streams presented here do not reproduce all observed features of the Ophiuchus stream.
@@
 
-It is still too early to say for sure that chaotic stream-fanning is occurring for the Ophiuchus stream; deeper follow-up imaging and spectroscopy over a larger area region around the stream will be needed to test the predictions of this work.
+It is still too early to say for sure that chaotic stream-fanning is occurring for the Ophiuchus stream; deeper follow-up imaging and spectroscopy over a larger area region around the stream will be needed to test the predictions of this work and compare with other possible scenarios.
+For example, recent work has shown that if the Ophichus stream progenitor orbit is in resonance with the bar, the debris can remain short for at least 1 Gyr \citep{hattori15}.
+In this model, there would be no nearby, high-dispersion debris, and the pattern speed of the bar would likely be related to the orbital frequencies of the progenitor orbit.
+
  With more information about the density distribution of stars in the stream and better proper motion measurements we would be able to (1) help distinguish models for the Milky Way bar independent from current methods and (2) begin to model the survivability of globular clusters orbiting in the central Milky Way \citep[e.g.,][]{gnedin97}.
@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{figures/mockstream0}
+\includegraphics[width=\textwidth]{figures/mockstream0}
 \caption{ Sky position, distance, and line-of-sight velocity in Heliocentric, Galactic coordinates for star particles (contours or grey points) from the maximum-likelihood mock streams in each potential.
@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{figures/mockstream1}
+\includegraphics[width=\textwidth]{figures/mockstream1}
 \caption{ Same as Figure~\ref{fig:mockstream0} for the other five barred potentials. }
@@
 
-%\begin{figure*}[p]
-%\begin{center}
-%\includegraphics[width=\textwidth]{figures/surface-density}
-%\caption{ Surface density of mock stream star particles in each potential.
-Contours are spaced logarithmically from $10^{-2}$ to $10$ particles per sq.
-deg---that is, each color represents a factor of 10 difference in surface density.
-In the static potential (top left), the density remains high along the center of the stream, but for some of the barred potentials the density drops sharply because of chaotic stream-fanning.}
-%\label{fig:surfacedensity}
-%\end{center}
-%\end{figure*}
+% ---------------------------------------------------------------------------------
+
+\begin{figure*}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/densitymaps}
+\caption{ Simulated maps of the 2D density of star particles from the maximum-likelihood mock streams in each potential with a noisy background of stars binned into 10' by 10' pixels.
+The background star density is assumed to be Poisson with $\lambda = 42$ \citep[see Figure 3 in][where the typical background density is $\approx\frac{60}{(0.2~{\deg})^2}$]{bernard14}.
+The mock stream particles are down-sampled so that the total number of particles in the region of sky that the stream is seen as an over-density matches the observed number of stars \citep[$N\approx500$][]{bernard14}.
+Color-scale is stretched so that white to black is 2nd to 98th percentile.}
+\label{fig:densitymaps}
+\end{center}
+\end{figure*}
 

### 2016-01-14 ms.tex
todo and new refs

@@
 
-\todo{Add sentences about Hattori ophiuchus paper}
-
 The model streams presented here do not reproduce all observed features of the Ophiuchus stream.
@@
 In this model, there would be no nearby, high-dispersion debris, and the pattern speed of the bar would likely be related to the orbital frequencies of the progenitor orbit.
-
+However, it has not been demonstrated whether this proposed scheme can explain the shortness of the stream over timescales closer to the age of the stellar population ($\approx$10 Gyr).
 With more information about the density distribution of stars in the stream and better proper motion measurements we would be able to (1) help distinguish models for the Milky Way bar independent from current methods and (2) begin to model the survivability of globular clusters orbiting in the central Milky Way \citep[e.g.,][]{gnedin97}.
@@
 If the Ophiuchus stream formed from a globular cluster on a chaotic orbit and we happen to be witnessing its final demise, what does this imply about the population of clusters that have already been fully disrupted?
-For a given model for the Galactic bar, we will study the orbit structure in the inner Milky Way and, in particular, study the fraction of strongly chaotic orbits that would lead to chaotic fanning or fast dispersal of tidal debris \citep{recent work by zotos or those greek peeps}.
+For a given model for the Galactic bar, we will study the orbit structure in the inner Milky Way and, in particular, study the fraction of strongly chaotic orbits that would lead to chaotic fanning or fast dispersal of tidal debris \citep[e.g.,][]{jung15}.
 The existence of strongly chaotic orbits in this region would limit the expected number of cold stellar streams in the inner Galaxy and enhance the rate of mixing of the debris.
-Gaia will ....
-compare amount of substructure with expectations .....
-\todo{Quote Gnedin and Ostriker, Murali and Weinberg}
+With constraints on the amount of kinematic substructure in the inner Galaxy from the \project{Gaia} mission \todo{KVJ: my brain is failing - need one sentence here about comparing to bulge formaton scenarios (Gnedin and Ostriker) and predictions for the evolution of globular clusters (Murali and Weinberg) -- any ideas?}
 
 \section{Conclusions}\label{sec:conclusions}
-
-\todo{Predictions and tests of this model!}
 
@@
 APW acknowledges the staff at the MPIA for their support and assistance.
-The authors wish to acknowledge Victor Debattista and Melissa Ness for useful discussions.
+The authors wish to acknowledge Victor Debattista, Melissa Ness, and Sarah Pearson for useful discussions.
 This research made use of Astropy, a community-developed core Python package for Astronomy \citep{astropy13}.

### 2016-01-20 ms.tex
update refs for sesar16

@@
-%\documentclass[numberedappendix]{emulateapj}
-\documentclass[letterpaper,12pt,preprint]{aastex}
+\documentclass[numberedappendix]{emulateapj}
+%\documentclass[letterpaper,12pt,preprint]{aastex}
 
@@
 
-\todo{Still working on the abstract!}
 % Context
 The Ophiuchus stream was recently discovered as an over-density of main sequence stars in Pan-STARRS (PS1) photometry and is peculiar in it's short apparent length given the age of its stellar population.
-The stream was later detected in Blue Horizontal Branch (BHB) stars by identifying members using line-of-sight velocities: Twenty BHB stars in the vicinity of the stream are kinematically distinct from the background halo stars and have distances consistent with the previously detected segment from PS1.
+The stream was later detected in Blue Horizontal Branch (BHB) stars using line-of-sight velocities: Twenty BHB stars in the vicinity of the stream are kinematically distinct from the background halo stars and have distances consistent with the previously detected segment from PS1.
 Four of these BHB stars beyond the apparent end of PS1 stars have similarly distinct velocities but have much larger velocity and spatial dispersions. % compared to the PS1 stream segment.
-In recent work, we have shown that streams formed on chaotic orbits may display similar peculiar morphologies: short segments of coherent stellar debris flanked by high-dispersion, low-density ``stream-fans'' at the ends of the stream.
+In recent work, we have shown that streams formed on chaotic orbits may display similar peculiar morphologies: Short segments of coherent stellar debris flanked by high-dispersion, low-density ``stream-fans'' at the ends of the stream.
 % Aims
@@
 % Methods
-[We fit orbits to the debris in one static and several barred potential models ...]
-We then assess the level of chaos for orbits like that of the Ophiuchus stream and generate mock streams along these orbits that qualitatively reproduce the observed kinematics of the stream.
+We fit orbits to the observed stream stars in on static and several barred models for the gravitational potential of the Milky Way. 
+We generate model stellar streams along samples from the orbit distributions and compare the maximum-likelihood model streams with the observed density and kinematics of the stream stars.
 % Results
-With the presence of a bar, many orbits near ... to that of are chaotic ...
-We find that the apparent shortness and nearby low-density, high-velocity-dispersion can be qualitatively explained if the stream formed along a strongly chaotic orbit.
+With the presence of a bar, many orbits consistent with the stream are strongly chaotic and produce model streams that qualitatively reproduce the apparent shortness and nearby low-density, high-velocity-dispersion stream stars.
 % Conclusions
-These results motivate efforts to obtain deep photometry of the region surrounding the Ophiuchus stream to detect the predicted low-surface-brightness fans of stellar debris.
-The existence of or lack of stream fans may provide an interesting constraint on the triaxiality and pattern speed of the bar.
+These results motivate efforts to obtain deep photometry of the region surrounding the Ophiuchus stream to detect the predicted low-surface-brightness fans of stellar debris and help distinguish between formation scenarios for the stream.
+The existence of or lack of stream-fans may provide an interesting constraint on the triaxiality and pattern speed of the bar.
 
@@
 The Ophiuchus stream \citep{bernard14, sesar15a} is a recently discovered stellar tidal stream that sits above the Galactic bulge at a Galactocentric radius and height $(R,z) \approx (1.5, 4.3)~{\rm kpc}$.
-All observational evidence suggests that the stream is a totally disrupted globular cluster: The stream stars have (1) a small positional dispersion orthogonal to the extended direction of the stream (width $\approx$10 pc, length $\approx$1.5 kpc); (2) no detectable over-density along the stream that could be the progenitor system; (3) a small velocity dispersion $\approx$0.4 ${\rm km}~{\rm s}^{-1}$; and (4) an old stellar population ($\approx$12 Gyr) estimated from isochrone fitting \citep[][hereafter S15a]{sesar15a}.
+All observational evidence suggests that the stream is a totally disrupted globular cluster: The stream stars have (1) a small positional dispersion orthogonal to the extended direction of the stream (width $\approx$10 pc, length $\approx$1.5 kpc); (2) no detectable over-density along the stream that could be the progenitor system; (3) a small velocity dispersion $\approx$0.4 ${\rm km}~{\rm s}^{-1}$; and (4) an old stellar population ($\approx$12 Gyr) estimated from isochrone fitting \citep[][hereafter S15]{sesar15a}.
 
@@
 For example, the de-projected length of the visible part of the stream is short given the age of its stellar population ($\approx$1.5 kpc).
-S15a fit an orbit to the kinematics of the stream stars in a static, axisymmetric model for the gravitational field of the inner Galaxy and ran N-body simulations of globular clusters on this orbit.
-S15a find that---on this orbit---the portion of the stream visible as an over-density in main-sequence stars must have been formed in the last $\lesssim$400 Myr for the stream to remain as short as it is observed.
+S15 fit an orbit to the kinematics of the stream stars in a static, axisymmetric model for the gravitational field of the inner Galaxy and ran N-body simulations of globular clusters on this orbit.
+S15 find that---on this orbit---the portion of the stream visible as an over-density in main-sequence stars must have been formed in the last $\lesssim$400 Myr for the stream to remain as short as it is observed.
 This dynamical age is at odds with the old stellar population: The stars appear to have formed $\approx$12 Gyr ago, but the cluster apparently fully disrupted at once in the last 400 Myrs.
-Another puzzle is the existence of blue horizontal branch (BHB) stars close to the stream (within a few degrees) with similar radial velocities, but with a large dispersion in both sky position and velocity \citep[][hereafter S15b]{sesar15b}.
+Another puzzle is the existence of blue horizontal branch (BHB) stars close to the stream (within a few degrees) with similar radial velocities, but with a large dispersion in both sky position and velocity \citep[][hereafter S16]{sesar16}.
 The stream has a very distinct and large line-of-sight velocity ($\approx$290 \kms) and is therefore easily detected above the background halo population.
 Four BHB stars have been detected with line-of-sight velocities $>230~\kms$ that lie close to an extrapolation of the stream on the sky, but have a velocity dispersion $\approx$75 times larger than the measured internal velocity dispersion of the stream stars.
-The velocities of these stars are highly inconsistent with belonging to the background halo population (see Section 4, S15b).
-These stars hint at the existence of associated low-density, high-dispersion features that were not modeled in S15a and are not predicted by the $N$-body simulations from this prior work.
+The velocities of these stars are highly inconsistent with belonging to the background halo population (see Section 4, S16).
+These stars hint at the existence of associated low-density, high-dispersion features that were not modeled in S15 and are not predicted by the $N$-body simulations from this prior work.
 
@@
 
-The orbit fit and $N$-body simulations in S15a used a static, axisymmetric potential to represent the Milky Way potential, but it is well-known that the Galactic bulge contains a triaxial, bar-like structure several kpc in size \citep[e.g.,][]{blitz91, weinberg92, dwek95, wegg13}.
+The orbit fit and $N$-body simulations in S15 used a static, axisymmetric potential to represent the Milky Way potential, but it is well-known that the Galactic bulge contains a triaxial, bar-like structure several kpc in size \citep[e.g.,][]{blitz91, weinberg92, dwek95, wegg13}.
 Given the proximity of the stream to the center of the Galaxy, the time-dependent, triaxial potential of the Galactic bar must be taken into account when modeling the orbit of the Ophiuchus stream.
@@
 	\item the apparent shortness and fast density truncation of the stream;
-	\item the increased positional dispersion of the four new candidate members from S15b;
-	\item the large velocity dispersion of the S15b stars.
+	\item the increased positional dispersion of the four new candidate members from S16;
+	\item the large velocity dispersion of the S16 stars.
 \end{enumerate}
@@
 
-In each of the potentials described above, we fit orbits to the measured kinematics of BHB stars that are high-likelihood members of the Ophiuchus stream \citep{sesar15a, sesar15b}.
+In each of the potentials described above, we fit orbits to the measured kinematics of BHB stars that are high-likelihood members of the Ophiuchus stream \citep{sesar15a, sesar16}.
 The details of this procedure and a definition of the likelihood function we use are presented in Appendix~\ref{sec:orbitfit}.
@@
 Data used for computing the likelihood are shown as black points with grey error bars.
-The four ``fanned'' BHB stars from S15b excluded from the likelihood computation are shown as grey squares.
+The four ``fanned'' BHB stars from S16 excluded from the likelihood computation are shown as grey squares.
 Error bars may sometimes be smaller than the point size.
@@
 We randomly sample 256 orbits from the orbital parameter posterior distributions and generate mock streams along each orbit.
-We use the method outlined above (Section \ref{sec:mocks}) to generate the streams and set the free parameters as follows: (1) we evolve the progenitors for 6 Gyr along each orbit to explore stream models where the shortness of the stream is \emph{not} due to an instantaneous disruption 400 Myrs ago, (2) we release star particles every 0.5 Myr (uniformly in time) to densely sample the final density distribution, (3) we set the progenitor mass to $m=10^4~\msun$ (as was estimated by S15a), and (4) we set the disruption time of the progenitor equal to the last time at which a pericenter and disk crossing coincide (in each case, this is at $t \approx -200~{\rm Myr}$).
+We use the method outlined above (Section \ref{sec:mocks}) to generate the streams and set the free parameters as follows: (1) we evolve the progenitors for 6 Gyr along each orbit to explore stream models where the shortness of the stream is \emph{not} due to an instantaneous disruption 400 Myrs ago, (2) we release star particles every 0.5 Myr (uniformly in time) to densely sample the final density distribution, (3) we set the progenitor mass to $m=10^4~\msun$ (as was estimated by S15), and (4) we set the disruption time of the progenitor equal to the last time at which a pericenter and disk crossing coincide (in each case, this is at $t \approx -200~{\rm Myr}$).
 After the disruption time, we continue releasing the star particles uniformly in time rather than releasing a ``burst'' of particles at once.
@@
 
-For each generated mock stream, we compute the likelihood of the data (now including all BHB stars from S15a and S15b, e.g., all points in Figure~\ref{fig:orbitfits}) given the star particles by estimating the phase-space model density using a kernel density estimate with a Gaussian kernel \citep[see, e.g.,][]{bonaca14}.
+For each generated mock stream, we compute the likelihood of the data (now including all BHB stars from S15 and S16, e.g., all points in Figure~\ref{fig:orbitfits}) given the star particles by estimating the phase-space model density using a kernel density estimate with a Gaussian kernel \citep[see, e.g.,][]{bonaca14}.
 For each $i$ data point $\bs{x}_i$ and each $k$ model point with coordinates $\bs{y}_k$, we compute the likelihood by converting the model point position and velocity into Heliocentric coordinates and evaluate
@@
 Figures~\ref{fig:mockstream0}--\ref{fig:mockstream1} show the final particle positions and line-of-sight velocities in heliocentric coordinates for the maximum likelihood mock streams (grey points) in each potential.
-Vertical, dashed lines show the approximate extent of the part of the stream visible in main-sequence stars (excluding the BHB stars from S15b).
+Vertical, dashed lines show the approximate extent of the part of the stream visible in main-sequence stars (excluding the BHB stars from S16).
 There are a few interesting features to note from these panels:
@@
 The number of mock stream star particles used to generate the map has been normalized such that the total number of stars within the observed extent of the stream (between $5.85 > l > 3.81~{\rm deg}$) is equal to the number of stars attributed to the Ophiuchus stream in the PS1 data \citep[$N \approx 500$][]{bernard14}.
-	\item The four high-dispersion BHB stars beyond the end of the stream (from S15b) don't match in position and velocity with the particle distribution from even the maximum likelihood stream model in the static potential.
+	\item The four high-dispersion BHB stars beyond the end of the stream (from S16) don't match in position and velocity with the particle distribution from even the maximum likelihood stream model in the static potential.
 However, in some barred potentials the chaotic evolution of the stream stars can lead to over-densities of stars with an increased positional dispersion and significantly discrepant velocities (e.g., bar8).
-	\item None of the stream models produce an appreciable density of stars with line-of-sight velocities near the S15b BHB star with the largest velocity ($\approx 320~\kms$).
+	\item None of the stream models produce an appreciable density of stars with line-of-sight velocities near the S16 BHB star with the largest velocity ($\approx 320~\kms$).
 This star is either an interesting Ophiuchus member star or is associated with some other kinematic substructure.
@@
 
-%We have applied the same distance-modulus cut to the points in these figures as used to select the BHB stars identified as stream members by S15b (see Section~2 in S15b).
+%We have applied the same distance-modulus cut to the points in these figures as used to select the BHB stars identified as stream members by S16 (see Section~2 in S16).
 
@@
 In this notation, $(l, b)$ are Galactic coordinates, $\DM$ is the distance modulus, $(\mu_l, \mu_b)$ are proper motions in the Galactic frame, and $v_r$ is the radial velocity.
-We assume that the sky coordinates for each star are known perfectly well (have zero uncertainty) and transform the data to a rotated, heliocentric coordinate system that is aligned with the stream and centered on the median sky position of the BHB stars in the densest part of the stream \cite[all BHB stars except the `fanned' stars: cand15, cand26, cand49, cand54 from][]{sesar15b}.
+We assume that the sky coordinates for each star are known perfectly well (have zero uncertainty) and transform the data to a rotated, heliocentric coordinate system that is aligned with the stream and centered on the median sky position of the BHB stars in the densest part of the stream \cite[all BHB stars except the `fanned' stars: cand15, cand26, cand49, cand54 from][]{sesar16}.
 We represent the longitude and latitude in these coordinates as $(\phi_1, \phi_2)$ and the rotation matrix to transform from Galactic to these coordinates is given in Appendix~\ref{sec:rotationmatrix}.

### 2016-01-22 ms.tex
brani and kathryn comments and abstract

@@
 
-\title{Spending too much time at the Galactic bar: Chaotic fanning of the Ophiuchus stream}
+\title{Spending too much time at the Galactic bar: chaotic fanning of the Ophiuchus stream}
 \author{Adrian M. Price-Whelan\altaffilmark{\colum,\adrn},
@@
 \begin{abstract}
-
-% Context
-The Ophiuchus stream was recently discovered as an over-density of main sequence stars in Pan-STARRS (PS1) photometry and is peculiar in it's short apparent length given the age of its stellar population.
-The stream was later detected in Blue Horizontal Branch (BHB) stars using line-of-sight velocities: Twenty BHB stars in the vicinity of the stream are kinematically distinct from the background halo stars and have distances consistent with the previously detected segment from PS1.
-Four of these BHB stars beyond the apparent end of PS1 stars have similarly distinct velocities but have much larger velocity and spatial dispersions. % compared to the PS1 stream segment.
-In recent work, we have shown that streams formed on chaotic orbits may display similar peculiar morphologies: Short segments of coherent stellar debris flanked by high-dispersion, low-density ``stream-fans'' at the ends of the stream.
-% Aims
-The stream's proximity to the triaxial, time-dependent potential of the Galactic bar and the peculiar kinematics and morphology of the stream suggest that it may have formed on a chaotic orbit.
-Here we study whether the presence of a triaxial, time-dependent bar can lead to chaotic orbits near (in phase-space) to the Ophiuchus stream. 
-% Methods
-We fit orbits to the observed stream stars in on static and several barred models for the gravitational potential of the Milky Way. 
-We generate model stellar streams along samples from the orbit distributions and compare the maximum-likelihood model streams with the observed density and kinematics of the stream stars.
-% Results
-With the presence of a bar, many orbits consistent with the stream are strongly chaotic and produce model streams that qualitatively reproduce the apparent shortness and nearby low-density, high-velocity-dispersion stream stars.
-% Conclusions
-These results motivate efforts to obtain deep photometry of the region surrounding the Ophiuchus stream to detect the predicted low-surface-brightness fans of stellar debris and help distinguish between formation scenarios for the stream.
-The existence of or lack of stream-fans may provide an interesting constraint on the triaxiality and pattern speed of the bar.
-
+The Ophiuchus stellar stream is peculiar: (1) its length is short given the age of its constituent stars, and (2) several probable member stars that lie close in both sky position and velocity have dispersions in these dimensions that far exceed those seen within the stream.
+The stream's proximity to the Galactic center suggests that the bar should have a significant influence on its dynamical history: The triaxiality and time-dependence of the bar may generate chaotic orbits in the vicinity of the stream that can greatly affect its morphology.
+We explore this hypothesis with models of stream formation along orbits consistent with Ophiuchus' properties in a Milky Way potential model that includes a rotating bar.
+We find that in all choices for the rotation parameters of the bar, orbits fit to the stream are strongly chaotic.
+Mock streams generated along these orbits qualitatively match the observed properties of the stream: because of chaos, stars stripped early generally form low-density, high-dispersion ``fans'' leaving only the most recently disrupted material detectable as a strong over-density.
+These results motivate efforts to obtain deep photometry and spectroscopy of the region surrounding the stream to detect this predicted low-surface-brightness tidal debris.
+The existence of or lack of these features would provide an interesting constraint on the properties of the Milky Way bar and would help distinguish between formation scenarios for the stream.
+This is the first time that chaos has been used to explain the properties of a stellar stream and is the first demonstration of the dynamical importance of chaos in the Galactic halo.
 \end{abstract}
@@
 
-There are a number of peculiarities about the observed kinematics of probable member stars of the Ophiuchus stream.
+There are a number of peculiarities about the observed kinematics of the Ophiuchus stream.
 For example, the de-projected length of the visible part of the stream is short given the age of its stellar population ($\approx$1.5 kpc).
@@
 S15 find that---on this orbit---the portion of the stream visible as an over-density in main-sequence stars must have been formed in the last $\lesssim$400 Myr for the stream to remain as short as it is observed.
-This dynamical age is at odds with the old stellar population: The stars appear to have formed $\approx$12 Gyr ago, but the cluster apparently fully disrupted at once in the last 400 Myrs.
+This dynamical age is at odds with the old stellar population: The stars appear to have formed $\approx$12 Gyr ago, but the abrupt end of the stream suggests that the cluster apparently fully disrupted at once in the last 400 Myrs.
 Another puzzle is the existence of blue horizontal branch (BHB) stars close to the stream (within a few degrees) with similar radial velocities, but with a large dispersion in both sky position and velocity \citep[][hereafter S16]{sesar16}.
@@
 The velocities of these stars are highly inconsistent with belonging to the background halo population (see Section 4, S16).
-These stars hint at the existence of associated low-density, high-dispersion features that were not modeled in S15 and are not predicted by the $N$-body simulations from this prior work.
+These stars hint at the existence of associated low-density, high-dispersion features that were not modeled in S15 and are not predicted by the $N$-body simulations from this prior work, which assume a sudden, total disruption of the stream progenitor.
 
@@
 
-The orbit fit and $N$-body simulations in S15 used a static, axisymmetric potential to represent the Milky Way potential, but it is well-known that the Galactic bulge contains a triaxial, bar-like structure several kpc in size \citep[e.g.,][]{blitz91, weinberg92, dwek95, wegg13}.
+The orbit fit and $N$-body simulations in S15 used a static, axisymmetric potential to represent the Milky Way potential, but it is well-known that the Galactic bulge contains a triaxial, rotating, bar-like structure several kpc in size \citep[e.g.,][]{blitz91, weinberg92, dwek95, wegg13}.
 Given the proximity of the stream to the center of the Galaxy, the time-dependent, triaxial potential of the Galactic bar must be taken into account when modeling the orbit of the Ophiuchus stream.
 The presence of a bar-like perturbation to the potential will change the orbit of the stream progenitor and the orbit structure in the inner galaxy \citep{zotos12, portail15b, gajda15}.
-Bar-like features can also introduce a significant number of chaotic orbits in their vicinity \citep{weinberg15}.
+Bar-like features can also introduce a significant number of chaotic orbits in their vicinity \citep{weinberg15} and generate resonances that may also affect stream formation \citep{hattori15}.
 
@@
 Our goal is not to perfectly represent the observed data but rather to explore the plausibility of explaining the peculiarities of the stream using chaotic stream fanning.
+Note that an alternate model was recently proposed that instead places the stream progenitor on an orbit in resonance with the bar.
+We discuss the differences between these two models in Section~\ref{sec:discussion}.
 
-In Section~\ref{sec:method} we describe the methods used in this work: in Section~\ref{sec:potential} we describe the models we use for the gravitational potential of the Galaxy, in Section~\ref{sec:mocks} we explain the simple method we use to generate mock streams, and in Section~\ref{sec:chaos-indicator} we briefly describe the methods we use for detecting and characterizing chaotic orbits.
-In Section~\ref{sec:results1} we discuss the differences in the orbit structure between a static, axisymmetric potential model and a model with a time-dependent bar potential, and in Section~\ref{sec:results2} we use mock streams to argue that [chaotic stream-fanning is a plausible explanation for the observational peculiarities of the Ophiuchus stream.] We conclude in Section~\ref{sec:conclusions}.
+In Section~\ref{sec:method} we describe the methods used in this work: in Section~\ref{sec:potential} we describe the models we use for the gravitational potential of the Galaxy, in Section~\ref{sec:orbitfit-nonapdx} we outline the probabilistic procedure we use to fit orbits to the stream data (explained in detail in Appendix~\ref{sec:orbitfit}), and in Section~\ref{sec:mocks} we explain the simple method we use to generate mock streams.
+In Section~\ref{sec:results1} we discuss the results from fitting orbits to the data in a static, axisymmetric potential model and several potential models with a time-dependent bar.
+In Section~\ref{sec:results2} we generate mock streams along these orbits to argue that chaotic stream-fanning is a plausible explanation for the observational peculiarities of the Ophiuchus stream.
+We discuss the implications of this work and possibilities for future work in Section~\ref{sec:discussion} and we conclude in Section~\ref{sec:conclusions}.
 
@@
 Our goal is to (1) assess whether the Galactic bar can produce chaotic orbits in the vicinity of the Ophiuchus stream and (2) determine if chaotic density evolution of tidal debris stripped from the progenitor of this stream can explain the apparent shortness of the stream and low-density, high-dispersion stars beyond the extent of main sequence stars observed in PS1.
-In this section, we describe the potential models we use to represent the galaxy and describe the methods we use to detect and quantify the strength of chaos for orbits.
-We then describe the method we use for fitting orbits to the stream stars.
-Finally, we describe the method we use to generate mock stellar streams.
+In this section, we describe the potential models we use to represent the galaxy and outline the methods we use to detect and quantify the strength of chaos for individual orbits.
+We then describe the likelihood function we use for fitting orbits to the stream stars.
+Finally, we describe how we generate mock stellar streams for a progenitor on a given orbit.
 
-In this work, we assume the Sun is at Galactocentric position $(x,y,z) = (-8.3,0,0)~{\rm kpc}$ \citep[e.g.,][]{schoenrich12} with velocity $(v_x,v_y,v_z) = (-11.1, 250, 7.25)~\kms$ \citep[e.g.,][]{schoenrich10, schoenrich12}.
+Throughout we assume the Sun is at Galactocentric position $(x,y,z) = (-8.3,0,0)~{\rm kpc}$ \citep[e.g.,][]{schoenrich12} with velocity $(v_x,v_y,v_z) = (-11.1, 250, 7.25)~\kms$ \citep[e.g.,][]{schoenrich10, schoenrich12}.
 
@@
 
-The BFE representation fixes the axis ratios of the bar---that is, the the exponential scale lengths along the three axes of the bar were adopted from \cite{dwek95} when the expansion coefficients were calculated in W12); all other potential parameter values are given in Table~\ref{tbl:potential-params-barred}.
+The BFE representation fixes the axis ratios of the bar---that is, the exponential scale lengths along the three axes of the bar were adopted from \cite{dwek95} when the expansion coefficients were calculated in W12; all other potential parameter values are given in Table~\ref{tbl:potential-params-barred}.
 The mass of the halo is fixed and the mass of the disk and bar are varied in order to qualitatively reproduce the flatness and amplitude of the circular velocity curve of the Milky Way \citep{bovy12}.
@@
 We consider a grid of nine parameter combinations of bar angle and pattern speed.
-These model names and parameter values are given in Table~\ref{tbl:bar-specific}.
+Model names and parameter values are given in Table~\ref{tbl:bar-specific}.
 
@@
 
-\subsection{Fitting orbits to the Ophiuchus stream}
+\subsection{Fitting orbits to the Ophiuchus stream}\label{sec:orbitfit-nonapdx}
 
@@
 
-To generate mock stellar streams, we use the method presented in \citep{fardal14}: star particles are `released' from a progenitor system near the Lagrange points with a dispersion in position and velocity that is set by the mass and orbit of the progenitor.
+To generate mock stellar streams, we use the method presented in \citet{fardal14}: Star particles are `released' from a progenitor system near the Lagrange points with a dispersion in position and velocity that is set by the mass and orbit of the progenitor.
 We draw samples from the posterior probability distributions over orbital parameters from fitting orbits to the stream star members and use these as the progenitor orbital parameters (Section~\ref{sec:orbitfit}).
@@
 Figure~\ref{fig:orbitfits} summarizes our results from fitting orbits to the BHB stream stars.
-Shown in each panel are the Ophiuchus stream star members (black points) and orbits integrated from samples from the posterior probability over orbital parameters (blue lines).
+Shown in each panel are the high-probability Ophiuchus stream stars (black points, to which orbits are fit) and orbits integrated from samples from the posterior probability over orbital parameters (blue lines).
+The four ``fanned'' BHB stars from S16 are shown as grey squares and are not included in the orbit fitting procedure.
 We only show one of the barred potentials: The end-to-end integration time of the orbit over the observed extent of the stream is only $\approx$6 Myr, so the derived orbits are extremely similar in all potentials (the time-dependence of the bar potential is not significant over such short timescales).
@@
 Orbits are integrated for 6 Gyr and shown in Galactocentric, Cartesian coordinates.
-Even though the mean orbital parameters have nearly identical values (e.g., the initial conditions are nearly identical in Heliocentric coordinates), the orbits in each potential model are  different in appearance.
+Even though the mean orbital parameters have nearly identical values (e.g., the initial conditions are nearly identical in heliocentric coordinates), the orbits in each potential model are  different in appearance.
 }
@@
 \caption{ Histograms of estimates of the Lyapunov time, $\lyapt=1/\lyapexp$, for posterior samples from the orbit fitting procedure (Section~\ref{sec:orbitfit}).
-Vertical lines show the Lyapunov time value of the mean orbit in that potential model.
 Orbits are typically more chaotic (have shorter Lyapunov times) for higher pattern speeds.
@@
 
-\section{Results 2: Stream models for the Ophiuchus stream}
+\section{Results 2: Stream models for the Ophiuchus stream}\label{sec:results2}
 
@@
 
-We randomly sample 256 orbits from the orbital parameter posterior distributions and generate mock streams along each orbit.
-We use the method outlined above (Section \ref{sec:mocks}) to generate the streams and set the free parameters as follows: (1) we evolve the progenitors for 6 Gyr along each orbit to explore stream models where the shortness of the stream is \emph{not} due to an instantaneous disruption 400 Myrs ago, (2) we release star particles every 0.5 Myr (uniformly in time) to densely sample the final density distribution, (3) we set the progenitor mass to $m=10^4~\msun$ (as was estimated by S15), and (4) we set the disruption time of the progenitor equal to the last time at which a pericenter and disk crossing coincide (in each case, this is at $t \approx -200~{\rm Myr}$).
-After the disruption time, we continue releasing the star particles uniformly in time rather than releasing a ``burst'' of particles at once.
+For each potential model, we randomly sample 256 orbits from the orbital parameter posterior distributions and generate mock streams along each orbit.
+We use the method outlined above (Section \ref{sec:mocks}) to generate the streams and set the free parameters as follows: (1) we evolve the progenitors for 6 Gyr along each orbit prior to the current position to explore stream models where the shortness of the stream is \emph{not} due to an instantaneous disruption 400 Myrs ago, (2) we release star particles every 0.5 Myr (uniformly in time) to densely sample the final density distribution, (3) we set the progenitor mass to $m=10^4~\msun$ (as was estimated by S15), and (4) we set the disruption time of the progenitor equal to the last time at which a pericenter and disk crossing coincide (in each case, this is at $t \approx -200~{\rm Myr}$).
+After the disruption time, we continue releasing the star particles uniformly in time (every 0.5 Myr) rather than releasing a ``burst'' of particles at once.
 We therefore expect that the density of the most recently disrupted debris will be systematically higher for the model streams as compared to the observed stream.
@@
 	\end{tabular}
-	\caption{Estimated mean and standard deviation of samples from the marginal posterior distributions over each parameter in our orbit fit model.
+	\caption{Estimated mean and standard deviation of samples from the marginal posterior distributions over each parameter in our orbit fit model (the posterior distributions are very close to Gaussian).
 For the barred potentials, all mean values are the same because the time-dependence of the bar doesn't impact the orbit fit over the short length of the stream.
@@
 For each generated mock stream, we compute the likelihood of the data (now including all BHB stars from S15 and S16, e.g., all points in Figure~\ref{fig:orbitfits}) given the star particles by estimating the phase-space model density using a kernel density estimate with a Gaussian kernel \citep[see, e.g.,][]{bonaca14}.
-For each $i$ data point $\bs{x}_i$ and each $k$ model point with coordinates $\bs{y}_k$, we compute the likelihood by converting the model point position and velocity into Heliocentric coordinates and evaluate
+For each $i$ data point $\bs{x}_i$ and each $k$ model point with coordinates $\bs{y}_k$, we compute the likelihood by converting the model point position and velocity into heliocentric coordinates and evaluate
 \begin{align}
@@
 This is a projection effect: The portion of the stream at larger $l$ is closer points almost directly towards the Sun so that the debris covers a larger area.
-This is more clearly seen in Figure~\ref{fig:mockstreamxyz}, which shows $x$-$z$ projections of the star particles in Galactocentric Cartesian coordinates (grey) along with the position of the Sun (symbol at $(x,z)=(-8.3,0)~{\rm kpc}$) and the ``window'' of the Heliocentric, sky-position plots of Figures~\ref{fig:mockstream0}--\ref{fig:mockstream1} (shown as blue lines).
 	\item The density of the mock stream in the static potential decreases slowly rather than abruptly as is observed.
-This is more clearly seen in terms of the density contrast between stream stars and background stars, visualized in Figure~\ref{fig:densitymaps}: This figure shows mock sky-density maps of stars generated by superimposing the maximum likelihood model streams over a noisy background of stars.
-The number of mock stream star particles used to generate the map has been normalized such that the total number of stars within the observed extent of the stream (between $5.85 > l > 3.81~{\rm deg}$) is equal to the number of stars attributed to the Ophiuchus stream in the PS1 data \citep[$N \approx 500$][]{bernard14}.
+This is shown in the top panels of each column where each contour level represents a factor of 10 difference in projected surface density.
+In the static potential model the length of dense debris extends much farther than the observed extent of the stream (vertical dashed lines), whereas in some of the barred potential models the stars released earlier have ``fanned'' and are associated with much lower density debris (e.g., bar8).
 	\item The four high-dispersion BHB stars beyond the end of the stream (from S16) don't match in position and velocity with the particle distribution from even the maximum likelihood stream model in the static potential.
 However, in some barred potentials the chaotic evolution of the stream stars can lead to over-densities of stars with an increased positional dispersion and significantly discrepant velocities (e.g., bar8).
-	\item None of the stream models produce an appreciable density of stars with line-of-sight velocities near the S16 BHB star with the largest velocity ($\approx 320~\kms$).
+	\item None of the stream models---static or barred---produce an appreciable density of stars with line-of-sight velocities near the S16 BHB star with the largest velocity ($\approx 320~\kms$).
 This star is either an interesting Ophiuchus member star or is associated with some other kinematic substructure.
 	\item For the barred potentials, the stream morphology is very sensitive to the properties of the bar (especially the pattern speed) and to the initial conditions of the orbit.
-We have found that the morphology can vary significantly between nearby orbits in the same potential model (because these are strongly chaotic orbits), but the overall characteristics remain similar: along more strongly chaotic orbits, the debris ``fans'' more and the apparent dense part of the model streams is shorter (see, for example, Figure~\ref{fig:mockstreamxyz}, panels bar3 vs.
-bar8).
+We have found that the morphology can vary significantly between nearby orbits in the same potential model (because these are strongly chaotic orbits), but the overall characteristics remain similar: along more strongly chaotic orbits, the debris ``fans'' more and the apparent dense part of the model streams is shorter.
 \end{enumerate}
+
+The viewing angle and stream geometry is more clearly demonstrated in Figure~\ref{fig:mockstreamxyz}, which shows $x$-$z$ projections of the star particles in Galactocentric Cartesian coordinates (grey) along with the position of the Sun (symbol at $(x,z)=(-8.3,0)~{\rm kpc}$) and the ``window'' of the heliocentric, sky-position plots of Figures~\ref{fig:mockstream0}--\ref{fig:mockstream1} (shown as blue lines).
+
+The density truncation of the mock streams in each potential model is more clearly seen in terms of the density contrast between stream stars and background stars, visualized in Figure~\ref{fig:densitymaps}: This figure shows mock sky-density maps of stars generated by superimposing the maximum likelihood model streams over a noisy background of stars.
+The number of mock stream star particles used to generate the map has been normalized such that the total number of stars within the observed extent of the stream (between $5.85 > l > 3.81~{\rm deg}$) is equal to the number of stars attributed to the Ophiuchus stream in the PS1 data \citep[$N \approx 500$][]{bernard14}.
 
@@
 
-It is still too early to say for sure that chaotic stream-fanning is occurring for the Ophiuchus stream; deeper follow-up imaging and spectroscopy over a larger area region around the stream will be needed to test the predictions of this work and compare with other possible scenarios.
+It is still too early to say for sure that chaotic stream-fanning is occurring for the Ophiuchus stream.
+Deeper follow-up imaging and spectroscopy over a larger area region around the stream will be needed to test the predictions of this work and compare with other possible scenarios.
 For example, recent work has shown that if the Ophichus stream progenitor orbit is in resonance with the bar, the debris can remain short for at least 1 Gyr \citep{hattori15}.
@@
 If the Ophiuchus stream formed from a globular cluster on a chaotic orbit and we happen to be witnessing its final demise, what does this imply about the population of clusters that have already been fully disrupted?
-For a given model for the Galactic bar, we will study the orbit structure in the inner Milky Way and, in particular, study the fraction of strongly chaotic orbits that would lead to chaotic fanning or fast dispersal of tidal debris \citep[e.g.,][]{jung15}.
 The existence of strongly chaotic orbits in this region would limit the expected number of cold stellar streams in the inner Galaxy and enhance the rate of mixing of the debris.
-With constraints on the amount of kinematic substructure in the inner Galaxy from the \project{Gaia} mission \todo{KVJ: my brain is failing - need one sentence here about comparing to bulge formaton scenarios (Gnedin and Ostriker) and predictions for the evolution of globular clusters (Murali and Weinberg) -- any ideas?}
+The fraction of strongly chaotic orbits that would lead to chaotic fanning or fast dispersal of tidal debris should therefore be related to the amount of kinematic substructure in the inner Galaxy.
+Indeed, first suggestions of kinematic substructure in the bulge have been found, but further modeling is needed to understand whether these hints could be signature of a widely dispersed globular cluster population.
+If so, a stronger theoretical understanding of the prevalence of these features could be combined with future kinematic surveys (from, e.g., \project{Gaia}) to place constraints on long-standing puzzles about the primordial globular cluster population \citep[e.g.,][]{murali97, gnedin97}.
 
@@
 \includegraphics[width=\textwidth]{figures/mockstream0}
-\caption{ Sky position, distance, and line-of-sight velocity in Heliocentric, Galactic coordinates for star particles (contours or grey points) from the maximum-likelihood mock streams in each potential.
+\caption{ Sky position, distance, and line-of-sight velocity in heliocentric, Galactic coordinates for star particles (contours or grey points) from the maximum-likelihood mock streams in each potential.
 Top panels show surface density of mock stream star particles in each potential.
@@
 \bibliographystyle{apj}
-\bibliography{refs}
+\bibliography{paper-helpers/references/refs}
 

### 2016-01-24 ms.tex
pre-rix comments

@@
 Mock streams generated along these orbits qualitatively match the observed properties of the stream: because of chaos, stars stripped early generally form low-density, high-dispersion ``fans'' leaving only the most recently disrupted material detectable as a strong over-density.
-These results motivate efforts to obtain deep photometry and spectroscopy of the region surrounding the stream to detect this predicted low-surface-brightness tidal debris.
-The existence of or lack of these features would provide an interesting constraint on the properties of the Milky Way bar and would help distinguish between formation scenarios for the stream.
+These results motivate efforts to obtain deep photometry and spectroscopy of the region surrounding the stream. 
+The existence of or lack of low-surface-brightness tidal debris would provide an interesting constraint on the properties of the Milky Way bar and would help distinguish between formation scenarios for the stream.
 This is the first time that chaos has been used to explain the properties of a stellar stream and is the first demonstration of the dynamical importance of chaos in the Galactic halo.
@@
 	\item Even in the static potential (Figure~\ref{fig:mockstream0}, leftmost column), there is a slight decrease in the density for the model stream towards higher Galactic longitudes, $l$.
-This is a projection effect: The portion of the stream at larger $l$ is closer points almost directly towards the Sun so that the debris covers a larger area.
+This is a projection effect: The portion of the stream at larger $l$ is closer and points almost directly towards the Sun so that the debris covers a larger area on the sky.
 	\item The density of the mock stream in the static potential decreases slowly rather than abruptly as is observed.
@@
 	\item The four high-dispersion BHB stars beyond the end of the stream (from S16) don't match in position and velocity with the particle distribution from even the maximum likelihood stream model in the static potential.
-However, in some barred potentials the chaotic evolution of the stream stars can lead to over-densities of stars with an increased positional dispersion and significantly discrepant velocities (e.g., bar8).
+In some barred potentials the chaotic evolution of the stream stars can lead to over-densities of stars with an increased positional dispersion and significantly discrepant velocities (e.g., bar8).
 	\item None of the stream models---static or barred---produce an appreciable density of stars with line-of-sight velocities near the S16 BHB star with the largest velocity ($\approx 320~\kms$).
@@
 
-The viewing angle and stream geometry is more clearly demonstrated in Figure~\ref{fig:mockstreamxyz}, which shows $x$-$z$ projections of the star particles in Galactocentric Cartesian coordinates (grey) along with the position of the Sun (symbol at $(x,z)=(-8.3,0)~{\rm kpc}$) and the ``window'' of the heliocentric, sky-position plots of Figures~\ref{fig:mockstream0}--\ref{fig:mockstream1} (shown as blue lines).
-
 The density truncation of the mock streams in each potential model is more clearly seen in terms of the density contrast between stream stars and background stars, visualized in Figure~\ref{fig:densitymaps}: This figure shows mock sky-density maps of stars generated by superimposing the maximum likelihood model streams over a noisy background of stars.
 The number of mock stream star particles used to generate the map has been normalized such that the total number of stars within the observed extent of the stream (between $5.85 > l > 3.81~{\rm deg}$) is equal to the number of stars attributed to the Ophiuchus stream in the PS1 data \citep[$N \approx 500$][]{bernard14}.
+
+The viewing angle and stream geometry is more clearly demonstrated in Figure~\ref{fig:mockstreamxyz}, which shows $x$-$z$ projections of the star particles in Galactocentric Cartesian coordinates (grey) along with the position of the Sun (symbol at $(x,z)=(-8.3,0)~{\rm kpc}$) and the ``window'' of the heliocentric, sky-position plots of Figures~\ref{fig:mockstream0}--\ref{fig:mockstream1} (shown as blue lines).
 

### 2016-01-24 ms.tex
rix comments

@@
 The Ophiuchus stellar stream is peculiar: (1) its length is short given the age of its constituent stars, and (2) several probable member stars that lie close in both sky position and velocity have dispersions in these dimensions that far exceed those seen within the stream.
-The stream's proximity to the Galactic center suggests that the bar should have a significant influence on its dynamical history: The triaxiality and time-dependence of the bar may generate chaotic orbits in the vicinity of the stream that can greatly affect its morphology.
+The stream's proximity to the Galactic center suggests that the bar must have a significant influence on its dynamical history: The triaxiality and time-dependence of the bar may generate chaotic orbits in the vicinity of the stream that can greatly affect its morphology.
 We explore this hypothesis with models of stream formation along orbits consistent with Ophiuchus' properties in a Milky Way potential model that includes a rotating bar.
@@
 Mock streams generated along these orbits qualitatively match the observed properties of the stream: because of chaos, stars stripped early generally form low-density, high-dispersion ``fans'' leaving only the most recently disrupted material detectable as a strong over-density.
-These results motivate efforts to obtain deep photometry and spectroscopy of the region surrounding the stream. 
-The existence of or lack of low-surface-brightness tidal debris would provide an interesting constraint on the properties of the Milky Way bar and would help distinguish between formation scenarios for the stream.
+Our models predict that there should be more low-surface-brightness tidal debris than detected so far, likely with a complex phase-space morphology. 
+The existence of or lack of these features around the Ophiuchus stream would provide an interesting constraint on the properties of the Milky Way bar and would help distinguish between formation scenarios for the stream.
 This is the first time that chaos has been used to explain the properties of a stellar stream and is the first demonstration of the dynamical importance of chaos in the Galactic halo.
+The existence of long, thin streams around the Milky Way---presumably formed along non- or weakly-chaotic orbits---may represent only a subset of the total population of disrupted satellites.
 \end{abstract}
@@
 The Ophiuchus stream \citep{bernard14, sesar15a} is a recently discovered stellar tidal stream that sits above the Galactic bulge at a Galactocentric radius and height $(R,z) \approx (1.5, 4.3)~{\rm kpc}$.
-All observational evidence suggests that the stream is a totally disrupted globular cluster: The stream stars have (1) a small positional dispersion orthogonal to the extended direction of the stream (width $\approx$10 pc, length $\approx$1.5 kpc); (2) no detectable over-density along the stream that could be the progenitor system; (3) a small velocity dispersion $\approx$0.4 ${\rm km}~{\rm s}^{-1}$; and (4) an old stellar population ($\approx$12 Gyr) estimated from isochrone fitting \citep[][hereafter S15]{sesar15a}.
+All observational evidence suggests that the stream is a completely disrupted globular cluster: The stream stars have (1) a small positional dispersion orthogonal to the extended direction of the stream (width $\approx$10 pc, length $\approx$1.5 kpc); (2) no detectable over-density along the stream that could be the progenitor system; (3) a small velocity dispersion $\approx$0.4 ${\rm km}~{\rm s}^{-1}$; and (4) an old stellar population ($\approx$12 Gyr) estimated from isochrone fitting \citep[][hereafter S15]{sesar15a}.
 
@@
 S15 find that---on this orbit---the portion of the stream visible as an over-density in main-sequence stars must have been formed in the last $\lesssim$400 Myr for the stream to remain as short as it is observed.
-This dynamical age is at odds with the old stellar population: The stars appear to have formed $\approx$12 Gyr ago, but the abrupt end of the stream suggests that the cluster apparently fully disrupted at once in the last 400 Myrs.
+This dynamical age is at odds with the old ($\approx$10--12 Gyr) stellar population: The abrupt end of the stream suggests that the cluster apparently fully disrupted at once in the last 400 Myrs.
 Another puzzle is the existence of blue horizontal branch (BHB) stars close to the stream (within a few degrees) with similar radial velocities, but with a large dispersion in both sky position and velocity \citep[][hereafter S16]{sesar16}.
 The stream has a very distinct and large line-of-sight velocity ($\approx$290 \kms) and is therefore easily detected above the background halo population.
-Four BHB stars have been detected with line-of-sight velocities $>230~\kms$ that lie close to an extrapolation of the stream on the sky, but have a velocity dispersion $\approx$75 times larger than the measured internal velocity dispersion of the stream stars.
-The velocities of these stars are highly inconsistent with belonging to the background halo population (see Section 4, S16).
+Four BHB stars have been detected with line-of-sight velocities $>230~\kms$ that lie close to an extrapolation of the stream on the sky.
+This makes them likely members of the stream, as their velocities are in stark contrast to the background halo population (see Section 4, S16).
+Yet, they have a velocity dispersion $\approx$75 times larger than the measured internal velocity dispersion of the stream stars.
 These stars hint at the existence of associated low-density, high-dispersion features that were not modeled in S15 and are not predicted by the $N$-body simulations from this prior work, which assume a sudden, total disruption of the stream progenitor.
@@
 
-Recent work has shown that chaos can dramatically alter the density evolution of tidal streams \citep[e.g.,][]{fardal14, apw15-chaos}.
+Recent work has shown that dynamical chaos can dramatically alter the density evolution of tidal streams \citep[e.g.,][]{fardal14, apw15-chaos}.
 Along certain chaotic orbits, the stream stars will spread much faster in 3D position than from ordinary phase-mixing and, depending on the orbital phase at which the stream is observed, may develop large, low-density ``fans'' of stars at the ends of a stream \citep{pearson15, apw15-chaos}.
@@
 \end{enumerate}
-Our goal is not to perfectly represent the observed data but rather to explore the plausibility of explaining the peculiarities of the stream using chaotic stream fanning.
-Note that an alternate model was recently proposed that instead places the stream progenitor on an orbit in resonance with the bar.
+We do not aim to perfectly represent the observed data, but rather to explore the plausibility of explaining the peculiarities of the stream using chaotic stream fanning.
+Note that an alternate model was recently proposed that instead places the stream progenitor on an orbit in resonance with the bar \citep{hattori15}.
 We discuss the differences between these two models in Section~\ref{sec:discussion}.
@@
 Note that the four higher dispersion stars (the four stars with highest longitude) were not used when computing the likelihood and are only shown for completeness.
-}
+Though these four stars are significant outliers relative to the extrapolated orbit, they are (1) at the correct distance and sky position relative to the stream and (2) have velocities $\approx$2.5$\sigma$ discrepant with the halo velocity distribution in this region.}
 \label{fig:orbitfits}
@@
 \includegraphics[width=0.5\textwidth]{figures/lyapunov-hist}
-\caption{ Histograms of estimates of the Lyapunov time, $\lyapt=1/\lyapexp$, for posterior samples from the orbit fitting procedure (Section~\ref{sec:orbitfit}).
-Orbits are typically more chaotic (have shorter Lyapunov times) for higher pattern speeds.
-There is little dependence on bar angle.}
+\caption{ Histograms of estimated Lyapunov time, $\lyapt=1/\lyapexp$, for posterior samples from the orbit fitting procedure (Section~\ref{sec:orbitfit}).
+More chaotic orbits, i.e.
+those with shorter Lyapunov times, typically occur in barred potentials with higher pattern speeds.
+There is little dependence on bar angle.
+All orbits in these barred potential models are strongly chaotic with $t_\lambda \ll t_{\rm Hubble}$ and $t_\lambda \sim t_{\rm orbit}$, where the typical orbital period for Ophiuchus stream stars is a few hundred Megayears.}
 \label{fig:lyapunov-hist}
@@
 
-The model streams presented here do not reproduce all observed features of the Ophiuchus stream.
+The model streams presented here do not reproduce all observed features of the Ophiuchus stream (the details of the potential and the orbit predict vastly different phase-space morphologies for the fanned part).
 Instead, these results illustrate that chaotic evolution of tidal debris can plausibly explain the peculiar features of the stream.

### 2016-01-25 ms.tex
bowie and bibliography

@@
 The authors wish to acknowledge Victor Debattista, Melissa Ness, and Sarah Pearson for useful discussions.
+APW acknowledges David Bowie (1947--2016) for his continuous inspiration and artistic vision.
 This research made use of Astropy, a community-developed core Python package for Astronomy \citep{astropy13}.
@@
 
-\bibliographystyle{apj}
-\bibliography{paper-helpers/references/refs}
+%\bibliographystyle{apj}
+%\bibliography{paper-helpers/references/refs}
+
+\begin{thebibliography}{}
+\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
+
+\bibitem[{{Antoja} {et~al.}(2014){Antoja}, {Helmi}, {Dehnen}, {Bienaym{\'e}},
+  {Bland-Hawthorn}, {Famaey}, {Freeman}, {Gibson}, {Gilmore}, {Grebel},
+  {Kordopatis}, {Kunder}, {Minchev}, {Munari}, {Navarro}, {Parker}, {Reid},
+  {Seabroke}, {Siebert}, {Steinmetz}, {Watson}, {Wyse}, \&
+  {Zwitter}}]{antoja14}
+{Antoja}, T., {Helmi}, A., {Dehnen}, W., {et~al.} 2014, \aap, 563, A60
+
+\bibitem[{{Astropy Collaboration} {et~al.}(2013){Astropy Collaboration},
+  {Robitaille}, {Tollerud}, {Greenfield}, {Droettboom}, {Bray}, {Aldcroft},
+  {Davis}, {Ginsburg}, {Price-Whelan}, {Kerzendorf}, {Conley}, {Crighton},
+  {Barbary}, {Muna}, {Ferguson}, {Grollier}, {Parikh}, {Nair}, {Unther},
+  {Deil}, {Woillez}, {Conseil}, {Kramer}, {Turner}, {Singer}, {Fox}, {Weaver},
+  {Zabalza}, {Edwards}, {Azalee Bostroem}, {Burke}, {Casey}, {Crawford},
+  {Dencheva}, {Ely}, {Jenness}, {Labrie}, {Lim}, {Pierfederici}, {Pontzen},
+  {Ptak}, {Refsdal}, {Servillat}, \& {Streicher}}]{astropy13}
+{Astropy Collaboration}, {Robitaille}, T.~P., {Tollerud}, E.~J., {et~al.} 2013,
+  \aap, 558, A33
+
+\bibitem[{{Bernard} {et~al.}(2014){Bernard}, {Ferguson}, {Schlafly}, {Abbas},
+  {Bell}, {Deacon}, {Martin}, {Rix}, {Sesar}, {Slater}, {Pe{\~n}arrubia},
+  {Wyse}, {Burgett}, {Chambers}, {Draper}, {Hodapp}, {Kaiser}, {Kudritzki},
+  {Magnier}, {Metcalfe}, {Morgan}, {Price}, {Tonry}, {Wainscoat}, \&
+  {Waters}}]{bernard14}
+{Bernard}, E.~J., {Ferguson}, A.~M.~N., {Schlafly}, E.~F., {et~al.} 2014,
+  \mnras, 443, L84
+
+\bibitem[{{Blitz} \& {Spergel}(1991)}]{blitz91}
+{Blitz}, L., \& {Spergel}, D.~N. 1991, \apj, 379, 631
+
+\bibitem[{{Bonaca} {et~al.}(2014){Bonaca}, {Geha}, {K{\"u}pper}, {Diemand},
+  {Johnston}, \& {Hogg}}]{bonaca14}
+{Bonaca}, A., {Geha}, M., {K{\"u}pper}, A.~H.~W., {et~al.} 2014, \apj, 795, 94
+
+\bibitem[{{Bovy}(2015)}]{bovy15-galpy}
+{Bovy}, J. 2015, \apjs, 216, 29
+
+\bibitem[{{Bovy} \& {Rix}(2013)}]{bovyrix13}
+{Bovy}, J., \& {Rix}, H.-W. 2013, \apj, 779, 115
+
+\bibitem[{{Bovy} \& {Tremaine}(2012)}]{bovytremaine12}
+{Bovy}, J., \& {Tremaine}, S. 2012, \apj, 756, 89
+
+\bibitem[{{Bovy} {et~al.}(2012){Bovy}, {Allende Prieto}, {Beers}, {Bizyaev},
+  {da Costa}, {Cunha}, {Ebelke}, {Eisenstein}, {Frinchaboy}, {Garc{\'{\i}}a
+  P{\'e}rez}, {Girardi}, {Hearty}, {Hogg}, {Holtzman}, {Maia}, {Majewski},
+  {Malanushenko}, {Malanushenko}, {M{\'e}sz{\'a}ros}, {Nidever}, {O'Connell},
+  {O'Donnell}, {Oravetz}, {Pan}, {Rocha-Pinto}, {Schiavon}, {Schneider},
+  {Schultheis}, {Skrutskie}, {Smith}, {Weinberg}, {Wilson}, \&
+  {Zasowski}}]{bovy12}
+{Bovy}, J., {Allende Prieto}, C., {Beers}, T.~C., {et~al.} 2012, \apj, 759, 131
+
+\bibitem[{{Dwek} {et~al.}(1995){Dwek}, {Arendt}, {Hauser}, {Kelsall}, {Lisse},
+  {Moseley}, {Silverberg}, {Sodroski}, \& {Weiland}}]{dwek95}
+{Dwek}, E., {Arendt}, R.~G., {Hauser}, M.~G., {et~al.} 1995, \apj, 445, 716
+
+\bibitem[{{Fardal} {et~al.}(2015){Fardal}, {Huang}, \& {Weinberg}}]{fardal14}
+{Fardal}, M.~A., {Huang}, S., \& {Weinberg}, M.~D. 2015, \mnras, 452, 301
+
+\bibitem[{{Foreman-Mackey} {et~al.}(2013){Foreman-Mackey}, {Hogg}, {Lang}, \&
+  {Goodman}}]{foremanmackey13}
+{Foreman-Mackey}, D., {Hogg}, D.~W., {Lang}, D., \& {Goodman}, J. 2013, \pasp,
+  125, 306
+
+\bibitem[{{Gajda} {et~al.}(2015){Gajda}, {Lokas}, \& {Athanassoula}}]{gajda15}
+{Gajda}, G., {Lokas}, E.~L., \& {Athanassoula}, E. 2015, ArXiv e-prints,
+  arXiv:1511.04253
+
+\bibitem[{{Gnedin} \& {Ostriker}(1997)}]{gnedin97}
+{Gnedin}, O.~Y., \& {Ostriker}, J.~P. 1997, \apj, 474, 223
+
+\bibitem[{{Goodman} \& {Weare}(2010)}]{goodman10}
+{Goodman}, J., \& {Weare}, J. 2010, Comm. App. Math. Comp. Sci., 5, 65
+
+\bibitem[{{Hattori} {et~al.}(2015){Hattori}, {Erkal}, \& {Sanders}}]{hattori15}
+{Hattori}, K., {Erkal}, D., \& {Sanders}, J.~L. 2015, ArXiv e-prints,
+  arXiv:1512.04536
+
+\bibitem[{{Hernquist}(1990)}]{hernquist90}
+{Hernquist}, L. 1990, \apj, 356, 359
+
+\bibitem[{{Hernquist} \& {Ostriker}(1992)}]{hernquist92}
+{Hernquist}, L., \& {Ostriker}, J.~P. 1992, \apj, 386, 375
+
+\bibitem[{Jones {et~al.}(2001--)Jones, Oliphant, Peterson, {et~al.}}]{scipy}
+Jones, E., Oliphant, T., Peterson, P., {et~al.} 2001--, {SciPy}: Open source
+  scientific tools for {Python}, [Online; accessed 2015-02-11]
+
+\bibitem[{{Koposov} {et~al.}(2010){Koposov}, {Rix}, \& {Hogg}}]{koposov10}
+{Koposov}, S.~E., {Rix}, H.-W., \& {Hogg}, D.~W. 2010, \apj, 712, 260
+
+\bibitem[{{Lowing} {et~al.}(2011){Lowing}, {Jenkins}, {Eke}, \&
+  {Frenk}}]{lowing11}
+{Lowing}, B., {Jenkins}, A., {Eke}, V., \& {Frenk}, C. 2011, \mnras, 416, 2697
+
+\bibitem[{{Miyamoto} \& {Nagai}(1975)}]{miyamoto75}
+{Miyamoto}, M., \& {Nagai}, R. 1975, \pasj, 27, 533
+
+\bibitem[{{Murali} \& {Weinberg}(1997)}]{murali97}
+{Murali}, C., \& {Weinberg}, M.~D. 1997, \mnras, 291, 717
+
+\bibitem[{{Navarro} {et~al.}(1996){Navarro}, {Frenk}, \& {White}}]{navarro96}
+{Navarro}, J.~F., {Frenk}, C.~S., \& {White}, S.~D.~M. 1996, \apj, 462, 563
+
+\bibitem[{{Pearson} {et~al.}(2015){Pearson}, {K{\"u}pper}, {Johnston}, \&
+  {Price-Whelan}}]{pearson15}
+{Pearson}, S., {K{\"u}pper}, A.~H.~W., {Johnston}, K.~V., \& {Price-Whelan},
+  A.~M. 2015, \apj, 799, 28
+
+\bibitem[{{Portail} {et~al.}(2015{\natexlab{a}}){Portail}, {Wegg}, \&
+  {Gerhard}}]{portail15b}
+{Portail}, M., {Wegg}, C., \& {Gerhard}, O. 2015{\natexlab{a}}, \mnras, 450,
+  L66
+
+\bibitem[{{Portail} {et~al.}(2015{\natexlab{b}}){Portail}, {Wegg}, {Gerhard},
+  \& {Martinez-Valpuesta}}]{portail15}
+{Portail}, M., {Wegg}, C., {Gerhard}, O., \& {Martinez-Valpuesta}, I.
+  2015{\natexlab{b}}, \mnras, 448, 713
+
+\bibitem[{Powell(1964)}]{powell64}
+Powell, M. J.~D. 1964, The Computer Journal, 7, 155
+
+\bibitem[{Price-Whelan {et~al.}(2016)Price-Whelan, Johnston, Valluri, Pearson,
+  KŸpper, \& Hogg}]{apw15-chaos}
+Price-Whelan, A.~M., Johnston, K.~V., Valluri, M., {et~al.} 2016, Monthly
+  Notices of the Royal Astronomical Society, 455, 1079
+
+\bibitem[{{Sch{\"o}nrich}(2012)}]{schoenrich12}
+{Sch{\"o}nrich}, R. 2012, \mnras, 427, 274
+
+\bibitem[{{Sch{\"o}nrich} {et~al.}(2010){Sch{\"o}nrich}, {Binney}, \&
+  {Dehnen}}]{schoenrich10}
+{Sch{\"o}nrich}, R., {Binney}, J., \& {Dehnen}, W. 2010, \mnras, 403, 1829
+
+\bibitem[{{Sesar} {et~al.}(2015){Sesar}, {Bovy}, {Bernard}, {Caldwell},
+  {Cohen}, {Fouesneau}, {Johnson}, {Ness}, {Ferguson}, {Martin},
+  {Price-Whelan}, {Rix}, {Schlafly}, {Burgett}, {Chambers}, {Flewelling},
+  {Hodapp}, {Kaiser}, {Magnier}, {Platais}, {Tonry}, {Waters}, \&
+  {Wyse}}]{sesar15a}
+{Sesar}, B., {Bovy}, J., {Bernard}, E.~J., {et~al.} 2015, \apj, 809, 59
+
+\bibitem[{{Sesar} {et~al.}(2016){Sesar}, {Price-Whelan}, {Cohen}, {Rix},
+  {Pearson}, {Johnston}, {Bernard}, {Ferguson}, {Martin}, {Slater}, {Chambers},
+  {Flewelling}, {Wainscoat}, \& {Waters}}]{sesar16}
+{Sesar}, B., {Price-Whelan}, A.~M., {Cohen}, J.~G., {et~al.} 2016, \apjl, 816,
+  L4
+
+\bibitem[{{Wang} {et~al.}(2013){Wang}, {Mao}, {Long}, \& {Shen}}]{wang13}
+{Wang}, Y., {Mao}, S., {Long}, R.~J., \& {Shen}, J. 2013, \mnras, 435, 3437
+
+\bibitem[{{Wang} {et~al.}(2012){Wang}, {Zhao}, {Mao}, \& {Rich}}]{wang12}
+{Wang}, Y., {Zhao}, H., {Mao}, S., \& {Rich}, R.~M. 2012, \mnras, 427, 1429
+
+\bibitem[{{Wegg} \& {Gerhard}(2013)}]{wegg13}
+{Wegg}, C., \& {Gerhard}, O. 2013, \mnras, 435, 1874
+
+\bibitem[{{Weinberg}(1992)}]{weinberg92}
+{Weinberg}, M.~D. 1992, \apj, 384, 81
+
+\bibitem[{{Weinberg}(2015)}]{weinberg15}
+---. 2015, ArXiv e-prints, arXiv:1508.05959
+
+\bibitem[{{Zotos}(2012)}]{zotos12}
+{Zotos}, E.~E. 2012, Research in Astronomy and Astrophysics, 12, 500
+
+\end{thebibliography}
 
@@
 
-% ---------------------------------------------------------------------------------
-%\clearpage
-%\begin{figure}[p]
-%\begin{center}
-%\includegraphics[width=\textwidth]{figures/barred-circ-vel}
-%\caption{Circular velocity curve along the Sun-Galactic center line for the barred MW potential model.
-Solid black line shows total (sum of all components), lines below show a decomposition by potential component.
-Vertical grey bar shows approximate position of the Sun, horizontal grey bar shows roughly the range in measured circular velocity of the Sun.}
-%\label{fig:circ-vel-barred}
-%\end{center}
-%\end{figure}
-%
-%% ---------------------------------------------------------------------------------
-%\clearpage
-%\begin{figure}[p]
-%\begin{center}
-%\includegraphics[width=\textwidth]{figures/barred-surface-density-contour}
-%\caption{Contours of constant surface density for two projections of the barred MW potential model.
-For both face-on view (left panel) and edge-on view (right panel), four contours are drawn per decade in surface density between $10^7$ and $10^{12}~\msun~{\rm kpc}^{-3}$.
-Note the perturbation from the bar potential within Galactocentric radius $r \lesssim 4~{\rm kpc}$.
-The Sun's position is indicated by the `$\odot$' symbol.}
-%\label{fig:surface-density-barred}
-%\end{center}
-%\end{figure}
-%
-%% ---------------------------------------------------------------------------------
-%\clearpage
-%\begin{figure}[p]
-%\begin{center}
-%\includegraphics[width=\textwidth]{figures/static-circ-vel}
-%\caption{Same as Figure~\ref{fig:circ-vel-barred} for the static, axisymmetric MW potential model. }
-%\label{fig:circ-vel-static}
-%\end{center}
-%\end{figure}
-%
-%% ---------------------------------------------------------------------------------
-%\clearpage
-%\begin{figure}[p]
-%\begin{center}
-%\includegraphics[width=\textwidth]{figures/static-surface-density-contour}
-%\caption{Same as Figure~\ref{fig:surface-density-barred} for the static, axisymmetric MW potential model. }
-%\label{fig:surface-density-static}
-%\end{center}
-%\end{figure}
-
-% ---------------------------------------------------------------------------------
-%\clearpage
-%\begin{figure}[p]
-%\begin{center}
-%\includegraphics[width=\textwidth]{figures/lyapunov}
-%\caption{ Estimates of the Lyapunov time, $\lyapt=1/\lyapexp$, for the mean orbits in each potential model (Section~\ref{sec:results1}).
-The orbits are integrated from the mean values of orbital parameters from orbit fitting (Section~\ref{sec:orbitfit}).
-In the static, axisymmetric potential model, an orbit fit to the Ophiuchus stream is regular (e.g., the Lyapunov time keeps increasing).
-In all barred potential models (bar1--9), the mean orbit is chaotic with a Lyapunov time between 0.5--1 Gyr.}
-%\label{fig:lyapunov}
-%\end{center}
-%\end{figure}
-
-% ---------------------------------------------------------------------------------
-%\clearpage
-%\begin{figure}[p]
-%\begin{center}
-%\includegraphics[width=\textwidth]{figures/mass-loss}
-%\caption{ Computed stripping time distributions needed for each mock stream model in order to reproduce the observed distribution of BHB stars. }
-%\label{fig:massloss}
-%\end{center}
-%\end{figure}
-
 \end{document}

### 2016-01-25 ms.tex
remove old comments

@@
 
-%
-%\subsection{Detecting and characterizing chaos with frequency diffusion}\label{sec:chaos-indicator}
-%
-%Regular orbits are quasi-periodic: any phase-space component of the orbit can be decomposed in terms of a Fourier series, e.g.,
-%\begin{equation}
-%	x(t) = \sum^\infty_k \, a_{x,k}\, e^{i\, \omega_k \, t}.
-%\end{equation}
-%The amplitudes, $\bs{a}_k$ are constants which determine the strength of each mode, and the frequencies in each Fourier mode, $\omega_k$, are linear, integer combinations of the fundamental frequencies of the orbit, $\bs{\Omega}$---that is,
-%\begin{equation}
-%	\omega_k = \bs{n}_k \cdot \bs{\Omega}
-%\end{equation}
-%where $\bs{n}_k$ is a integer vector with dimensionality equal to the number of degrees of freedom of the gravitational potential \citep[e.g.,][]{binney82, binneytremaine}.
-The fundamental frequencies of a regular orbit are directly related to the orbital actions, a set of integrals of motion intricately related to the [...] of Hamiltonian and nonlinear dynamics \citep[e.g.,][]{??}.
-The fundamental frequencies can be approximated for numerically-integrated orbital time series by measuring the positions and amplitudes of spectral lines in the Fourier transform of the orbital components.
-A very accurate procedure for numerically computing the frequencies was developed by \citet{laskar93} and has been re-implemented in \python\ and released as open-source software \citep[\superfreq;][]{superfreq}.
-%
-%Chaotic orbits have no actions and therefore no fundamental frequencies.
-The frequencies and actions drift erratically in time with rates that depend on the structure of resonances surrounding such orbits.
-The evolution of the frequencies changes the configuration-space appearance of finite sub-sections of chaotic orbits and greatly enhances the density evolution of ensembles of nearby chaotic orbits; this has a dramatic effect on the morphological evolution of tidal streams \citep[e.g.,][]{apw15-chaos}.
-Numerically-integrated trajectories can be identified as chaotic orbits by the detection of frequency diffusion \citep[e.g.,][]{laskar96, valluri98}.
-%
-%The accuracy of the frequency-finding method (\superfreq) allows for detecting very small changes in the frequencies of even mildly chaotic orbits.
-Measuring the rate of frequency diffusion is therefore a powerful tool for detecting and characterizing dynamical chaos \citep[e.g.,][]{many}.
-From measuring the frequencies in many consecutive, overlapping time-windows along a numerically integrated orbit, we define the frequency diffusion rate, $DERP$, as
-%\begin{equation}
-%	yo
-%\end{equation}
-%where the index $k$ refers to the window number and $\tau$ is the shift in time between two consecutive windows.
-We note that this definition differs from previous, coarse approximations of the diffusion rate that simply differenced the frequencies computed in two consecutive time windows.
-We use this frequency diffusion rate to identify chaotic orbits integrated in the potential models described above.
-
 \subsection{Generating mock streams}\label{sec:mocks}
@@
 	\toprule
-%	name & $\phi_2$ [deg] & $d$ [kpc] & $\mu_l$ [mas yr$^{-1}$] & $\mu_b$ [mas yr$^{-1}$] & $v_r$ [km s$^{-1}$] & $s_{\phi_2}$ [deg] & $s_{d}$ [kpc] & $s_{v_r}$ [km s$^{-1}$]\\\midrule
-%static & $-0.03\pm0.05$ & $8.35\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.1\pm0.9$ & $0.20\pm0.04$ & $0.31\pm0.10$ & $2.9\pm0.8$\\
-%bar1--9 & $-0.03\pm0.05$ & $8.40\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.0\pm1.0$ & $0.21\pm0.05$ & $0.31\pm0.14$ & $3.2\pm0.9$\\
 	name & $\phi_2$ [deg] & $d$ [kpc] & $\mu_l$ [mas yr$^{-1}$] & $\mu_b$ [mas yr$^{-1}$] & $v_r$ [km s$^{-1}$]\\\midrule
@@
 	\toprule
-%	name & $\phi_2$ [deg] & $d$ [kpc] & $\mu_l$ [mas yr$^{-1}$] & $\mu_b$ [mas yr$^{-1}$] & $v_r$ [km s$^{-1}$] & $s_{\phi_2}$ [deg] & $s_{d}$ [kpc] & $s_{v_r}$ [km s$^{-1}$]\\\midrule
-%static & $-0.03\pm0.05$ & $8.35\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.1\pm0.9$ & $0.20\pm0.04$ & $0.31\pm0.10$ & $2.9\pm0.8$\\
-%bar1--9 & $-0.03\pm0.05$ & $8.40\pm0.05$ & $-7.2\pm0.1$ & $0.9\pm0.1$ & $289.0\pm1.0$ & $0.21\pm0.05$ & $0.31\pm0.14$ & $3.2\pm0.9$\\
 	$s_{\phi_2}$ [deg] & $s_{d}$ [kpc] & $s_{v_r}$ [km s$^{-1}$]\\\midrule
@@
 
-%We have applied the same distance-modulus cut to the points in these figures as used to select the BHB stars identified as stream members by S16 (see Section~2 in S16).
-
 \section{Discussion}\label{sec:discussion}
@@
 We consider the second scenario to be more plausible: our understanding of the formation and evolution of the Ophiuchus stream is that the progenitor object has been orbiting and steadily losing stars over the last several Gigayears, but only the stars stripped from the most recent few pericentric passages remain coherent enough to be detected as a stream-like over-density in the PS1 data.
-%[Gaia, Don't do distance cut that Brani did, chemistry]
 
@@
 With more information about the density distribution of stars in the stream and better proper motion measurements we would be able to (1) help distinguish models for the Milky Way bar independent from current methods and (2) begin to model the survivability of globular clusters orbiting in the central Milky Way \citep[e.g.,][]{gnedin97}.
-
-%\todo{Where to put this?}
-%
-%also, cite gnedin97
-%
-%A typical inner MW globular cluster with a mass $m_{\rm cl} \approx 10^4~\msun$ has a half-mass radius $r_h \approx 2~{\rm pc}$.
-We can ask, given an assumed mass density profile for the inner galaxy, what Galactocentric pericentric radius the Ophiuchus progenitor would have to reach such that the tidal radius of the cluster is equal to the half-mass radius.
-%
-%Let's assume the density profile of the inner $\approx 1~{\rm kpc}$ of the MW is a Jaffe spheroid so that
-%\begin{align}
-%	\rho(r) &= \frac{\rho_0}{(r/a)^2 \, (1 + r/a)^2}\\
-%	M(<r) &= 4\pi \, \rho_0 \, a^3 \, \frac{r/a}{1 + r/a}
-%\end{align}
-%with scale radius $a \approx 0.5~{\rm kpc}$ and mass within the scale radius $M(<a) \approx 10^{10}~\msun$ ($\rho_0 \approx \frac{10^{10}~\msun}{2\pi a^3} \approx 10^{10}~\msun~{\rm kpc}^{-3}$).
-The tidal radius is approximately
-%\begin{align}
-%	r_{\rm tide} \approx \left(\frac{m_{\rm cl}}{3M(<r)}\right)^{1/3} \, r
-%\end{align}
-%so that 
-%
-%then
-%\begin{align}
-%	2~{\rm pc} &= \left(\frac{m_{\rm cl}}{3M(<r)}\right)^{1/3} \, r\\
-%\end{align}
-%
-%
-%if we set
-%\begin{align}
-%	r_{\rm tide} = r_h = 2~{\rm pc} 
-%\end{align}
 

### 2016-01-25 ms.tex
back to single column
oops, figure caption

@@
 
-% \keywords{}
+\keywords{
+  Galaxy: halo
+  ---
+  globular clusters: general
+  ---
+  stars: kinematics and dynamics
+  ---
+  Galaxy: structure
+  ---
+  Galaxy: kinematics and dynamics
+}
 
@@
 \begin{center}
-\includegraphics[width=0.7\textwidth]{figures/potentials-four}
+%\includegraphics[width=0.7\textwidth]{figures/potentials-four}
+\includegraphics[width=\textwidth]{figures/potentials-four}
 \caption{{\it Left column:} Circular velocity curves along the Sun-Galactic center line for a representative barred MW potential model (top left) and for the static MW potential model (bottom left).
@@
 \begin{center}
-\includegraphics[width=0.5\textwidth]{figures/orbitfits}
+%\includegraphics[width=0.5\textwidth]{figures/orbitfits}
+\includegraphics[width=0.75\textwidth]{figures/orbitfits}
 \caption{ Results from fitting orbits to BHB stars associated with the Ophiuchus stream in the static and barred potential models.
@@
 \begin{center}
-\includegraphics[width=0.5\textwidth]{figures/orbit-yz}
+%\includegraphics[width=0.5\textwidth]{figures/orbit-yz}
+\includegraphics[width=\textwidth]{figures/orbit-yz}
 \caption{  Projections of orbits integrated from the mean orbital parameters estimated from the orbit fitting posterior distributions in each potential model.
@@
 \begin{center}
-\includegraphics[width=0.5\textwidth]{figures/lyapunov-hist}
+%\includegraphics[width=0.5\textwidth]{figures/lyapunov-hist}
+\includegraphics[width=\textwidth]{figures/lyapunov-hist}
 \caption{ Histograms of estimated Lyapunov time, $\lyapt=1/\lyapexp$, for posterior samples from the orbit fitting procedure (Section~\ref{sec:orbitfit}).
@@
 The position of the Sun is shown as the symbol at $(x,z)=(-8,0)~{\rm kpc}$.
-The corners of the sky position and distance plots of Figures~\ref{fig:mockstream0}--\ref{fig:mockstream1} are shown transformed to these coordinates as dark markers near $(x,z)=(-4,3)~{\rm kpc}$, $(x,z)=(0,4)~{\rm kpc}$.
+The volume of the sky position and distance plots of Figures~\ref{fig:mockstream0}--\ref{fig:mockstream1} are shown transformed to these coordinates as the blue wedge near $(x,z)=(-2,4)~{\rm kpc}$.
 This demonstrates that the stream is nearly aligned our viewing angle.

### 2016-01-25 ms.tex
fix links

@@
 For the barred potentials, all mean values are the same because the time-dependence of the bar doesn't impact the orbit fit over the short length of the stream.
-We have made samples from the full posterior distribution available with this article and provide code to transform to and from stream coordinates (see http://adrian.pw/ophiuchus/coordinates for more information).\label{tbl:param-means} }
+We have made samples from the full posterior distribution available with this article and provide code to transform to and from stream coordinates (see http://adrian.pw/ophiuchus for more information).\label{tbl:param-means} }
 \end{center}
@@
 \end{equation*}
-but the precise transformation and coordinate frame is implemented in \python\ using the \package{Astropy} coordinates package.\footnote{See \url{http://adrian.pw/ophiuchus/coordinates} for more information}  This code is hosted on \project{GitHub}.\footnote{\url{https://github.com/adrn/ophiuchus}}
+but the precise transformation and coordinate frame is implemented in \python\ using the \package{Astropy} coordinates package.\footnote{See \url{http://adrian.pw/ophiuchus} for more information}  This code is hosted on \project{GitHub}.\footnote{\url{https://github.com/adrn/ophiuchus}}
 

### 2016-03-01 ms.tex
minor comment

@@
 For example, recent work has shown that if the Ophichus stream progenitor orbit is in resonance with the bar, the debris can remain short for at least 1 Gyr \citep{hattori15}.
-In this model, there would be no nearby, high-dispersion debris, and the pattern speed of the bar would likely be related to the orbital frequencies of the progenitor orbit.
+In their model, there would be no nearby, high-dispersion debris, and the pattern speed of the bar would be related to the orbital frequencies of the progenitor orbit.
 However, it has not been demonstrated whether this proposed scheme can explain the shortness of the stream over timescales closer to the age of the stellar population ($\approx$10 Gyr).

### 2016-04-12 ms.tex
move stuff to paper helpers
first round of incorporations

@@
 \input{../paper-helpers/preamble.tex}
-
-% packages
-\usepackage{amssymb,amsmath,amsbsy}
-\usepackage{booktabs}
-\usepackage[caption=false]{subfig}
-\usepackage{multirow}
-
-\newcommand{\DM}{{\rm DM}}
-\newcommand{\norm}{\mathcal{N}}
-
-% TO DO
-\usepackage{graphicx} 
-\usepackage{color}
-\newcommand{\todo}[1]{{\color{red} TODO: #1}}
 
@@
 
-To generate mock stellar streams, we use the method presented in \citet{fardal14}: Star particles are `released' from a progenitor system near the Lagrange points with a dispersion in position and velocity that is set by the mass and orbit of the progenitor.
+To generate mock stellar streams, we use \changes{a method similar to that} presented in \citet{fardal14}: Star particles are `released' from a progenitor system near the Lagrange points with a dispersion in position and velocity that is set by the mass and orbit of the progenitor.
 We draw samples from the posterior probability distributions over orbital parameters from fitting orbits to the stream star members and use these as the progenitor orbital parameters (Section~\ref{sec:orbitfit}).
@@
 From the endpoint of the backwards-integration (e.g., the past position), we begin integrating the orbit forward in time, but now at each time-step a star particle is released near each of the Lagrange points of the progenitor.
-The position of the Lagrange points and the scale of the dispersion in position and velocity are set by the progenitor mass, $m$.
+\changes{This simplification assumes that the mass-loss rate is constant in time, which is not assumed in \citet{fardal14} but has been shown to closely match $N$-body simulations of globular cluster disruption \citep{kuepper12}.} The position of the Lagrange points and the scale of the dispersion in position and velocity are set by the progenitor mass, $m$.
 The star particles are drawn from Gaussians centered on the Lagrange points (in position) and the progenitor (in velocity) and the full parametrization of the release distribution is given in \cite{fardal14}.
@@
 At this time, we set the offset of the Lagrange point to 0: In terms of the parametrization in \citep{fardal14}, we set $k_r = 0$ and $k_{vt}=0$ but preserve the dispersion in the release radius and velocity.
-Any star released after $\tau_d$ is released with a small dispersion around the progenitor orbit but no offset.
-This mass-loss history is intended to mimic the expected gradual evaporation of a globular cluster over a tidally-limited boundary (i.e.
+\changes{That is, any star released after $\tau_d$ is drawn from a Gaussian with positional and velocity dispersion fixed to their respective values at $\tau_d$ with no offset from the progenitor orbit.} This mass-loss history is intended to mimic the expected gradual evaporation of a globular cluster over a tidally-limited boundary (i.e.
 driven by two-body relaxation and gravitational shocks over many Gigayears) with final disruption likely to occur once the tidal boundary is less than the core radius of the cluster.
@@
 Though the coordinate and velocity parameter values in observed coordinates are very similar between each potential model, the resulting orbits are quite different.
-For the posterior samples in each potential, we take the mean values of the coordinate and velocity parameters and convert to Galactocentric coordinates (e.g., Table~\ref{tbl:param-means}).
+For the posterior samples in each potential, we take the mean values of the coordinate and velocity parameters \changes{at $\phi_1 = 0$}and convert to Galactocentric coordinates (e.g., Table~\ref{tbl:param-means}).
 Figure~\ref{fig:orbits-yz} shows projections of these ``mean'' orbits in each potential model.
@@
 All orbits sampled from each barred potential are strongly chaotic with Lyapunov times that range from $\lyapt \approx 400$--$1100~{\rm Myr}$.
-It is clear from these panels that the orbits around Ophiuchus are generally more strongly chaotic (have lower Lyapunov times) for larger pattern speeds.
+\changes{From the left column to the right column, the distribution of Lyapunov times shifts slightly towards lower values suggesting that the orbits around Ophiuchus are more strongly chaotic for larger pattern speeds.}
 
@@
 \includegraphics[width=\textwidth]{figures/orbit-yz}
-\caption{  Projections of orbits integrated from the mean orbital parameters estimated from the orbit fitting posterior distributions in each potential model.
+\caption{  Projections of orbits integrated from the mean orbital parameters \changes{at $\phi_1 = 0$} estimated from the orbit fitting posterior distributions in each potential model.
 Orbits are integrated for 6 Gyr and shown in Galactocentric, Cartesian coordinates.
@@
 	\toprule
-	name & $\phi_2$ [deg] & $d$ [kpc] & $\mu_l$ [mas yr$^{-1}$] & $\mu_b$ [mas yr$^{-1}$] & $v_r$ [km s$^{-1}$]\\\midrule
+	name & $\phi_2$ [deg] & $d_\odot$ [kpc] & $\mu_l$ [mas yr$^{-1}$] & $\mu_b$ [mas yr$^{-1}$] & $v_r$ [km s$^{-1}$]\\\midrule
 	static & $-0.03\pm0.05$ & $8.3\pm0.05$ & $-7.4\pm0.1$ & $0.9\pm0.1$ & $288.9\pm0.9$\\
@@
 	\end{tabular}
-	\caption{Estimated mean and standard deviation of samples from the marginal posterior distributions over each parameter in our orbit fit model (the posterior distributions are very close to Gaussian).
+	\caption{Estimated mean and standard deviation of samples from the marginal posterior distributions over each parameter in our orbit fit model \changes{at $\phi_1 = 0$}(the posterior distributions are very close to Gaussian).
 For the barred potentials, all mean values are the same because the time-dependence of the bar doesn't impact the orbit fit over the short length of the stream.
@@
 \caption{ Star particles (grey points) from mock streams generated on the mean orbits in each potential model shown in projections of Galatocentric, Cartesian coordinates.
-The position of the Sun is shown as the symbol at $(x,z)=(-8,0)~{\rm kpc}$.
+The position of the Sun is shown as the symbol at $(x,z)=(-8.3,0)~{\rm kpc}$.
 The volume of the sky position and distance plots of Figures~\ref{fig:mockstream0}--\ref{fig:mockstream1} are shown transformed to these coordinates as the blue wedge near $(x,z)=(-2,4)~{\rm kpc}$.
@@
 
-%\bibliographystyle{apj}
-%\bibliography{paper-helpers/references/refs}
-
-\begin{thebibliography}{}
-\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
-
-\bibitem[{{Antoja} {et~al.}(2014){Antoja}, {Helmi}, {Dehnen}, {Bienaym{\'e}},
-  {Bland-Hawthorn}, {Famaey}, {Freeman}, {Gibson}, {Gilmore}, {Grebel},
-  {Kordopatis}, {Kunder}, {Minchev}, {Munari}, {Navarro}, {Parker}, {Reid},
-  {Seabroke}, {Siebert}, {Steinmetz}, {Watson}, {Wyse}, \&
-  {Zwitter}}]{antoja14}
-{Antoja}, T., {Helmi}, A., {Dehnen}, W., {et~al.} 2014, \aap, 563, A60
-
-\bibitem[{{Astropy Collaboration} {et~al.}(2013){Astropy Collaboration},
-  {Robitaille}, {Tollerud}, {Greenfield}, {Droettboom}, {Bray}, {Aldcroft},
-  {Davis}, {Ginsburg}, {Price-Whelan}, {Kerzendorf}, {Conley}, {Crighton},
-  {Barbary}, {Muna}, {Ferguson}, {Grollier}, {Parikh}, {Nair}, {Unther},
-  {Deil}, {Woillez}, {Conseil}, {Kramer}, {Turner}, {Singer}, {Fox}, {Weaver},
-  {Zabalza}, {Edwards}, {Azalee Bostroem}, {Burke}, {Casey}, {Crawford},
-  {Dencheva}, {Ely}, {Jenness}, {Labrie}, {Lim}, {Pierfederici}, {Pontzen},
-  {Ptak}, {Refsdal}, {Servillat}, \& {Streicher}}]{astropy13}
-{Astropy Collaboration}, {Robitaille}, T.~P., {Tollerud}, E.~J., {et~al.} 2013,
-  \aap, 558, A33
-
-\bibitem[{{Bernard} {et~al.}(2014){Bernard}, {Ferguson}, {Schlafly}, {Abbas},
-  {Bell}, {Deacon}, {Martin}, {Rix}, {Sesar}, {Slater}, {Pe{\~n}arrubia},
-  {Wyse}, {Burgett}, {Chambers}, {Draper}, {Hodapp}, {Kaiser}, {Kudritzki},
-  {Magnier}, {Metcalfe}, {Morgan}, {Price}, {Tonry}, {Wainscoat}, \&
-  {Waters}}]{bernard14}
-{Bernard}, E.~J., {Ferguson}, A.~M.~N., {Schlafly}, E.~F., {et~al.} 2014,
-  \mnras, 443, L84
-
-\bibitem[{{Blitz} \& {Spergel}(1991)}]{blitz91}
-{Blitz}, L., \& {Spergel}, D.~N. 1991, \apj, 379, 631
-
-\bibitem[{{Bonaca} {et~al.}(2014){Bonaca}, {Geha}, {K{\"u}pper}, {Diemand},
-  {Johnston}, \& {Hogg}}]{bonaca14}
-{Bonaca}, A., {Geha}, M., {K{\"u}pper}, A.~H.~W., {et~al.} 2014, \apj, 795, 94
-
-\bibitem[{{Bovy}(2015)}]{bovy15-galpy}
-{Bovy}, J. 2015, \apjs, 216, 29
-
-\bibitem[{{Bovy} \& {Rix}(2013)}]{bovyrix13}
-{Bovy}, J., \& {Rix}, H.-W. 2013, \apj, 779, 115
-
-\bibitem[{{Bovy} \& {Tremaine}(2012)}]{bovytremaine12}
-{Bovy}, J., \& {Tremaine}, S. 2012, \apj, 756, 89
-
-\bibitem[{{Bovy} {et~al.}(2012){Bovy}, {Allende Prieto}, {Beers}, {Bizyaev},
-  {da Costa}, {Cunha}, {Ebelke}, {Eisenstein}, {Frinchaboy}, {Garc{\'{\i}}a
-  P{\'e}rez}, {Girardi}, {Hearty}, {Hogg}, {Holtzman}, {Maia}, {Majewski},
-  {Malanushenko}, {Malanushenko}, {M{\'e}sz{\'a}ros}, {Nidever}, {O'Connell},
-  {O'Donnell}, {Oravetz}, {Pan}, {Rocha-Pinto}, {Schiavon}, {Schneider},
-  {Schultheis}, {Skrutskie}, {Smith}, {Weinberg}, {Wilson}, \&
-  {Zasowski}}]{bovy12}
-{Bovy}, J., {Allende Prieto}, C., {Beers}, T.~C., {et~al.} 2012, \apj, 759, 131
-
-\bibitem[{{Dwek} {et~al.}(1995){Dwek}, {Arendt}, {Hauser}, {Kelsall}, {Lisse},
-  {Moseley}, {Silverberg}, {Sodroski}, \& {Weiland}}]{dwek95}
-{Dwek}, E., {Arendt}, R.~G., {Hauser}, M.~G., {et~al.} 1995, \apj, 445, 716
-
-\bibitem[{{Fardal} {et~al.}(2015){Fardal}, {Huang}, \& {Weinberg}}]{fardal14}
-{Fardal}, M.~A., {Huang}, S., \& {Weinberg}, M.~D. 2015, \mnras, 452, 301
-
-\bibitem[{{Foreman-Mackey} {et~al.}(2013){Foreman-Mackey}, {Hogg}, {Lang}, \&
-  {Goodman}}]{foremanmackey13}
-{Foreman-Mackey}, D., {Hogg}, D.~W., {Lang}, D., \& {Goodman}, J. 2013, \pasp,
-  125, 306
-
-\bibitem[{{Gajda} {et~al.}(2015){Gajda}, {Lokas}, \& {Athanassoula}}]{gajda15}
-{Gajda}, G., {Lokas}, E.~L., \& {Athanassoula}, E. 2015, ArXiv e-prints,
-  arXiv:1511.04253
-
-\bibitem[{{Gnedin} \& {Ostriker}(1997)}]{gnedin97}
-{Gnedin}, O.~Y., \& {Ostriker}, J.~P. 1997, \apj, 474, 223
-
-\bibitem[{{Goodman} \& {Weare}(2010)}]{goodman10}
-{Goodman}, J., \& {Weare}, J. 2010, Comm. App. Math. Comp. Sci., 5, 65
-
-\bibitem[{{Hattori} {et~al.}(2015){Hattori}, {Erkal}, \& {Sanders}}]{hattori15}
-{Hattori}, K., {Erkal}, D., \& {Sanders}, J.~L. 2015, ArXiv e-prints,
-  arXiv:1512.04536
-
-\bibitem[{{Hernquist}(1990)}]{hernquist90}
-{Hernquist}, L. 1990, \apj, 356, 359
-
-\bibitem[{{Hernquist} \& {Ostriker}(1992)}]{hernquist92}
-{Hernquist}, L., \& {Ostriker}, J.~P. 1992, \apj, 386, 375
-
-\bibitem[{Jones {et~al.}(2001--)Jones, Oliphant, Peterson, {et~al.}}]{scipy}
-Jones, E., Oliphant, T., Peterson, P., {et~al.} 2001--, {SciPy}: Open source
-  scientific tools for {Python}, [Online; accessed 2015-02-11]
-
-\bibitem[{{Koposov} {et~al.}(2010){Koposov}, {Rix}, \& {Hogg}}]{koposov10}
-{Koposov}, S.~E., {Rix}, H.-W., \& {Hogg}, D.~W. 2010, \apj, 712, 260
-
-\bibitem[{{Lowing} {et~al.}(2011){Lowing}, {Jenkins}, {Eke}, \&
-  {Frenk}}]{lowing11}
-{Lowing}, B., {Jenkins}, A., {Eke}, V., \& {Frenk}, C. 2011, \mnras, 416, 2697
-
-\bibitem[{{Miyamoto} \& {Nagai}(1975)}]{miyamoto75}
-{Miyamoto}, M., \& {Nagai}, R. 1975, \pasj, 27, 533
-
-\bibitem[{{Murali} \& {Weinberg}(1997)}]{murali97}
-{Murali}, C., \& {Weinberg}, M.~D. 1997, \mnras, 291, 717
-
-\bibitem[{{Navarro} {et~al.}(1996){Navarro}, {Frenk}, \& {White}}]{navarro96}
-{Navarro}, J.~F., {Frenk}, C.~S., \& {White}, S.~D.~M. 1996, \apj, 462, 563
-
-\bibitem[{{Pearson} {et~al.}(2015){Pearson}, {K{\"u}pper}, {Johnston}, \&
-  {Price-Whelan}}]{pearson15}
-{Pearson}, S., {K{\"u}pper}, A.~H.~W., {Johnston}, K.~V., \& {Price-Whelan},
-  A.~M. 2015, \apj, 799, 28
-
-\bibitem[{{Portail} {et~al.}(2015{\natexlab{a}}){Portail}, {Wegg}, \&
-  {Gerhard}}]{portail15b}
-{Portail}, M., {Wegg}, C., \& {Gerhard}, O. 2015{\natexlab{a}}, \mnras, 450,
-  L66
-
-\bibitem[{{Portail} {et~al.}(2015{\natexlab{b}}){Portail}, {Wegg}, {Gerhard},
-  \& {Martinez-Valpuesta}}]{portail15}
-{Portail}, M., {Wegg}, C., {Gerhard}, O., \& {Martinez-Valpuesta}, I.
-  2015{\natexlab{b}}, \mnras, 448, 713
-
-\bibitem[{Powell(1964)}]{powell64}
-Powell, M. J.~D. 1964, The Computer Journal, 7, 155
-
-\bibitem[{Price-Whelan {et~al.}(2016)Price-Whelan, Johnston, Valluri, Pearson,
-  KŸpper, \& Hogg}]{apw15-chaos}
-Price-Whelan, A.~M., Johnston, K.~V., Valluri, M., {et~al.} 2016, Monthly
-  Notices of the Royal Astronomical Society, 455, 1079
-
-\bibitem[{{Sch{\"o}nrich}(2012)}]{schoenrich12}
-{Sch{\"o}nrich}, R. 2012, \mnras, 427, 274
-
-\bibitem[{{Sch{\"o}nrich} {et~al.}(2010){Sch{\"o}nrich}, {Binney}, \&
-  {Dehnen}}]{schoenrich10}
-{Sch{\"o}nrich}, R., {Binney}, J., \& {Dehnen}, W. 2010, \mnras, 403, 1829
-
-\bibitem[{{Sesar} {et~al.}(2015){Sesar}, {Bovy}, {Bernard}, {Caldwell},
-  {Cohen}, {Fouesneau}, {Johnson}, {Ness}, {Ferguson}, {Martin},
-  {Price-Whelan}, {Rix}, {Schlafly}, {Burgett}, {Chambers}, {Flewelling},
-  {Hodapp}, {Kaiser}, {Magnier}, {Platais}, {Tonry}, {Waters}, \&
-  {Wyse}}]{sesar15a}
-{Sesar}, B., {Bovy}, J., {Bernard}, E.~J., {et~al.} 2015, \apj, 809, 59
-
-\bibitem[{{Sesar} {et~al.}(2016){Sesar}, {Price-Whelan}, {Cohen}, {Rix},
-  {Pearson}, {Johnston}, {Bernard}, {Ferguson}, {Martin}, {Slater}, {Chambers},
-  {Flewelling}, {Wainscoat}, \& {Waters}}]{sesar16}
-{Sesar}, B., {Price-Whelan}, A.~M., {Cohen}, J.~G., {et~al.} 2016, \apjl, 816,
-  L4
-
-\bibitem[{{Wang} {et~al.}(2013){Wang}, {Mao}, {Long}, \& {Shen}}]{wang13}
-{Wang}, Y., {Mao}, S., {Long}, R.~J., \& {Shen}, J. 2013, \mnras, 435, 3437
-
-\bibitem[{{Wang} {et~al.}(2012){Wang}, {Zhao}, {Mao}, \& {Rich}}]{wang12}
-{Wang}, Y., {Zhao}, H., {Mao}, S., \& {Rich}, R.~M. 2012, \mnras, 427, 1429
-
-\bibitem[{{Wegg} \& {Gerhard}(2013)}]{wegg13}
-{Wegg}, C., \& {Gerhard}, O. 2013, \mnras, 435, 1874
-
-\bibitem[{{Weinberg}(1992)}]{weinberg92}
-{Weinberg}, M.~D. 1992, \apj, 384, 81
-
-\bibitem[{{Weinberg}(2015)}]{weinberg15}
----. 2015, ArXiv e-prints, arXiv:1508.05959
-
-\bibitem[{{Zotos}(2012)}]{zotos12}
-{Zotos}, E.~E. 2012, Research in Astronomy and Astrophysics, 12, 500
-
-\end{thebibliography}
+\bibliographystyle{apj}
+\bibliography{paper-helpers/references/refs}
 
@@
 This may be a bad assumption.} We add two additional nuisance parameters for controlling the amount of time to integrate forwards, $t_f$, and backwards, $t_b$, from the given initial conditions, which ultimately controls the length of the section of orbit that is compared to the stream star data.
-For brevity in the equations below, we define $\bs{s} = (s_{\phi_2}, s_{\DM}, s_{v_r})$ and $\bs{\theta} = (\bs{w}_0, \Phi, t_b, t_f)$.
+For brevity in the equations below, we define $\bs{s} = (s_{\phi_2}, s_{\DM}, s_{v_r})$, $\bs{\sigma}_i = (\sigma_{\DM}, \sigma_{\mu_l}, \sigma_{\mu_b}, \sigma_{v_r})_i$, and $\bs{\theta} = (\bs{w}_0, \Phi, t_b, t_f)$.
 

### 2016-04-12 ms.tex
added a pm figure

@@
 
+\begin{figure}[!tbp]
+\begin{center}
+%\includegraphics[width=0.5\textwidth]{figures/orbitfits}
+\includegraphics[width=0.75\textwidth]{figures/orbitfits-pm}
+\caption{ \changes{Same as Figure~\ref{fig:orbitfits} for the proper motion components.
+Proper motions have not been measured for the ``fanned'' stars, so these are not shown.
+The proper motion measurements have large uncertainties and do not contribute significantly to the orbit fitting likelihood.}}
+\label{fig:orbitfits-pm}
+\end{center}
+\end{figure}
+
 We make one modification to this method based on the idea that the Ophiuchus stream progenitor has been fully-disrupted.
@@
 
-Figure~\ref{fig:orbitfits} summarizes our results from fitting orbits to the BHB stream stars.
+Figures~\ref{fig:orbitfits}--\ref{fig:orbitfits-pm} summarize our results from fitting orbits to the BHB stream stars.
 Shown in each panel are the high-probability Ophiuchus stream stars (black points, to which orbits are fit) and orbits integrated from samples from the posterior probability over orbital parameters (blue lines).
@@
 We only show one of the barred potentials: The end-to-end integration time of the orbit over the observed extent of the stream is only $\approx$6 Myr, so the derived orbits are extremely similar in all potentials (the time-dependence of the bar potential is not significant over such short timescales).
-The orbital periods are typically $\approx$170--200 Myr with pericenters $r_p \approx 4~{\rm kpc}$ and apocenters $r_a \approx 12$--$15~{\rm kpc}$.
+\changes{As was previously noted, there is a marginal discrepancy between the orbit fits and the observed proper motions, but it is possible there is a systematic offset in the proper motion measurements \citep[see Figure~10 in][]{sesar15a}.
+We conclude that it is too soon to tell whether there is in fact an inconsistency.} The orbital periods are typically $\approx$170--200 Myr with pericenters $r_p \approx 4~{\rm kpc}$ and apocenters $r_a \approx 12$--$15~{\rm kpc}$.
 Though the coordinate and velocity parameter values in observed coordinates are very similar between each potential model, the resulting orbits are quite different.
@@
 All orbits sampled from each barred potential are strongly chaotic with Lyapunov times that range from $\lyapt \approx 400$--$1100~{\rm Myr}$.
-\changes{From the left column to the right column, the distribution of Lyapunov times shifts slightly towards lower values suggesting that the orbits around Ophiuchus are more strongly chaotic for larger pattern speeds.}
+\changes{From the left column to the right column, the distribution of Lyapunov times shifts slightly towards lower values suggesting that the orbits around Ophiuchus are more strongly chaotic for larger pattern speeds (within the range considered in this work).}
 

### 2016-04-12 ms.tex
added proper motion panesl

@@
 \input{../paper-helpers/preamble.tex}
+\usepackage{pdflscape}
 
@@
 
-Figures~\ref{fig:mockstream0}--\ref{fig:mockstream1} show the final particle positions and line-of-sight velocities in heliocentric coordinates for the maximum likelihood mock streams (grey points) in each potential.
-Vertical, dashed lines show the approximate extent of the part of the stream visible in main-sequence stars (excluding the BHB stars from S16).
-There are a few interesting features to note from these panels:
+Figures~\ref{fig:mockstream0}--\ref{fig:mockstream1} show the final particle positions and line-of-sight velocities in heliocentric coordinates for the maximum likelihood mock streams (points) in each potential.
+Vertical lines show the approximate extent of the part of the stream visible in main-sequence stars (excluding the BHB stars from S16).
+\changes{Figure~\ref{fig:mockstream-pm0}--\ref{fig:mockstream-pm1} show the same for proper motion components.} There are a few interesting features to note from these panels:
 \begin{enumerate}
@@
 	\item The density of the mock stream in the static potential decreases slowly rather than abruptly as is observed.
-This is shown in the top panels of each column where each contour level represents a factor of 10 difference in projected surface density.
+This is more easily seen in Figure~\ref{fig:mockstream-density} where each contour level represents a factor of 10 difference in projected surface density.
 In the static potential model the length of dense debris extends much farther than the observed extent of the stream (vertical dashed lines), whereas in some of the barred potential models the stars released earlier have ``fanned'' and are associated with much lower density debris (e.g., bar8).
@@
 Instead, these results illustrate that chaotic evolution of tidal debris can plausibly explain the peculiar features of the stream.
-If the cluster progenitor was on a regular orbit, it would have to have disrupted entirely within the last $\approx$300--600 Myr in order to explain the shortness and density profile.
+\changes{If the cluster progenitor was on a regular orbit fit to the observed stream stars in a static, axisymmetric potential model for the Milky Way}, it would have to have disrupted entirely within the last $\approx$300--600 Myr in order to explain the shortness and density profile.
 In addition, the four most recently identified BHB stars with similar distances and line-of-sight velocities would have to be (highly unlikely) chance alignments of halo stars.
@@
 % --------------------------------------------------------------------------------- 
-
+\begin{landscape}
 \begin{figure*}[p]
 \begin{center}
-\includegraphics[width=\textwidth]{figures/mockstream0}
-\caption{ Sky position, distance, and line-of-sight velocity in heliocentric, Galactic coordinates for star particles (contours or grey points) from the maximum-likelihood mock streams in each potential.
-Top panels show surface density of mock stream star particles in each potential.
-Contours are spaced logarithmically from $10^{-2}$ to $10$ particles per sq.
-deg---that is, each color represents a factor of 10 difference in surface density.
-In the static potential (top left), the density remains high along the center of the stream, but for some of the barred potentials the density drops sharply because of chaotic stream-fanning.
-Vertical, dashed lines show the approximate extent of the densest part of the stream visible in main-sequence stars \citep[the segment originally detected in ][]{bernard14}.}
+\includegraphics[width=1.2\textwidth]{figures/mockstream0}
+\caption{ Sky position, distance, and line-of-sight velocity in heliocentric, Galactic coordinates for star particles (points) from the maximum-likelihood mock streams in each potential.
+Vertical, dashed lines show the approximate extent of the densest part of the stream visible in main-sequence stars \citep[the segment originally detected in ][]{bernard14}.
+\changes{Star particles are colored by the time of stripping relative to present-day.
+Green, square points show the four ``fanned'' stars.}}
 \label{fig:mockstream0}
@@
 \end{figure*}
+\end{landscape}
 
+\begin{landscape}
 \begin{figure*}[p]
 \begin{center}
-\includegraphics[width=\textwidth]{figures/mockstream1}
+\includegraphics[width=1.2\textwidth]{figures/mockstream1}
 \caption{ Same as Figure~\ref{fig:mockstream0} for the other five barred potentials. }
@@
 \end{figure*}
+\end{landscape}
+
+\begin{landscape}
+\begin{figure*}[p]
+\begin{center}
+\includegraphics[width=1.2\textwidth]{figures/mockstream-pm0}
+\caption{ Proper motion components in Galactic coordinates for star particles (points) from the maximum-likelihood mock streams in each potential.
+Star particles are colored by the time of stripping relative to present-day.
+Green, square points show the four ``fanned'' stars.}
+\label{fig:mockstream-pm0}
+\end{center}
+\end{figure*}
+\end{landscape}
+
+\begin{landscape}
+\begin{figure*}[p]
+\begin{center}
+\includegraphics[width=1.2\textwidth]{figures/mockstream-pm1}
+\caption{ Same as Figure~\ref{fig:mockstream-pm0} for the other five barred potentials. }
+\label{fig:mockstream-pm1}
+\end{center}
+\end{figure*}
+\end{landscape}
+
+\begin{landscape}
+\begin{figure*}[p]
+\begin{center}
+\includegraphics[width=1.3\textwidth]{figures/mockstream-density}
+\caption{ Surface density of mock stream star particles in each potential.
+Contours are spaced logarithmically from $10^{-2}$ to $10$ particles per sq.
+deg---that is, each color represents a factor of 10 difference in surface density.
+In the static potential (top left), the density remains high along the center of the stream, but for some of the barred potentials the density drops sharply because of chaotic stream-fanning.
+}
+\label{fig:mockstream-density}
+\end{center}
+\end{figure*}
+\end{landscape}
 
@@
 
+\begin{landscape}
 \begin{figure*}[p]
 \begin{center}
-\includegraphics[width=\textwidth]{figures/densitymaps}
+\includegraphics[width=1.3\textwidth]{figures/densitymaps}
 \caption{ Simulated maps of the 2D density of star particles from the maximum-likelihood mock streams in each potential with a noisy background of stars binned into 10' by 10' pixels.
@@
 \end{figure*}
+\end{landscape}
 
@@
 
+\begin{landscape}
 \begin{figure*}[p]
 \begin{center}
-\includegraphics[width=\textwidth]{figures/mockstream-xyz}
+\includegraphics[width=1.3\textwidth]{figures/mockstream-xyz}
 \caption{ Star particles (grey points) from mock streams generated on the mean orbits in each potential model shown in projections of Galatocentric, Cartesian coordinates.

### 2016-04-15 ms.tex
whitespace and new line in section 3

@@
 We have tried computing the frequency diffusion rate for these orbits as an independent check of their chaotic timescale but have found that, over consecutive integration windows, the frequency recovery fails or is unreliable because the frequency spectrum changes dramatically over timescales of $\approx$10--20 orbital periods.
+\changes{To understand whether time-dependence or the triaxiality of the bar is the dominant cause of chaos, we have also performed the orbit-fit and Lyapunov time experiments in a potential with a fixed bar ($\Omega_p = 0$) at the same bar angles explored above.
+We find that the Lyapunov times for orbits fit to the stream in these potentials are consistent with being regular.
+Computing the frequency diffusion rate instead shows that these orbits are likely weakly chaotic: the frequency diffusion times \citep{apw15-chaos} are $\approx$$10^4~{\rm Gyr}$.
+We therefore conclude that the time-dependence of the bar is the primary source of chaos in the inner Galaxy.}
 

### 2016-04-17 ms.tex
shorten abstract

@@
 \begin{abstract}
-The Ophiuchus stellar stream is peculiar: (1) its length is short given the age of its constituent stars, and (2) several probable member stars that lie close in both sky position and velocity have dispersions in these dimensions that far exceed those seen within the stream.
-The stream's proximity to the Galactic center suggests that the bar must have a significant influence on its dynamical history: The triaxiality and time-dependence of the bar may generate chaotic orbits in the vicinity of the stream that can greatly affect its morphology.
+The Ophiuchus stellar stream is peculiar: (1) its length is short given the age of its constituent stars, and (2) several probable member stars have dispersions in sky position and velocity that far exceed those seen within the stream.
+The stream's proximity to the Galactic center suggests that its dynamical history is significantly influenced by the Galactic bar.
 We explore this hypothesis with models of stream formation along orbits consistent with Ophiuchus' properties in a Milky Way potential model that includes a rotating bar.
-We find that in all choices for the rotation parameters of the bar, orbits fit to the stream are strongly chaotic.
+In all choices for the rotation parameters of the bar, orbits fit to the stream are strongly chaotic.
 Mock streams generated along these orbits qualitatively match the observed properties of the stream: because of chaos, stars stripped early generally form low-density, high-dispersion ``fans'' leaving only the most recently disrupted material detectable as a strong over-density.
-Our models predict that there should be more low-surface-brightness tidal debris than detected so far, likely with a complex phase-space morphology.
-The existence of or lack of these features around the Ophiuchus stream would provide an interesting constraint on the properties of the Milky Way bar and would help distinguish between formation scenarios for the stream.
+Our models predict that there should be a significant amount of low-surface-brightness tidal debris around the stream with a complex phase-space morphology.
+The existence of or lack of these features could provide interesting constraints on the Milky Way bar and would rule out formation scenarios for the stream.
 This is the first time that chaos has been used to explain the properties of a stellar stream and is the first demonstration of the dynamical importance of chaos in the Galactic halo.
-The existence of long, thin streams around the Milky Way---presumably formed along non- or weakly-chaotic orbits---may represent only a subset of the total population of disrupted satellites.
+The existence of long, thin streams around the Milky Way, presumably formed along non- or weakly-chaotic orbits, may represent only a subset of the total population of disrupted satellites.
 \end{abstract}
